### 2024-11-20 wp/tables.tex

%l !Rnw weave = knitr

\documentclass[11pt]{article}

\usepackage[]{graphicx} % omit 'demo' option in real doc.
\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\usepackage{epstopdf}
\usepackage{float}
\epstopdfDeclareGraphicsRule{.tif}{png}{.png}{convert #1 \OutputFile}
\AppendGraphicsExtensions{.tif}

\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

\usepackage{hyperref}
\hypersetup{
        colorlinks=true,
        breaklinks,
        allcolors=[RGB]{128,0,0}
}

\let\oldFootnote\footnote
\newcommand\nextToken\relax

\renewcommand\footnote[1]{%
    \oldFootnote{#1}\futurelet\nextToken\isFootnote}

\newcommand\isFootnote{%
    \ifx\footnote\nextToken\textsuperscript{,}\fi}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{crimson}

\usepackage{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\usepackage{url}
\usepackage{dcolumn}
\usepackage{ctable}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{rotating}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{seqsplit}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{soul}
\usepackage{float}
%\usepackage[multiple]{footmisc}
%\usepackage{longtable}
\usepackage{adjustbox}
\usepackage{rotating}

\usepackage{bbding} % checkmark symbol
\usepackage{comment} % checkmark symbol

\renewcommand{\textfraction}{0.05}
\renewcommand{\topfraction}{0.8}
\renewcommand{\bottomfraction}{0.8}
\renewcommand{\floatpagefraction}{0.75}

%Bibliography
\usepackage{natbib}
%\usepackage[autostyle]{csquotes}
%\usepackage[backend=bibtex, style=authoryear, natbib=true]{biblatex}
%\addbibresource{CCH.bib}

%%For Hypotheses and Subhypotheses
%\usepackage{ntheorem}
\newtheorem{hyp}{Hypothesis}
\makeatletter
\newcounter{subhyp}
\let\savedc@hyp\c@hyp
\newenvironment{subhyp}
 {%
  \setcounter{subhyp}{0}%
  \stepcounter{hyp}%
  \edef\saved@hyp{\thehyp}% Save the current value of hyp
  \let\c@hyp\c@subhyp     % Now hyp is subhyp
  \renewcommand{\thehyp}{\saved@hyp\alph{hyp}}%
 }
 {}
\newcommand{\normhyp}{%
  \let\c@hyp\savedc@hyp % revert to the old one
  \renewcommand\thehyp{\arabic{hyp}}%
}
\makeatother
%%%

\setcounter{MaxMatrixCols}{10}

\theoremstyle{plain}
\newtheorem{prop}{\protect\propositionname}
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}
\makeatother
\providecommand{\examplename}{Example}
\providecommand{\propositionname}{Proposition}
\providecommand{\theoremname}{Theorem}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\bb}{\begin{block}}
\newcommand{\eb}{\end{block}}
\newcommand{\bmath}{\begin{eqnarray}}
\newcommand{\emath}{\end{eqnarray}}
\newcommand{\bmathnn}{\begin{eqnarray*}}
\newcommand{\emathnn}{\end{eqnarray*}}
\newtheorem{theorem}{{Theorem}}%[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{prediction}{Prediction}
\newtheorem{open_question}{Open question}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{claim}[theorem]{{Claim}}
\newtheorem{example}[theorem]{{Example}}
\newtheorem{corollary}[theorem]{{Corollary}}
\newtheorem{remark}[theorem]{{Remark}}
\newtheorem{assumptions}[theorem]{{Assumptions}}
\newtheorem*{thm1star}{{Theorem $\mathbf{1^*}$}}
\newtheorem*{thm2star}{{Theorem $\mathbf{2^*}$}}
\newtheorem*{theorem*}{Theorem}
\newtheorem*{lemma*}{Lemma}

%For nice tables
\usepackage{booktabs}
\usepackage{array}
\usepackage{threeparttable}
\usepackage{tabulary}
\newcolumntype{M}[1]{{\centering\arraybackslash}m{#1}}

%Expectation opperator
\newcommand{\Expect}{{\rm I\kern-.3em E}}

%Significance commands
\newcommand*{\SuperScriptSameStyle}[1]{%
  \ensuremath{%
    \mathchoice
      {{}^{\displaystyle #1}}%
      {{}^{\textstyle #1}}%
      {{}^{\scriptstyle #1}}%
      {{}^{\scriptscriptstyle #1}}%
  }%
}

\newcommand*{\oneS}{\SuperScriptSameStyle{*}}
\newcommand*{\twoS}{\SuperScriptSameStyle{**}}
\newcommand*{\threeS}{\SuperScriptSameStyle{*{*}*}}
\newcommand{\vh}[1]{\textcolor{red}{(VH: #1)}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

\begin{document}

\title{Internet and Nation Building in Africa\\  \vspace{0.75cm} \underline{Preliminary and incomplete }}

\vspace{1cm}

\author{
\textbf{Illan Barriola} \\ CRED -- Paris II \\
\and \textbf{R\'{e}dha Chaba}\\ LEMMA -- Paris II \\
}

\date{ \today}

\maketitle

\begin{table}[H]
    \sffamily
    \caption{{Effect of distance from the capital on political trust}}
    \centering
    %\footnotesize
    \resizebox{12cm}{!}{
    \begin{tabular}{@{\extracolsep{5pt}} l c c c c}
    \\
    \toprule
    \toprule
    &  \multicolumn{4}{c}{{OLS}}  \\
    \cmidrule(r){2-5}
    &  \multicolumn{4}{c}{{Political trust}}  \\
    \cmidrule(r){2-5}
    & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}}\\
        \cmidrule(r){2-3}
        \cmidrule(r){4-5}
        & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}}\\
     \midrule

     Distance from the capital&       0.284***&       0.270***&       0.681***&       0.397* \\
     \medskip
     &      (0.04)   &      (0.03)   &      (0.16)   &      (0.21)  \\
     \midrule
     \smallskip
    Individual \& regional controls  & Yes & Yes & Yes & Yes  \\
    \smallskip
    Country controls & Yes& No& Yes& No\\
    \smallskip
    Round FE & Yes & No& Yes & No\\
    \smallskip
    Country X Round FE       & No & Yes& No & Yes\\
    \smallskip
    Ethnic homeland FE & No & No & Yes& Yes\\
    \smallskip
    Observations           & 98,235   &       98,235   &        4,189   &        4,189   \\
    Adjusted-R$^2$        &       0.158   &       0.197   &       0.159   &       0.189   \\

                          \bottomrule
    \multicolumn{5}{p{12.5cm}}{\footnotesize \emph{Notes}: The border sample includes individuals residing within a 20-kilometer buffer around a country border that overlaps with a historical ethnic homeland, as defined by Murdock (1959). Robust standard errors clustered at the region x round level  for the base sample and ethnic homeland x region x round level for the border sample are in parentheses. The set of individual controls
    includes values of: normalized distance from the largest non-capital city, age, age squared, sex,
    education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, radio news consumption, newspaper news consumption. The set of regional controls includes values of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
    \end{tabular}
    }
    \end{table}

    \begin{table}[H]
        \sffamily
        \caption{Effect of internet coverage on political trust by distance}
        \centering
        %\footnotesize
        \resizebox{17cm}{!}{
        \begin{tabular}{@{\extracolsep{5pt}} l c c c c c c}
        \\
        \toprule
        \toprule
        & \multicolumn{5}{c}{{Base sample}}\\
        \cmidrule(r){2-6}
        & \multicolumn{1}{c}{{OLS}} & \multicolumn{3}{c}{{First Stage}} & \multicolumn{1}{c}{{2SLS}}\\
            \cmidrule(r){2-2}
            \cmidrule(r){3-5}
            \cmidrule(r){6-6}
            & \multicolumn{1}{c}{{Political trust}} & \multicolumn{1}{c}{{Internet coverage}} & \multicolumn{1}{c}{{}} &  \multicolumn{1}{c}{{Distance $\times$ Internet coverage}} & \multicolumn{1}{c}{{Political trust}}\\
            \cmidrule(r){2-2}
            \cmidrule(r){3-3}
            \cmidrule(r){5-5}    
            \cmidrule(r){6-6}    
            & \multicolumn{1}{c}{{(1)}} & \multicolumn{1}{c}{{(2)}} & \multicolumn{1}{c}{{}} &  \multicolumn{1}{c}{{(3)}} & \multicolumn{1}{c}{{(4)}}\\

            \midrule

    Distance from the capital &0.463***&&&& 1.361***\\      
    \smallskip
    &      (0.05)   &&&&     (0.49)   \\
    Internet coverage     &       0.029   &&&&       1.470**  \\
    \smallskip
    & (0.06)   &&&&      (0.64)    \\ 
    \smallskip
    Distance from the capital $\times$ Internet coverage&      -0.480***&&&               &       -2.295**\\
    \smallskip  &       (0.10)   &&&               &      (1.09) \\

   Lightning strikes && -0.002*** && -0.000  \\
    \smallskip&& (0.00) && (0.00) \\
    Distance from the capital city $\times$ Lightning strikes&& 0.000 && -0.001**\\
    \medskip&& (0.00) && (0.00)\\

         \midrule
        SW \emph{F} - Lightning strikes &-&-& 19.66&- &-&-\\
        \smallskip
        SW \emph{F} - Distance $\times$ Lightning strikes &-&-& 11.13 &-&-&-\\
        \smallskip
        Individual \& regional controls  & Yes & Yes &&  Yes & Yes\\
        \smallskip
        Country X Round FE       & Yes & Yes&- & Yes & Yes \\
        \smallskip
        Observations       &       98,235    &99,414 && 99,414&       98,235  \\
        Adjusted-R$^2$    &       0.198    &-&-&-&  \\
                              \bottomrule
        \multicolumn{6}{p{27cm}}{\footnotesize \emph{Notes}: % Set font size to 8pt with 10pt line spacing
        \emph{Notes}: Robust standard errors clustered at the region x round level. The set of individual controls
        includes values of: normalized distance from the largest non-capital city, age, age squared, sex,
        education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, newspaper news consumption, radio consumption. The set of regional controls includes values of: nighttime light, population density, region area, president birthplace dummy. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
    \end{tabular}}
        \end{table}

        \begin{table}[H]
            \sffamily
            \caption{{Other outcomes}}
            \centering
            %\footnotesize
            \resizebox{18cm}{!}{
            \begin{tabular}{@{\extracolsep{5pt}} l c c}
            \\
            \toprule
            \toprule
            & \multicolumn{2}{c}{2SLS}\\
            \cmidrule(r){2-3}
            &\multicolumn{1}{c}{{Vote against ruling party}}&  \multicolumn{1}{c}{{Country performance}}  \\
            \cmidrule(r){2-2}
            \cmidrule(r){3-3}
            & \multicolumn{1}{c}{{Base sample}} & \multicolumn{1}{c}{{Base sample}}\\
                \cmidrule(r){2-2}
                \cmidrule(r){3-3}
                & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}\\
             \midrule

             Distance from the capital&   -1.306***&       1.560***\\
             \smallskip
             &      (0.40)   &      (0.52)   \\
             Internet coverage     &      -1.370***&       1.565**  \\
             \smallskip
             &      (0.47)   &      (0.62)   \\
             Distance from the capital $\times$ Internet coverage&       2.614***&      -3.065***\\
             \medskip
             &       (0.92)   &      (1.16)   \\
             \midrule
             \smallskip
            Individual \& regional controls  & Yes & Yes\\
            \smallskip
            Country X Round FE       & Yes& Yes\\
            \smallskip
            Observations      &       99,414   &       98,512  \\            \bottomrule
            \multicolumn{3}{p{18cm}}{\footnotesize \emph{Notes}: Robust standard errors clustered at the region x round level are in parentheses. The set of individual controls
            includes values of: normalized distance from the largest non-capital city, age, age squared, sex,
            education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, newspaper news consumption, radio consumption. The set of regional controls includes values of: nighttime light, population density, region area, president birthplace dummy. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
            \end{tabular}
            }
            \end{table}

            \begin{table}[H]
              \sffamily
              \caption{{Media and institutions freedom}}
              \centering
              %\footnotesize
              \resizebox{16cm}{!}{
              \begin{tabular}{@{\extracolsep{5pt}} l c c c c}
              \\
              \toprule
              \toprule
              & \multicolumn{4}{c}{2SLS: Political trust}\\
              \cmidrule(r){2-5}
              & \multicolumn{4}{c}{{Base sample}}\\
              \cmidrule(r){2-5}
              &\multicolumn{2}{c}{{Media}}&  \multicolumn{2}{c}{{Institutions}}\\
              \cmidrule(r){2-3}
                  \cmidrule(r){4-5}
                  & \multicolumn{1}{c}{{Free}} & \multicolumn{1}{c}{{Captured}}& \multicolumn{1}{c}{{Free}} & \multicolumn{1}{c}{{Captured}}\\
                  \cmidrule(r){2-2}
                  \cmidrule(r){3-3}
                  \cmidrule(r){4-4}
                  \cmidrule(r){5-5}
                  & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}}\\
               \midrule

               Distance from the capital&     -4.922   &       1.133*** &       0.078   &       2.151***\\
               \smallskip
               &     (11.77)   &      (0.41)  &      (0.83)   &      (0.77)\\
               Internet coverage        &      -3.475   &       1.220* &       0.112   &       2.964***\\
               \smallskip
               &     (10.01)   &      (0.68)  &      (1.37)   &      (1.06) \\
            Distance from the capital $\times$ Internet coverage&       9.210   &      -2.160* &       0.391   &      -4.202**\\
            \medskip
            &     (21.19)   &      (1.22) &      (1.59)   &      (1.90)\\

               \midrule
               \smallskip
              Individual \& regional controls  & Yes & Yes & Yes & Yes\\
              \smallskip
              Country X Round FE       & Yes& Yes & Yes& Yes\\
              \smallskip
              Observations          &       48,174   &       50,061 &       45,238   &       52,997 \\
              \bottomrule
              \multicolumn{5}{p{15cm}}{\footnotesize \emph{Notes}: Robust standard errors clustered at the region x round level are in parentheses. The set of individual controls
              includes values of: normalized distance from the largest non-capital city, age, age squared, sex,
              education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, newspaper news consumption, radio consumption. The set of regional controls includes values of: nighttime light, population density, region area, president birthplace dummy. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
              \end{tabular}
              }
              \end{table}

              \begin{table}[H]
                \sffamily
                \caption{{Individual heterogeneity}}
                \centering
                %\footnotesize
                \resizebox{16cm}{!}{
                \begin{tabular}{@{\extracolsep{5pt}} l c c c c c c c c}
                \\
                \toprule
                \toprule
                & \multicolumn{8}{c}{2SLS: Political trust}\\
                \cmidrule(r){2-9}
                & \multicolumn{8}{c}{{Base sample}}\\
                \cmidrule(r){2-9}
                &\multicolumn{2}{c}{{Education}}&  \multicolumn{2}{c}{{Urban/Rural}}&  \multicolumn{4}{c}{{Age}}\\
                \cmidrule(r){2-3}
                    \cmidrule(r){4-5}
                    \cmidrule(r){6-9}
                    & \multicolumn{1}{c}{{$<$ Secondary}} & \multicolumn{1}{c}{{$\geq$ Secondary}}& \multicolumn{1}{c}{{Urban}} & \multicolumn{1}{c}{{Rural}} & \multicolumn{1}{c}{{18-25}}& \multicolumn{1}{c}{{26-35}}& \multicolumn{1}{c}{{36-44}}& \multicolumn{1}{c}{{45+}}\\
                    \cmidrule(r){2-2}
                    \cmidrule(r){3-3}
                    \cmidrule(r){4-4}
                    \cmidrule(r){5-5}
                    \cmidrule(r){6-6}
                    \cmidrule(r){7-7}
                    \cmidrule(r){8-8}
                    \cmidrule(r){9-9}
                    & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}} &  \multicolumn{1}{c}{{(5)}} &  \multicolumn{1}{c}{{(6)}} &  \multicolumn{1}{c}{{(7)}} &  \multicolumn{1}{c}{{(8)}}\\
                 \midrule

                 Distance from the capital&       1.208***&       1.394   &       1.133** &       1.961***&       1.295** &       0.965   &       1.438** &       1.723*** \\
                 \smallskip
                 &      (0.31)   &      (1.17)   &      (0.52)   &      (0.70)   &      (0.57)   &      (0.63)   &      (0.56)   &      (0.59)   \\
                 Internet coverage  &       1.706*  &       1.271   &       0.397   &       3.374***&       1.439*  &       0.982   &       1.236*  &       2.540***\\
                 \smallskip
                 &      (0.92)   &      (1.32)   &      (0.63)   &      (1.18)   &      (0.74)   &      (0.76)   &      (0.72)   &      (0.87)   \\
                   Distance from the capital $\times$ Internet coverage&      -1.909** &      -2.422   &      -1.785*  &      -3.447*  &      -1.994   &      -1.477   &      -2.577** &      -3.103**  \\
                   \medskip
                   &      (0.83)   &      (2.41)   &      (0.98)   &      (1.81)   &      (1.24)   &      (1.38)   &      (1.28)   &      (1.39)   \\
                 \midrule
                 \smallskip
                Individual \& regional controls  & Yes & Yes & Yes & Yes& Yes & Yes & Yes & Yes\\
                \smallskip
                Country X Round FE       & Yes& Yes & Yes& Yes & Yes& Yes & Yes& Yes\\
                \smallskip
                Observations         &       36,389   &       62,039   &       37,348   &       60,887   &       25,442   &       29,913   &       20,074   &       23,246   \\
                \bottomrule
                \multicolumn{9}{p{24cm}}{\footnotesize \emph{Notes}: Robust standard errors clustered at the region x round level are in parentheses. The set of individual controls
                includes values of: normalized distance from the largest non-capital city, age, age squared, sex,
                education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, newspaper news consumption, radio consumption. The set of regional controls includes values of: nighttime light, population density, region area, president birthplace dummy. Education, rural/urban, and age controls are omitted from columns (1-2), (3-4), and (5-8), respectively.*** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
                \end{tabular}
                }
                \end{table}

            \begin{table}[H]
              \sffamily
              \caption{{Generalized trust}}
              \centering
              %\footnotesize
              \resizebox{16cm}{!}{
              \begin{tabular}{@{\extracolsep{5pt}} l c c}
              \\
              \toprule
              \toprule
              & \multicolumn{2}{c}{2SLS}\\
              \cmidrule(r){2-3}
              & \multicolumn{2}{c}{{Base sample: rounds 5 $\&$ 8}}\\
              \cmidrule(r){2-3}
                  & \multicolumn{1}{c}{{Political trust}} & \multicolumn{1}{c}{{Generalized trust}}\\
                  \cmidrule(r){2-2}
                  \cmidrule(r){3-3}
                  & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}\\
               \midrule
               Distance from the capital&       1.358** &      -0.157   \\
               \smallskip
               &      (0.62)   &      (0.17)    \\
               Internet coverage    &       1.508   &      -0.101    \\
               \smallskip
               &      (0.94)   &      (0.23)   \\
               Distance from the capital $\times$ Internet coverage&      -2.161*  &       0.542   \\
               \medskip
               &      (1.22)   &      (0.33)   \\

               \midrule
               \smallskip
              Individual \& regional controls  & Yes & Yes\\
              \smallskip
              Country X Round FE       & Yes& Yes\\
              \smallskip
              Observations         &     49,404   &       49,362\\
              \bottomrule
              \multicolumn{3}{p{14cm}}{\footnotesize \emph{Notes}: Robust standard errors clustered at the region x round level are in parentheses. The set of individual controls
              includes values of: normalized distance from the largest non-capital city, age, age squared, sex,
              education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, newspaper news consumption, radio consumption. The set of regional controls includes values of: nighttime light, population density, region area, president birthplace dummy. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
              \end{tabular}
              }
              \end{table}

              \begin{table}[H]
                \sffamily
                \caption{{Effect of distance from the capital on institutional trust (President, Parliament, Electoral commission)}}
                \centering
                %\footnotesize
                \resizebox{16cm}{!}{
                \begin{tabular}{@{\extracolsep{5pt}} l c c c c c c}
                \\
                \toprule
                \toprule
                &  \multicolumn{6}{c}{{Trust in institutions (President, Parliament, Electoral commission)}}  \\
                \cmidrule(r){2-7}
                &  \multicolumn{5}{c}{{OLS}}&  \multicolumn{1}{c}{{2SLS}}  \\
                \cmidrule(r){2-6}
                \cmidrule(r){7-7}
                & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}}& \multicolumn{1}{c}{{Base sample}}& \multicolumn{1}{c}{{Base sample}}\\
                    \cmidrule(r){2-3}
                    \cmidrule(r){4-5}
                    \cmidrule(r){6-6}
                    \cmidrule(r){7-7}
                    & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}}&  \multicolumn{1}{c}{{(5)}}&  \multicolumn{1}{c}{{(6)}}\\
                 \midrule

                 Distance from the capital&    0.295***&       0.301***&       0.563***&       0.306 &       0.483***&       1.738***\\
                 \smallskip
                 &      (0.04)   &      (0.03)   &      (0.16)   &      (0.20) &      (0.05)   &      (0.58) \\
                 Internet coverage &&&&&       0.016   &       2.035*** \\
                 \smallskip

                 &&&&&      (0.06)   &      (0.71)\\
                 Distance from the capital $\times$ Internet coverage &&&&&      -0.458***&-2.992**\\
                 \medskip
                 &&&&&      (0.10) & (1.32)\\
                 \midrule
                 \smallskip
                Individual \& regional controls  & Yes & Yes & Yes & Yes& Yes & Yes  \\
                \smallskip
                Country controls & Yes& No& Yes& No& No& No\\
                \smallskip
                Round FE & Yes & No& Yes & No& No& No\\
                \smallskip
                Country X Round FE  & No & Yes& No & Yes& Yes& Yes\\
                \smallskip
                Ethnic homeland FE & No & No & Yes& Yes& No& No\\
                \smallskip
                Observations           &   98,499   &       98,499   &        4,202   &        4,202   & 98,499   &       98,499 \\
                Adjusted-R$^2$        &        0.164   &       0.198   &       0.148   &       0.184  & 0.200   &       - \\

                                      \bottomrule
                \multicolumn{7}{p{20cm}}{\footnotesize \emph{Notes}: The border sample includes individuals residing within a 20-kilometer buffer around a country border that overlaps with a historical ethnic homeland, as defined by Murdock (1959). Robust standard errors clustered at the region x round level  for the base sample and ethnic homeland x region x round level for the border sample are in parentheses. The set of individual controls
                includes values of: normalized distance from the largest non-capital city, age, age squared, sex,
                education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, radio news consumption, newspaper news consumption. The set of regional controls includes values of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
                \end{tabular}
                }
                \end{table}

                \begin{table}[H]
                  \sffamily
                  \caption{{Effect of distance from the capital on trust in president}}
                  \centering
                  %\footnotesize
                  \resizebox{16cm}{!}{
                  \begin{tabular}{@{\extracolsep{5pt}} l c c c c c c}
                  \\
                  \toprule
                  \toprule
                  &  \multicolumn{6}{c}{{Trust in president}}  \\
                  \cmidrule(r){2-7}
                  &  \multicolumn{5}{c}{{OLS}}&  \multicolumn{1}{c}{{2SLS}}  \\
                  \cmidrule(r){2-6}
                  \cmidrule(r){7-7}
                  & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}}& \multicolumn{1}{c}{{Base sample}}& \multicolumn{1}{c}{{Base sample}}\\
                      \cmidrule(r){2-3}
                      \cmidrule(r){4-5}
                      \cmidrule(r){6-6}
                      \cmidrule(r){7-7}
                      & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}}&  \multicolumn{1}{c}{{(5)}}&  \multicolumn{1}{c}{{(6)}}\\
                   \midrule

                   Distance from the capital&       0.287***&       0.262***&       0.799***&       0.343   &       0.473***&       1.517**\\
                   \smallskip
                   &      (0.05)   &      (0.03)   &      (0.19)   &      (0.25)   &      (0.06)   &      (0.63) \\
                   Internet coverage &&&&       &       0.047   &       1.659**  \\
                   \smallskip
                   &&&&    &      (0.07)   &      (0.81) \\
                   Distance from the capital $\times$ Internet coverage &&&&&      -0.521***&-2.654*\\
                   \medskip
                   &&&&&      (0.12)  & (1.41) \\
                   \midrule
                   \smallskip
                  Individual \& regional controls  & Yes & Yes & Yes & Yes& Yes & Yes  \\
                  \smallskip
                  Country controls & Yes& No& Yes& No& No& No\\
                  \smallskip
                  Round FE & Yes & No& Yes & No& No& No\\
                  \smallskip
                  Country X Round FE  & No & Yes& No & Yes& Yes& Yes\\
                  \smallskip
                  Ethnic homeland FE & No & No & Yes& Yes& No& No\\
                  \smallskip
                  Observations            &       97,261   &       97,261   &        4,149   &        4,149   &       97,261   &       97,261  \\
                  Adjusted-R$^2$        &       0.137   &       0.180   &       0.180   &       0.215   &       0.182    &       - \\

                                        \bottomrule
                  \multicolumn{7}{p{20cm}}{\footnotesize \emph{Notes}: The border sample includes individuals residing within a 20-kilometer buffer around a country border that overlaps with a historical ethnic homeland, as defined by Murdock (1959). Robust standard errors clustered at the region x round level  for the base sample and ethnic homeland x region x round level for the border sample are in parentheses. The set of individual controls
                  includes values of: normalized distance from the largest non-capital city, age, age squared, sex,
                  education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, radio news consumption, newspaper news consumption. The set of regional controls includes values of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
                  \end{tabular}
                  }
                  \end{table}

                  \begin{table}[H]
                    \sffamily
                    \caption{{Effect of distance from the capital on trust in parliament}}
                    \centering
                    %\footnotesize
                    \resizebox{16cm}{!}{
                    \begin{tabular}{@{\extracolsep{5pt}} l c c c c c c}
                    \\
                    \toprule
                    \toprule
                    &  \multicolumn{6}{c}{{Trust in parliament}}  \\
                    \cmidrule(r){2-7}
                    &  \multicolumn{5}{c}{{OLS}}&  \multicolumn{1}{c}{{2SLS}}  \\
                    \cmidrule(r){2-6}
                    \cmidrule(r){7-7}
                    & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}}& \multicolumn{1}{c}{{Base sample}}& \multicolumn{1}{c}{{Base sample}}\\
                        \cmidrule(r){2-3}
                        \cmidrule(r){4-5}
                        \cmidrule(r){6-6}
                        \cmidrule(r){7-7}
                        & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}}&  \multicolumn{1}{c}{{(5)}}&  \multicolumn{1}{c}{{(6)}}\\
                     \midrule

                     Distance from the capital&       0.282***&       0.279***&       0.560***&       0.471** &       0.453***&       1.124***\\
                     \smallskip
                     &      (0.04)   &      (0.03)   &      (0.17)   &      (0.21)   &      (0.05)   &      (0.39) \\
                     Internet coverage &&&&       &        0.007   &       1.169**  \\
                     \smallskip
                     &&&&    &      (0.06)   &      (0.52)   \\
                     Distance from the capital $\times$ Internet coverage &&&&&      -0.442***& -1.776**\\
                     \medskip
                     &&&&&     (0.10)   &  (0.89) \\
                     \midrule
                     \smallskip
                    Individual \& regional controls  & Yes & Yes & Yes & Yes& Yes & Yes  \\
                    \smallskip
                    Country controls & Yes& No& Yes& No& No& No\\
                    \smallskip
                    Round FE & Yes & No& Yes & No& No& No\\
                    \smallskip
                    Country X Round FE  & No & Yes& No & Yes& Yes& Yes\\
                    \smallskip
                    Ethnic homeland FE & No & No & Yes& Yes& No& No\\
                    \smallskip
                    Observations            &      96,182   &       96,182   &        4,111   &        4,111   &       96,182   &       96,182  \\
                    Adjusted-R$^2$        &        0.130   &       0.160   &       0.102   &       0.127   &       0.161   &       - \\

                                          \bottomrule
                    \multicolumn{7}{p{20cm}}{\footnotesize \emph{Notes}: The border sample includes individuals residing within a 20-kilometer buffer around a country border that overlaps with a historical ethnic homeland, as defined by Murdock (1959). Robust standard errors clustered at the region x round level  for the base sample and ethnic homeland x region x round level for the border sample are in parentheses. The set of individual controls
                    includes values of: normalized distance from the largest non-capital city, age, age squared, sex,
                    education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, radio news consumption, newspaper news consumption. The set of regional controls includes values of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
                    \end{tabular}
                    }
                    \end{table}

                  \begin{table}[H]
                    \sffamily
                    \caption{{Effect of distance from the capital on trust in electoral commission}}
                    \centering
                    %\footnotesize
                    \resizebox{16cm}{!}{
                    \begin{tabular}{@{\extracolsep{5pt}} l c c c c c c}
                    \\
                    \toprule
                    \toprule
                    &  \multicolumn{6}{c}{{Trust in electoral commission}}  \\
                    \cmidrule(r){2-7}
                    &  \multicolumn{5}{c}{{OLS}}&  \multicolumn{1}{c}{{2SLS}}  \\
                    \cmidrule(r){2-6}
                    \cmidrule(r){7-7}
                    & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}}& \multicolumn{1}{c}{{Base sample}}& \multicolumn{1}{c}{{Base sample}}\\
                        \cmidrule(r){2-3}
                        \cmidrule(r){4-5}
                        \cmidrule(r){6-6}
                        \cmidrule(r){7-7}
                        & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}}&  \multicolumn{1}{c}{{(5)}}&  \multicolumn{1}{c}{{(6)}}\\
                     \midrule

                     Distance from the capital&       0.319***&       0.365***&       0.366*  &       0.206   &       0.529***&       2.588***\\
                     \smallskip
                     &     (0.04)   &      (0.03)   &      (0.19)   &      (0.23)   &      (0.06)   &      (0.84) \\
                     Internet coverage &&&&       &        0.003   &       3.252***  \\
                     \smallskip
                     &&&&    &      (0.07)   &      (0.95)   \\
                     Distance from the capital $\times$ Internet coverage &&&&&     -0.418***&  -4.606**\\
                     \medskip
                     &&&&&      (0.11)    &  (1.93) \\
                     \midrule
                     \smallskip
                    Individual \& regional controls  & Yes & Yes & Yes & Yes& Yes & Yes  \\
                    \smallskip
                    Country controls & Yes& No& Yes& No& No& No\\
                    \smallskip
                    Round FE & Yes & No& Yes & No& No& No\\
                    \smallskip
                    Country X Round FE  & No & Yes& No & Yes& Yes& Yes\\
                    \smallskip
                    Ethnic homeland FE & No & No & Yes& Yes& No& No\\
                    \smallskip
                    Observations            &      94,926   &       94,926   &        4,035   &        4,035   &       94,926   &       94,926   \\
                    Adjusted-R$^2$        &         0.118   &       0.144   &       0.094   &       0.137   &       0.145   &       - \\

                                          \bottomrule
                    \multicolumn{7}{p{20cm}}{\footnotesize \emph{Notes}: The border sample includes individuals residing within a 20-kilometer buffer around a country border that overlaps with a historical ethnic homeland, as defined by Murdock (1959). Robust standard errors clustered at the region x round level  for the base sample and ethnic homeland x region x round level for the border sample are in parentheses. The set of individual controls
                    includes values of: normalized distance from the largest non-capital city, age, age squared, sex,
                    education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, radio news consumption, newspaper news consumption. The set of regional controls includes values of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
                    \end{tabular}
                    }
                    \end{table}

\begin{table}[H]
  \sffamily
  \caption{{Effect of distance from the capital on political trust}}
  \centering
  %\footnotesize
  \resizebox{12cm}{!}{
  \begin{tabular}{@{\extracolsep{5pt}} l c c c c c c c c}
  \\
  \toprule
  \toprule
  &  \multicolumn{8}{c}{{OLS}}  \\
  \cmidrule(r){2-9}
  &  \multicolumn{4}{c}{{Vote against ruling party}}&  \multicolumn{4}{c}{{Country performance}}  \\
  \cmidrule(r){2-5}
  \cmidrule(r){6-9}
  & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}}  & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}}\\
      \cmidrule(r){2-3}
      \cmidrule(r){4-5}
      \cmidrule(r){6-7}
      \cmidrule(r){8-9}
      & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}}&  \multicolumn{1}{c}{{(5)}}&  \multicolumn{1}{c}{{(6)}}&  \multicolumn{1}{c}{{(7)}}&  \multicolumn{1}{c}{{(8)}}\\
      \midrule

   Distance from the capital&      -0.164***&      -0.101***&      -0.462***&      -0.334***&       0.183***&       0.153***&       0.272*  &      -0.034   \\
   \medskip
   &      (0.02)   &      (0.02)   &      (0.09)   &      (0.09)   &      (0.04)   &      (0.03)   &      (0.16)   &      (0.20)   \\
   \midrule
   \smallskip
  Individual \& regional controls  & Yes & Yes & Yes & Yes& Yes & Yes & Yes & Yes  \\
  \smallskip
  Country controls & Yes& No& Yes& No& Yes& No& Yes& No\\
  \smallskip
  Round FE & Yes & No& Yes & No& Yes & No& Yes & No\\
  \smallskip
  Country X Round FE       & No & Yes& No & Yes& No & Yes& No & Yes\\
  \smallskip
  Ethnic homeland FE & No & No & Yes& Yes& No & No & Yes& Yes\\
  \smallskip
  Observations           &       99,414   &       99,414   &        4,247   &        4,247   &       98,512   &       98,512   &        4,203   &        4,203   \\
  Adjusted-R$^2$         &       0.082   &       0.147   &       0.167   &       0.249   &       0.140   &       0.161   &       0.166   &       0.184   \\
                        \bottomrule
  \multicolumn{9}{p{21cm}}{\footnotesize \emph{Notes}: The border sample includes individuals residing within a 20-kilometer buffer around a country border that overlaps with a historical ethnic homeland, as defined by Murdock (1959). Robust standard errors clustered at the region x round level  for the base sample and ethnic homeland x region x round level for the border sample are in parentheses. The set of individual controls
  includes values of: normalized distance from the largest non-capital city, age, age squared, sex,
  education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, radio news consumption, newspaper news consumption. The set of regional controls includes values of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
  \end{tabular}
  }
  \end{table}

\end{document}