### 2025-02-24 script/afrobarometer.Rmd

```{r}
setwd("~/working_paper/trust/")
getwd()
```

```{r packages, echo=FALSE, }
library(haven)
library(dplyr)
select <- dplyr::select
library(magrittr)
library(stringr)
library(geosphere)
library(purrr)
library(kableExtra)
library(tidyr)
```

# Read Afrobarometer data from different waves (rounds 5-8)

```{r echo=FALSE}
afro_w5 <- read.csv("data/data_raw/afrobarometer/afb_full_r5.csv_")
afro_w6 <- read.csv("data/data_raw/afrobarometer/afb_full_r6.csv_")
afro_w7 <- read_sav("data/data_raw/afrobarometer/R7_Merged data_34ctry_4Dec18.sav")
afro_w8 <- read_sav("data/data_raw/afrobarometer/R8.Merge_34ctry_2nov21.release.w.local.info.sav")
```

# Process Wave 5 data

```{r}
# Fix country codes: Replace MAU with MUS (Mauritius) and MRC with MAR (Morocco)

afro_w5 %<>%
  mutate(respno = case_when(
    str_starts(respno, "MAU") ~ str_replace(respno, "^MAU", "MUS"),
    TRUE ~ respno
  ))

afro_w5 %<>%
  mutate(respno = case_when(
    str_starts(respno, "MRC") ~ str_replace(respno, "^MRC", "MAR"),
    TRUE ~ respno
  ))

# Add round number and create unique identifier

afro_w5%<>%
  mutate(round=rep(5, times = length("respno")))%>%
  mutate(id_unique=paste0(respno, round))

# Rename variables to standardized names across waves

afro_w5 %<>%
  rename(
    age = q1,
    assoc = q25b,
    check_citiz = q33,
    check_parl = q36,
    commu = q26a,
    conditions_country = q3a,
    conditions_eco = q3b,
    contact_counc = q30a,
    #contact_trad
    corr_judge = q60g,
    #corr_level
    corr_local = q60d,
    corr_parl = q60b,
    corr_pres = q60a,
    #corr_trad
    disc_pol = q15,
    educ = q97,
    elec_fair = q28,
    emp = q96,
    extent = q42,
    feelnat = q85b,
    food = q8a,
    internet_news = q13d,
    internet_use = q91b,
    latitude = latitude,
    listen_counc = q62b,
    #listen_trad
    longitude = longitude,
    military_reject = q31b,
    money = q8e,
    one_man_reject = q31c,
    one_party_reject = q31a,
    paper_news = q13c,
    perf_corr = q65k,
    perf_crime = q65f,
    perf_educ = q65h,
    perf_elect = q65o,
    perf_eco = q65a,
    perf_health = q65g,
    perf_ineq = q65e,
    perf_job = q65c,
    perf_local = q71c,
    perf_parl = q71b,
    perf_poor = q65b,
    perf_pres = q71a,
    perf_price = q65d,
    perf_road = q65n,
    perf_water = q65i,
    pref_demo = q32,
    prob = q26b,
    protest = q26d,
    radio_news = q13a,
    rural = urbrur,
    satis_demo = q43,
    security = q9a,
    sex = q101,
    #sm_news
    tele_news = q13b,
    tele_pos = q90b,
    trust_elec = q59c,
    trust_gen = q87,
    trust_govpart = q59f,
    trust_local = q59e,
    trust_oppart = q59g,
    trust_parl = q59b,
    trust_pres = q59a,
    #trust_trad
    unfair_eth = q85a,
    vote = q27
  )

glimpse(afro_w5)

saveRDS(afro_w5, file="data/data_rds/afro_w5.rds")
```

# Process Wave 6 data

```{r}
# Fix country codes: Replace MAU with MUS (Mauritius) and MRC with MAR (Morocco)

afro_w6 %<>%
  mutate(respno = case_when(
    str_starts(respno, "MAU") ~ str_replace(respno, "^MAU", "MUS"),
    TRUE ~ respno
  ))

afro_w6 %<>%
  mutate(respno = case_when(
    str_starts(respno, "MOR") ~ str_replace(respno, "^MOR", "MAR"),
    TRUE ~ respno
  ))

# Add round number and create unique identifier

afro_w6%<>%
  mutate(round=rep(6, times = length("respno")))%>%
  mutate(id_unique=paste0(respno, round))

# Rename variables to standardized names across waves

afro_w6 %<>%
  rename(
    age = q1,
    assoc = q19b,
    check_citiz=q31,
    check_parl=q34,
    commu = q20a,
    conditions_country = q4a,
    conditions_eco = q4b,
    contact_counc = q24a,
    contact_trad = q24e,
    corr_judge = q53g,
    corr_level = q54,
    corr_local = q53d, 
    corr_parl = q53b,
    corr_pres = q53a, 
    corr_trad = q53h,
    disc_pol = q14,
    educ = q97,
    elec_fair = q22,
    emp = q95,
    extent = q40,
    feelnat = q88b,
    food = q8a,
    internet_news = q12d,
    internet_use = q92b,
    latitude = latitude,
    listen_counc = q59b,
    #listen_trad
    longitude = longitude,
    military_reject = q28b,
    money = q8e,
    one_man_reject = q28c,
    one_party_reject = q28a,
    paper_news = q12c,
    perf_corr=q66k,
    perf_crime=q66f,
    perf_educ=q66h,
    perf_elect=q66m,
    perf_eco = q66a,
    perf_health=q66g,
    perf_ineq=q66e,
    perf_job=q66c,
    perf_local = q68c,
    perf_parl=q68b,
    perf_poor=q66b,
    perf_pres=q68a,
    perf_price=q66d,
    perf_road=q66l,
    perf_water=q66i,
    pref_demo = q30,
    prob = q20b,
    protest = q27e,
    radio_news = q12a,
    rural = urbrur,
    satis_demo = q41,
    security = q10a,
    sex = q101,
    sm_news = q12e,
    tele_news = q12b,
    tele_pos = q91b,
    trust_elec = q52c,
    #trust_gen = 
    trust_govpart = q52f,
    trust_local = q52e,
    trust_oppart = q52g,
    trust_parl = q52b,
    trust_pres = q52a, 
    trust_trad = q52k,
    unfair_eth = q88a,
    vote = q21)

glimpse(afro_w6)

saveRDS(afro_w6, file="data/data_rds/afro_w6.rds")
```

# Process Wave 7 data

```{r}
# Fix country codes: Replace MAU with MUS (Mauritius) and MRC with MAR (Morocco)

afro_w7 %<>%
  mutate(RESPNO = case_when(
    str_starts(RESPNO, "MAU") ~ str_replace(RESPNO, "^MAU", "MUS"),
    TRUE ~ RESPNO
  ))

afro_w7 %<>%
  mutate(RESPNO = case_when(
    str_starts(RESPNO, "MOR") ~ str_replace(RESPNO, "^MOR", "MAR"),
    TRUE ~ RESPNO
  ))

# Add round number and create unique identifier

afro_w7%<>%
  mutate(respno=RESPNO)%>%
  mutate(round=rep(7, times = length("respno")))%>%
  mutate(id_unique=paste0(RESPNO, round))

# Rename variables to standardized names across waves

afro_w7 %<>%
  rename(
    age = Q1,
    assoc = Q20B,
    check_citiz=Q29,
    check_parl=Q32,
    commu = Q21A,
    conditions_country = Q4A,
    conditions_eco = Q4B,
    contact_counc = Q25A,
    contact_trad = Q25E,
    corr_judge = Q44F,
    corr_level = Q45,
    corr_local = Q44D, 
    corr_parl = Q44B,
    corr_pres = Q44A, 
    corr_trad = Q44G,
    disc_pol = Q13,
    educ = Q97,
    elec_fair = Q23,
    emp = Q94,
    extent = Q35,
    feelnat = Q85B,
    food = Q8A,
    internet_news = Q12D,
    internet_use = Q91B,
    latitude = EA_GPS_LA,
    listen_counc = Q54B,
    #listen_trad
    longitude = EA_GPS_LO,
    military_reject = Q27B,
    money = Q8E,
    one_man_reject = Q27C,
    one_party_reject = Q27A,
    paper_news = Q12C,
    perf_corr = Q56K,
    perf_crime = Q56F,
    perf_eco = Q56A,
    perf_educ = Q56H,
    perf_elect = Q56M,
    perf_health = Q56G,
    perf_ineq = Q56E,
    perf_job = Q56C,
    perf_local = Q58C,
    perf_parl = Q58B,
    perf_pres = Q58A,
    perf_price = Q56D,
    perf_road = Q56L,
    perf_trad = Q58D,
    perf_water = Q56I,
    pref_demo = Q28,
    prob = Q21B,
    protest = Q26E,
    radio_news = Q12A,
    rural = URBRUR,
    satis_demo = Q36,
    security = Q10A,
    sex = Q101,
    sm_news = Q12E,
    tele_news = Q12B,
    tele_pos = Q89B,
    trust_elec =  Q43C,
    #trust_gen = 
    trust_govpart = Q43E,
    trust_local = Q43D,
    trust_oppart = Q43F,
    trust_parl = Q43B,
    trust_pres = Q43A, 
    trust_trad = Q43J,
    unfair_eth = Q85A,
    vote = Q22)

glimpse(afro_w7)

saveRDS(afro_w7, file="data/data_rds/afro_w7.rds")
```

# Process Wave 8 data

```{r}
# Fix country codes: Replace MAU with MUS (Mauritius) and MRC with MAR (Morocco)

afro_w8 %<>%
  mutate(RESPNO = case_when(
    str_starts(RESPNO, "MAU") ~ str_replace(RESPNO, "^MAU", "MUS"),
    TRUE ~ RESPNO
  ))

afro_w8 %<>%
  mutate(RESPNO = case_when(
    str_starts(RESPNO, "MOR") ~ str_replace(RESPNO, "^MOR", "MAR"),
    TRUE ~ RESPNO
  ))

# Add round number and create unique identifier

afro_w8%<>%
  mutate(respno=RESPNO)%>%
  mutate(round=rep(8, times = length("RESPNO")))%>%
  mutate(id_unique=paste0(RESPNO, round))

# Rename variables to standardized names across waves

afro_w8 %<>%
  rename(
    age = Q1,
    #assoc = ,
    check_citiz=Q22,
    check_parl=Q29,
    commu = Q11A,
    conditions_country = Q6A,
    conditions_eco = Q4B,
    contact_counc=Q12A,
    contact_trad=Q12D,
    corr_judge = Q42F,
    corr_level = Q43A,
    corr_local = Q42D,
    corr_parl = Q42B,
    corr_pres = Q42A,
    corr_trad = Q42H,
    disc_pol = Q9,
    educ = Q97,
    elec_fair = Q14,
    emp = Q95A,
    extent = Q36,
    feelnat = Q82B,
    food = Q7A,
    internet_news = Q55D,
    internet_use = Q92I,
    latitude = EA_GPS_LA,
    listen_counc=Q38B,
    listen_trad=Q38C,
    longitude = EA_GPS_LO,
    military_reject = Q20B,
    money = Q7E,
    one_man_reject = Q20C,
    one_party_reject = Q20A,
    paper_news = Q55C,
    perf_corr = Q50J,
    perf_crime = Q50F,
    perf_eco = Q50A,
    perf_educ = Q50H,
    perf_elect = Q50L,
    perf_health = Q50G,
    perf_ineq = Q50E,
    perf_job = Q50C,
    perf_local = Q51C,
    perf_parl = Q51B,
    perf_pres = Q51A,
    perf_price = Q50D,
    perf_road = Q50K,
    perf_trad = Q51D,
    perf_water = Q50I,
    pref_demo = Q21,
    prob = Q11B,
    protest = Q11C,
    radio_news = Q55A,
    rural = URBRUR,
    satis_demo = Q37,
    security = Q8A,
    sex = Q101,
    sm_news = Q55E,
    tele_news = Q55B,
    tele_pos = Q92B,
    trust_elec = Q41C,
    trust_gen = Q83,
    trust_govpart = Q41E,
    trust_local = Q41D,
    trust_oppart = Q41F,
    trust_parl = Q41B,
    trust_pres = Q41A,
    trust_trad = Q41K,
    unfair_eth = Q84C,
    vote = Q13)

glimpse(afro_w8)

saveRDS(afro_w8, file="data/data_rds/afro_w8.rds")
```

# Define variables to keep in final dataset

```{r}
variables_to_keep <- c(
  "age", "assoc", "check_citiz", "check_parl", "commu", "conditions_country", 
  "conditions_eco", "contact_counc", "contact_trad", "corr_judge", 
  "corr_level", "corr_local", "corr_parl", "corr_pres", "corr_trad", 
  "disc_pol", "educ", "elec_fair", "emp", "extent", "feelnat", "food", 
  "internet_news", "internet_use", "latitude", "listen_counc", "listen_trad", 
  "longitude", "military_reject", "money", "one_man_reject", "one_party_reject", 
  "paper_news", "perf_corr", "perf_crime", "perf_eco", "perf_educ", "perf_elect", 
  "perf_health", "perf_ineq", "perf_job", "perf_local", "perf_parl", "perf_pres", 
  "perf_price", "perf_road", "perf_trad", "perf_water", "pref_demo", "prob", 
  "protest", "radio_news", "rural", "satis_demo", "security", "sex", "sm_news", 
  "tele_news", "tele_pos", "trust_elec", "trust_gen", "trust_govpart", "trust_local", 
  "trust_oppart", "trust_parl", "trust_pres", "trust_trad", "unfair_eth", "vote", "id_unique", "round", "respno"
)
```

# Filter all datasets to keep only selected variables

```{r}
datasets <- list(afro_w5, afro_w6, afro_w7, afro_w8)

filtered_dataset_names <- c("afro_w5_filt", "afro_w6_filt", "afro_w7_filt", "afro_w8_filt")

for (i in 1:length(datasets)) {
  filtered_dataset <- datasets[[i]] %>%
    select(any_of(variables_to_keep))

  assign(filtered_dataset_names[i], filtered_dataset)
}

rm(afro_w5, afro_w6, afro_w7, afro_w8, datasets, filtered_dataset, filtered_dataset_names, i, variables_to_keep)
```

# Stack all waves together and remove duplicates

```{r}
afro_stack <- bind_rows(afro_w5_filt, afro_w6_filt, afro_w7_filt, afro_w8_filt)

afro_stack%<>%
  filter(!duplicated(id_unique))%>%
  filter(!is.na(latitude), !is.na(longitude))

rm(afro_w5_filt, afro_w6_filt, afro_w7_filt, afro_w8_filt)
```

# Add ISO country codes and standardize them

```{r}
afro_stack%<>%
  mutate(iso = substr(id_unique, 1, 3))

# Convert non-standard country codes to ISO format

afro_stack %<>%
  mutate(iso=case_when(
    iso=="BFO"~"BFA",
    iso=="BOT"~"BWA",
    iso=="CAM"~"CMR",
    iso=="CDI"~"CIV",
    iso=="CVE"~"CPV",
    iso=="LES"~"LSO",
    iso=="LIB"~"LBR",
    iso=="MAD"~"MDG",
    iso=="SAF"~"ZAF",
    iso=="SRL"~"SLE",
    iso=="ZAM"~"ZMB",
    iso=="ZIM"~"ZWE",
    iso=="GAM"~"GMB",
    iso=="ANG"~"AGO",
    iso=="ALG"~"DZA",
    iso=="GUI"~"GIN",
    iso=="MLW"~"MWI",
    iso=="NGR"~"NER",
    iso=="NIG"~"NGA",
    iso=="SUD"~"SDN",
    iso=="TAN"~"TZA",
    iso=="TOG"~"TGO",
    TRUE ~ iso))

# Create country mapping data frame

country_mapping <- data.frame(
  ccode = 1:39,
  cname = c("algeria", "angola", "benin", "botswana", "burkina_faso", "burundi", 
            "cabo_verde", "cameroon", "egypt", "eswatini", "ethiopia", "gabon", 
            "gambia", "ghana", "guinea", "ivory_coast", "kenya", "lesotho", 
            "liberia", "madagascar", "malawi", "mali", "mauritius", "morocco", 
            "mozambique", "namibia", "niger", "nigeria", "sao_tome", "senegal", 
            "sierra_leone", "south_africa", "sudan", "tanzania", "togo", 
            "tunisia", "uganda", "zambia", "zimbabwe")
)

iso <- c(
  "algeria" = "DZA", "angola" = "AGO", "benin" = "BEN", "botswana" = "BWA",
  "burkina_faso" = "BFA", "burundi" = "BDI", "cabo_verde" = "CPV", "cameroon" = "CMR",
  "egypt" = "EGY", "eswatini" = "SWZ", "ethiopia" = "ETH", "gabon" = "GAB",
  "gambia" = "GMB", "ghana" = "GHA", "guinea" = "GIN", "ivory_coast" = "CIV",
  "kenya" = "KEN", "lesotho" = "LSO", "liberia" = "LBR", "madagascar" = "MDG",
  "malawi" = "MWI", "mali" = "MLI", "mauritius" = "MUS", "morocco" = "MAR",
  "mozambique" = "MOZ", "namibia" = "NAM", "niger" = "NER", "nigeria" = "NGA",
  "sao_tome" = "STP", "senegal" = "SEN", "sierra_leone" = "SLE",
  "south_africa" = "ZAF", "sudan" = "SDN", "tanzania" = "TZA", "togo" = "TGO",
  "tunisia" = "TUN", "uganda" = "UGA", "zambia" = "ZMB", "zimbabwe" = "ZWE"
)

country_mapping %<>%
  mutate(iso = iso[cname])

afro_stack %<>%
  left_join(country_mapping, by = "iso")

rm(country_mapping, iso)
```

# Remove geographic outliers using IQR method

```{r}
rm_out <- function(data, variable = "latitude") {
  init_n <- nrow(data)

  Q1 <- quantile(data[[variable]], 0.25, na.rm = TRUE)
  Q3 <- quantile(data[[variable]], 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR

  data_iqr <- data %>%
    filter(!!sym(variable) >= lower_bound & !!sym(variable) <= upper_bound)

  n_rm <- init_n - nrow(data_iqr)
  pct_rm <- round(100 * n_rm / init_n, 2)

  cat("Initial observations:", init_n, "\n",
      "Observations removed:", n_rm, "\n",
      "Percentage removed:", pct_rm, "%\n",
      "Final observations:", nrow(data_iqr), "\n")

  return(data_iqr)
}

afro_stack %<>% rm_out("latitude")

saveRDS(afro_stack, file="data/data_rds/afro_stack.rds")

rm(rm_out)
```

# Create summary statistics table

```{r}
vars_sum <- c(
  "age", "assoc", "check_citiz", "check_parl", "commu", "conditions_country", 
  "conditions_eco", "contact_counc", "contact_trad", "corr_judge", 
  "corr_level", "corr_local", "corr_parl", "corr_pres", "corr_trad", 
  "disc_pol", "educ", "elec_fair", "emp", "extent", "feelnat", "food", 
  "internet_news", "internet_use", "latitude", "listen_counc", "listen_trad", 
  "longitude", "military_reject", "money", "one_man_reject", "one_party_reject", 
  "paper_news", "perf_corr", "perf_crime", "perf_eco", "perf_educ", "perf_elect", 
  "perf_health", "perf_ineq", "perf_job", "perf_local", "perf_parl", "perf_pres", 
  "perf_price", "perf_road", "perf_trad", "perf_water", "pref_demo", "prob", 
  "protest", "radio_news", "rural", "satis_demo", "security", "sex", "sm_news", 
  "tele_news", "tele_pos", "trust_elec", "trust_gen", "trust_govpart", "trust_local", 
  "trust_oppart", "trust_parl", "trust_pres", "trust_trad", "unfair_eth", "vote"
)

sum_tab <- function(data, variables_list, rounds_col = "round") {

  calc_stats <- function(var_name) {
    if (!var_name %in% names(data)) {
      return(NULL)
    }

    if (is.numeric(data[[var_name]])) {
      data %>%
        group_by(!!sym(rounds_col)) %>%
        summarise(
          Mean = mean(!!sym(var_name), na.rm = TRUE),
          Median = median(!!sym(var_name), na.rm = TRUE),
          SD = sd(!!sym(var_name), na.rm = TRUE),
          Min = min(!!sym(var_name), na.rm = TRUE),
          Max = max(!!sym(var_name), na.rm = TRUE),
          N = sum(!is.na(!!sym(var_name))),
          .groups = 'drop'
        ) %>%
        mutate(Variable = var_name)
    } else {
      data %>%
        group_by(!!sym(rounds_col)) %>%
        summarise(
          Mean = n_distinct(!!sym(var_name), na.rm = TRUE),
          Median = NA,
          SD = NA,
          Min = NA,
          Max = NA,
          N = sum(!is.na(!!sym(var_name))),
          .groups = 'drop'
        ) %>%
        mutate(Variable = var_name)
    }
  }

  all_stats <- map_df(variables_list, calc_stats)

  final_table <- all_stats %>%
    pivot_wider(
      id_cols = Variable,
      names_from = !!sym(rounds_col),
      values_from = c(Mean, Median, SD, Min, Max, N),
      names_sep = "_"
    )

  latex_table <- final_table %>%
    kbl(format = "latex",
        booktabs = TRUE,
        caption = "Summary Statistics by Round",
        digits = 2,
        align = c("l", rep("c", ncol(final_table) - 1))) %>%
    kable_styling(
      latex_options = c("striped", "hold_position", "scale_down"),
      font_size = 9
    ) %>%
    column_spec(1, bold = TRUE) %>%
    footnote(
      general = paste("Number of variables:", length(variables_list)),
      footnote_as_chunk = TRUE
    )

  return(latex_table)
}

summary_table <- sum_tab(
   data = afro_stack,
   variables_list = vars_sum,
   rounds_col = "round"
 )

print(summary_table)

writeLines(summary_table, "variables_summary.tex")

rm(vars_sum, sum_tab, summary_table)
```