### 2025-02-24 script/lightning_cell.Rmd

```{r}
setwd("~/Working paper/Trust/Script")
getwd()
```

```{r}
library(haven)
library(dplyr)
select <- dplyr::select
library(magrittr)
library(ncdf4)
library(sf)
library(raster)
library(exactextractr)
library(rgdal)
library(ggplot2)
library(terra)
```

```{r}
ncpath <- "data/data_Raw/"
ncname <- "VHRFC"  
ncfname <- paste(ncpath, ncname, ".nc", sep="")
ncin <- nc_open(ncfname)
print(ncin)

FRD <- ncvar_get(ncin, "VHRFC_LIS_FRD")
dim(FRD)

lon <- ncvar_get(ncin, "Longitude")
lat <- ncvar_get(ncin, "Latitude")

r_FRD <- raster(t(FRD), 
                xmn=min(lon), xmx=max(lon), 
                ymn=min(lat), ymx=max(lat),
                crs=CRS("+proj=longlat +datum=WGS84"))

r_FRD <- flip(r_FRD, direction = "y")

nc_close(ncin)
```

```{r}
countries <- c("BEN", "BFA", "BWA", "CIV", "CMR", "GIN", "LBR", "MLI", "NAM", 
               "NER", "SLE", "GHA", "KEN", "MOZ", "MWI", "NGA", "TZA", "UGA", 
               "ZMB", "ZWE")

accumulated_results <- data.frame()

for (country in countries) {
shapefile_path <- paste0("C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/gadm41_",country,"_0.shp")

country_shp <- readOGR(shapefile_path)

if (!identical(crs(r_FRD), crs(country_shp))) {
  country_shp <- spTransform(country_shp, crs(r_FRD))
}

cropped <- crop(r_FRD, extent(country_shp))
masked <- mask(cropped, country_shp)

data_file <- paste0("C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/",country,"_pd_2011_1km_UNadj_ASCII_XYZ.csv")

pop <- raster(data_file)

print(pop)

pop_wgs84 <- projectRaster(pop, crs = CRS("+proj=longlat +datum=WGS84"))

pop_resampled <- resample(pop_wgs84, r_FRD, method="bilinear")

pop_wgs84_cropped <- crop(pop_resampled, extent(country_shp))
pop_wgs84_masked <- mask(pop_wgs84_cropped, country_shp)

plot(pop_wgs84_masked)

pop_resampled <- resample(pop_wgs84_masked, masked, method="bilinear")

if (!all(res(masked) == res(pop_wgs84_masked))) {
  pop_wgs84_resampled <- resample(pop_wgs84_masked, masked, method = "bilinear")
} else {
  pop_wgs84_resampled <- pop_wgs84_masked
}

if (!compareRaster(masked, pop_wgs84_resampled)) {
  stop("Les rasters n'ont pas la même résolution et/ou la même étendue après ajustement.")
}

Lightning_risk <- masked / (1 + pop_resampled)

data_trust <- readRDS("Data/Data_RDS/data_merge2.rds")

data_trust%<>%
  filter(ISO==country)

zones_sf <- st_as_sf(data_trust, coords = c("longitude", "latitude"), crs = 4326)
eclairs_raster <- rast(Lightning_risk)

eclairs_values <- extract(eclairs_raster, zones_sf)

zones_sf$eclairs_count <- eclairs_values

pop_density_values <- extract(pop_resampled, zones_sf)

zones_sf$population_density <- pop_density_values

final_df <- zones_sf

light_cell_vector <- unlist(final_df$light_cell)

light_cell_vector <- final_df$eclairs_count[[2]]

final_df$light_cell <- light_cell_vector

final_df%<>%
  select(-eclairs_count)

final_df%<>%
  select(geometry, light_cell, population_density)

accumulated_results <- rbind(accumulated_results, final_df)
}

accumulated_results2 <- accumulated_results

accumulated_results2 %<>% 
  mutate(longitude = st_coordinates(geometry)[,1],
         latitude = st_coordinates(geometry)[,2])

accumulated_results2 %<>%
  st_drop_geometry()

accumulated_results2%<>%
  distinct(longitude, latitude, .keep_all = TRUE)

data_trust <- readRDS("Data/Data_RDS/data_merge2.rds")

data_trust2 <- data_trust

data_trust2 %<>% left_join(accumulated_results2, by = c("latitude", "longitude"))

data_trust2%<>%
  rename(pop_cell=population_density)

glimpse(data_trust2)

saveRDS(data_trust2, file="Data/Data_RDS/final_df2.rds")
write_dta(data_trust2, "Data/Data_dta/final_df2.dta")

```

```{r}
data_trust <- readRDS("Data/Data_RDS/data_merge2.rds")
```

```{r}
zones_sf <- st_as_sf(data_trust, coords = c("longitude", "latitude"), crs = 4326)
eclairs_raster <- rast(r_FRD)

#zones_buffer <- st_buffer(zones_sf, dist = 10000)

eclairs_values <- extract(eclairs_raster, zones_sf)

#eclairs_values <- extract(eclairs_raster, vect(zones_buffer), fun = mean)

zones_sf$eclairs_count <- eclairs_values

final_df <- st_drop_geometry(zones_sf)

light_cell_vector <- unlist(final_df$light_cell)

light_cell_vector <- final_df$eclairs_count[[2]]

final_df$light_cell <- light_cell_vector

final_df%<>%
  select(-eclairs_count)
glimpse(final_df)

saveRDS(final_df, file="Data/Data_RDS/final_df.rds")
write_dta(final_df, "Data/Data_dta/final_df.dta")

```