### 2025-02-24 script/nighttime_adm2.Rmd

```{r}
setwd("~/Working paper/Trust/Script")
getwd()
```

```{r}
library(raster)
library(rgdal)
library(dplyr)
library(magrittr)
library(sf)
library(exactextractr)
library(ggplot2)
library(haven)
```

#Data

##2019

```{r}
tiff_file <- "C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/nighttime_recode/Harmonized_DN_NTL_2019.tif"

night_wgs84 <- raster(tiff_file)

#print(night)

#night_wgs84 <- projectRaster(night, crs = CRS("+proj=longlat +datum=WGS84"))

values(night_wgs84)[values(night_wgs84) < 0] <- 0

countries_adm2 <- c("BEN", "BFA", "BWA", "CIV", "CMR", "GIN", "LBR", "MLI", "NAM", "NER", "SLE", "SEN", "TGO","GHA","KEN","MOZ", "MWI","NGA","TZA","UGA","ZMB", "ZWE")

accumulated_results <- data.frame()

for (country in countries_adm2) {

  # Lecture du fichier shapefile
  shapefile <- paste0("C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/gadm41_", country,"_2.shp")
  regions <- st_read(shapefile)
  regions <- st_make_valid(regions)

shapefile_path <- paste0("C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/gadm41_",country,"_0.shp")

country_shp <- readOGR(shapefile_path)

if (!identical(crs(night_wgs84), crs(country_shp))) {
  country_shp <- spTransform(country_shp, crs(night_wgs84))
}

# Recadrer et masquer le raster night_wgs84_recode selon les frontières du country
night_wgs84_cropped <- crop(night_wgs84, extent(country_shp))
night_wgs84_masked <- mask(night_wgs84_cropped, country_shp)

mean_night <- exact_extract(night_wgs84_masked, regions, 'mean')

regions$mean_night <- mean_night

regions$mean_night_log <- log(mean_night+1)

regions %<>%
  as.data.frame()

regions %<>%
  rename(ISO=GID_0)%>%
  select(ISO, NAME_2, mean_night, mean_night_log)%>%
  mutate(round=8)

accumulated_results <- rbind(accumulated_results, regions)

}

accumulated_results%<>%
  rename(mean_night_2=mean_night,
         mean_night_log_2=mean_night_log)

glimpse(accumulated_results)

data_trust <- readRDS("Data/Data_RDS/final_df5.rds")

data_trust %<>% left_join(accumulated_results, by=c("ISO", "NAME_2", "round"))

data_trust%<>%
  distinct(id_unique, .keep_all = TRUE)

glimpse(data_trust)
```

##2016

```{r}
tiff_file <- "C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/nighttime_recode/Harmonized_DN_NTL_2016.tif"

night_wgs84 <- raster(tiff_file)

#print(night)

#night_wgs84 <- projectRaster(night, crs = CRS("+proj=longlat +datum=WGS84"))

values(night_wgs84)[values(night_wgs84) < 0] <- 0

countries_adm2 <- c("BEN", "BFA", "BWA", "CIV", "CMR", "GIN", "LBR", "MLI", "NAM", "NER", "SLE", "SEN", "TGO", "GHA","KEN","MOZ", "MWI","NGA","TZA","UGA","ZMB", "ZWE")

accumulated_results <- data.frame()

for (country in countries_adm2) {

  # Lecture du fichier shapefile
  shapefile <- paste0("C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/gadm41_", country,"_2.shp")
  regions <- st_read(shapefile)
  regions <- st_make_valid(regions)

shapefile_path <- paste0("C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/gadm41_",country,"_0.shp")

country_shp <- readOGR(shapefile_path)

if (!identical(crs(night_wgs84), crs(country_shp))) {
  country_shp <- spTransform(country_shp, crs(night_wgs84))
}

# Recadrer et masquer le raster night_wgs84_recode selon les frontières du country
night_wgs84_cropped <- crop(night_wgs84, extent(country_shp))
night_wgs84_masked <- mask(night_wgs84_cropped, country_shp)

mean_night <- exact_extract(night_wgs84_masked, regions, 'mean')

regions$mean_night <- mean_night

regions$mean_night_log <- log(mean_night+1)

regions %<>%
  as.data.frame()

regions %<>%
  rename(ISO=GID_0)%>%
  select(ISO, NAME_2, mean_night, mean_night_log)%>%
  mutate(round=7)

accumulated_results <- rbind(accumulated_results, regions)

}

accumulated_results%<>%
  rename(mean_night_2=mean_night,
         mean_night_log_2=mean_night_log)

glimpse(accumulated_results)

data_trust %<>% left_join(accumulated_results, by=c("ISO", "NAME_2", "round"))

data_trust%<>%
  distinct(id_unique, .keep_all = TRUE)

data_trust %<>%
  mutate(mean_night_2 = coalesce(mean_night_2.x, mean_night_2.y))%>%
  mutate(mean_night_log_2 = coalesce(mean_night_log_2.x, mean_night_log_2.y))%>%
  select(-c(mean_night_log_2.x, mean_night_log_2.y, mean_night_2.x, mean_night_2.y))

glimpse(data_trust)

```

##2014
```{r}
tiff_file <- "C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/nighttime_recode/Harmonized_DN_NTL_2014.tif"

night_wgs84 <- raster(tiff_file)

#print(night)

#night_wgs84 <- projectRaster(night, crs = CRS("+proj=longlat +datum=WGS84"))

values(night_wgs84)[values(night_wgs84) < 0] <- 0

countries_adm2 <- c("BEN", "BFA", "BWA", "CIV", "CMR", "GIN", "LBR", "MLI", "NAM", "NER", "SLE", "SEN", "TGO", "GHA","KEN","MOZ", "MWI","NGA","TZA","UGA","ZMB", "ZWE")

accumulated_results <- data.frame()

for (country in countries_adm2) {

  # Lecture du fichier shapefile
  shapefile <- paste0("C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/gadm41_", country,"_2.shp")
  regions <- st_read(shapefile)
  regions <- st_make_valid(regions)

shapefile_path <- paste0("C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/gadm41_",country,"_0.shp")

country_shp <- readOGR(shapefile_path)

if (!identical(crs(night_wgs84), crs(country_shp))) {
  country_shp <- spTransform(country_shp, crs(night_wgs84))
}

# Recadrer et masquer le raster night_wgs84_recode selon les frontières du country
night_wgs84_cropped <- crop(night_wgs84, extent(country_shp))
night_wgs84_masked <- mask(night_wgs84_cropped, country_shp)

mean_night <- exact_extract(night_wgs84_masked, regions, 'mean')

regions$mean_night <- mean_night

regions$mean_night_log <- log(mean_night+1)

regions %<>%
  as.data.frame()

regions %<>%
  rename(ISO=GID_0)%>%
  select(ISO, NAME_2, mean_night, mean_night_log)%>%
  mutate(round=6)

accumulated_results <- rbind(accumulated_results, regions)

}

accumulated_results%<>%
  rename(mean_night_2=mean_night,
         mean_night_log_2=mean_night_log)

glimpse(accumulated_results)

data_trust %<>% left_join(accumulated_results, by=c("ISO", "NAME_2", "round"))

data_trust%<>%
  distinct(id_unique, .keep_all = TRUE)

data_trust %<>%
  mutate(mean_night_2 = coalesce(mean_night_2.x, mean_night_2.y))%>%
  mutate(mean_night_log_2 = coalesce(mean_night_log_2.x, mean_night_log_2.y))%>%
  select(-c(mean_night_log_2.x, mean_night_log_2.y, mean_night_2.x, mean_night_2.y))

glimpse(data_trust)
```

##2011

```{r}
tiff_file <- "C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/nighttime_recode/Harmonized_DN_NTL_2011.tif"

night_wgs84 <- raster(tiff_file)

#print(night)

#night_wgs84 <- projectRaster(night, crs = CRS("+proj=longlat +datum=WGS84"))

values(night_wgs84)[values(night_wgs84) < 0] <- 0

countries_adm2 <- c("BEN", "BFA", "BWA", "CIV", "CMR", "GIN", "LBR", "MLI", "NAM", "NER", "SLE", "SEN", "TGO", "GHA","KEN","MOZ", "MWI","NGA","TZA","UGA","ZMB", "ZWE")

accumulated_results <- data.frame()

for (country in countries_adm2) {

  # Lecture du fichier shapefile
  shapefile <- paste0("C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/gadm41_", country,"_2.shp")
  regions <- st_read(shapefile)
  regions <- st_make_valid(regions)

shapefile_path <- paste0("C:/Users/Redha CHABA/Documents/Working paper/Trust/Script/data/data_Raw/gadm41_",country,"_0.shp")

country_shp <- readOGR(shapefile_path)

if (!identical(crs(night_wgs84), crs(country_shp))) {
  country_shp <- spTransform(country_shp, crs(night_wgs84))
}

# Recadrer et masquer le raster night_wgs84_recode selon les frontières du country
night_wgs84_cropped <- crop(night_wgs84, extent(country_shp))
night_wgs84_masked <- mask(night_wgs84_cropped, country_shp)

mean_night <- exact_extract(night_wgs84_masked, regions, 'mean')

regions$mean_night <- mean_night

regions$mean_night_log <- log(mean_night+1)

regions %<>%
  as.data.frame()

regions %<>%
  rename(ISO=GID_0)%>%
  select(ISO, NAME_2, mean_night, mean_night_log)%>%
  mutate(round=5)

accumulated_results <- rbind(accumulated_results, regions)

}

accumulated_results%<>%
  rename(mean_night_2=mean_night,
         mean_night_log_2=mean_night_log)

glimpse(accumulated_results)

data_trust %<>% left_join(accumulated_results, by=c("ISO", "NAME_2", "round"))

data_trust%<>%
  distinct(id_unique, .keep_all = TRUE)

data_trust %<>%
  mutate(mean_night_2 = coalesce(mean_night_2.x, mean_night_2.y))%>%
  mutate(mean_night_log_2 = coalesce(mean_night_log_2.x, mean_night_log_2.y))%>%
  select(-c(mean_night_log_2.x, mean_night_log_2.y, mean_night_2.x, mean_night_2.y))

glimpse(data_trust)

```

```{r}
saveRDS(data_trust, file="Data/Data_RDS/final_df5.rds")
write_dta(data_trust, "Data/Data_dta/final_df5.dta")
```