### 2025-02-24 tables/v_1/tables.tex

%l !Rnw weave = knitr

\documentclass[11pt]{article}

\usepackage[dvipsnames]{xcolor}

\usepackage[]{graphicx} % omit 'demo' option in real doc.
\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\usepackage{epstopdf}
\usepackage{float}
\epstopdfDeclareGraphicsRule{.tif}{png}{.png}{convert #1 \OutputFile}
\AppendGraphicsExtensions{.tif}

\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

\let\oldFootnote\footnote
\newcommand\nextToken\relax

\renewcommand\footnote[1]{%
    \oldFootnote{#1}\futurelet\nextToken\isFootnote}

\newcommand\isFootnote{%
    \ifx\footnote\nextToken\textsuperscript{,}\fi}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{crimson}

\usepackage{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\usepackage{url}
\usepackage{dcolumn}
\usepackage{ctable}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{rotating}
\usepackage{subcaption}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{soul}
\usepackage{float}
\usepackage{bbding} % checkmark symbol
\usepackage{comment} % checkmark symbol

\renewcommand{\textfraction}{0.05}
\renewcommand{\topfraction}{0.8}
\renewcommand{\bottomfraction}{0.8}
\renewcommand{\floatpagefraction}{0.75}

%Bibliography
\usepackage{natbib}

\newtheorem{hyp}{Hypothesis}
\makeatletter
\newcounter{subhyp}
\let\savedc@hyp\c@hyp
\newenvironment{subhyp}
 {%
  \setcounter{subhyp}{0}%
  \stepcounter{hyp}%
  \edef\saved@hyp{\thehyp}% Save the current value of hyp
  \let\c@hyp\c@subhyp     % Now hyp is subhyp
  \renewcommand{\thehyp}{\saved@hyp\alph{hyp}}%
 }
 {}
\newcommand{\normhyp}{%
  \let\c@hyp\savedc@hyp % revert to the old one
  \renewcommand\thehyp{\arabic{hyp}}%
}
\makeatother

\setcounter{MaxMatrixCols}{10}

\theoremstyle{plain}
\newtheorem{prop}{\protect\propositionname}
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}
\makeatother
\providecommand{\examplename}{Example}
\providecommand{\propositionname}{Proposition}
\providecommand{\theoremname}{Theorem}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\bb}{\begin{block}}
\newcommand{\eb}{\end{block}}
\newcommand{\bmath}{\begin{eqnarray}}
\newcommand{\emath}{\end{eqnarray}}
\newcommand{\bmathnn}{\begin{eqnarray*}}
\newcommand{\emathnn}{\end{eqnarray*}}
\newtheorem{theorem}{{Theorem}}%[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{prediction}{Prediction}
\newtheorem{open_question}{Open question}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{claim}[theorem]{{Claim}}
\newtheorem{example}[theorem]{{Example}}
\newtheorem{corollary}[theorem]{{Corollary}}
\newtheorem{remark}[theorem]{{Remark}}
\newtheorem{assumptions}[theorem]{{Assumptions}}
\newtheorem*{thm1star}{{Theorem $\mathbf{1^*}$}}
\newtheorem*{thm2star}{{Theorem $\mathbf{2^*}$}}
\newtheorem*{theorem*}{Theorem}
\newtheorem*{lemma*}{Lemma}

%For nice tables
\usepackage{booktabs}
\usepackage{array}
\usepackage{threeparttable}
\usepackage{tabulary}
\newcolumntype{M}[1]{{\centering\arraybackslash}m{#1}}

%Expectation operator
\newcommand{\Expect}{{\rm I\kern-.3em E}}

%Significance commands
\newcommand*{\SuperScriptSameStyle}[1]{%
  \ensuremath{%
    \mathchoice
      {{}^{\displaystyle #1}}%
      {{}^{\textstyle #1}}%
      {{}^{\scriptstyle #1}}%
      {{}^{\scriptscriptstyle #1}}%
  }%
}

\newcommand*{\oneS}{\SuperScriptSameStyle{*}}
\newcommand*{\twoS}{\SuperScriptSameStyle{**}}
\newcommand*{\threeS}{\SuperScriptSameStyle{*{*}*}}
\newcommand{\vh}[1]{\textcolor{red}{(VH: #1)}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

\usepackage{titlesec}
\usepackage[dvipsnames]{xcolor}

\titleformat{\section}
{\color{WildStrawberry}\normalfont\LARGE\bfseries}
{\color{WildStrawberry}\thesection}{1em}{}

\titleformat{\subsection}
{\color{ForestGreen}\normalfont\Large\bfseries}
{\color{ForestGreen}\thesubsection}{1em}{}

\titleformat{\subsubsection}
{\color{NavyBlue}\normalfont\large\bfseries}
{\color{NavyBlue}\thesubsubsection}{1em}{}

\titleformat{\paragraph}[runin]
{\normalfont\normalsize\bfseries}
{}{0pt}{}

\titleformat{\subparagraph}[runin]
{\normalfont\normalsize\bfseries}
{}{0pt}{}

% Adjust spacing after headings
\titlespacing*{\section}{0pt}{\baselineskip}{\baselineskip}
\titlespacing*{\subsection}{0pt}{\baselineskip}{\baselineskip}
\titlespacing*{\subsubsection}{0pt}{\baselineskip}{\baselineskip}
\titlespacing*{\paragraph}{0pt}{\baselineskip}{\baselineskip}
\titlespacing*{\subparagraph}{0pt}{\baselineskip}{\baselineskip}

% Adjust numbering depth
\setcounter{secnumdepth}{3} % Number subsubsections

\usepackage{hyperref}
\usepackage{chngcntr}
\usepackage{lscape}
\usepackage{rotate}
\usepackage{pdflscape}

\usepackage{booktabs}      % For professional table lines
\usepackage{multirow}      % For merged cells across rows
\usepackage{makecell}      % For better cell formatting
\usepackage{array}         % For advanced array and table formatting
\usepackage{colortbl}      % If you want to use colored cells

\begin{document}

\title{The Death of Distance: Mobile Internet
and Political Trust in Africa \\  \vspace{0.75cm} \underline{Tables}}

\vspace{1cm}

\date{ \today}

\maketitle

\begin{table}[H]
    \sffamily
    \caption{{Effect of distance from the capital on political trust}}
    \centering
    %\footnotesize
    \resizebox{8cm}{!}{
    \begin{tabular}{@{\extracolsep{5pt}} l c c c c}
    \\
    \toprule
    \toprule
    &  \multicolumn{4}{c}{{OLS}}  \\
    \cmidrule(r){2-5}
    &  \multicolumn{4}{c}{{Political trust}}  \\
    \cmidrule(r){2-5}
    & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}}\\
        \cmidrule(r){2-3}
        \cmidrule(r){4-5}
        & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}}\\
     \midrule

     Distance from the capital&       0.300***&       0.298***&       0.618***&       0.341** \\
     \smallskip
     &      (0.04)   &      (0.03)   &      (0.15)   &      (0.16)   \\
     \midrule
     \smallskip
    Individual \& regional controls  & Yes & Yes & Yes & Yes  \\
    \smallskip
    Country controls & Yes& No& Yes& No\\
    \smallskip
    Round FE & Yes & No& Yes & No\\
    \smallskip
    Country X Round FE       & No & Yes& No & Yes\\
    \smallskip
    Ethnic homeland FE & No & No & Yes& Yes\\
    \smallskip
    Observations           &      107,117   &      111,570   &       10,790   &       11,924   \\
    Adjusted-R$^2$         &       0.105   &       0.156   &       0.140   &       0.172   \\
                          \bottomrule
    \multicolumn{5}{p{12.5cm}}{\footnotesize \emph{Notes}: The border sample includes individuals residing within a 40-kilometer buffer around a country border that overlaps with a historical ethnic homeland, as defined by Murdock (1959). Robust standard errors clustered at the region x round level  for the base sample and ethnic homeland x region x round level for the border sample are in parentheses. The set of individual controls
    includes measures of: normalized distance from the largest non-capital city, age, age squared, sex,
    education, employment status, rural/urban situation, personal economic conditions perception, interest in politics, TV news consumption, radio news consumption, newspaper consumption. (3) and (4) also include a measure of ethnic discrimination. The set of regional controls includes measures of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
    \end{tabular}
    }
    \end{table}

    \begin{table}[H]
        \sffamily
        \caption{Effect of internet coverage on political trust by distance}
        \centering
        %\footnotesize
        \resizebox{12cm}{!}{
        \begin{tabular}{@{\extracolsep{5pt}} l c c c c c c}
        \\
        \toprule
        \toprule
        & \multicolumn{5}{c}{{Base sample}}\\
        \cmidrule(r){2-6}
        & \multicolumn{1}{c}{{OLS}} & \multicolumn{3}{c}{{First Stage}} & \multicolumn{1}{c}{{2SLS}}\\
            \cmidrule(r){2-2}
            \cmidrule(r){3-5}
            \cmidrule(r){6-6}
            & \multicolumn{1}{c}{{Political trust}} & \multicolumn{1}{c}{{Internet coverage}} & \multicolumn{1}{c}{{}} &  \multicolumn{1}{c}{{Distance $\times$ Internet coverage}} & \multicolumn{1}{c}{{Political trust}}\\
            \cmidrule(r){2-2}
            \cmidrule(r){3-3}
            \cmidrule(r){5-5}    
            \cmidrule(r){6-6}    
            & \multicolumn{1}{c}{{(1)}} & \multicolumn{1}{c}{{(2)}} & \multicolumn{1}{c}{{}} &  \multicolumn{1}{c}{{(3)}} & \multicolumn{1}{c}{{(4)}}\\

            \midrule

    Distance from the capital&       0.453***&&&&       1.523***\\
    \smallskip
    &      (0.05)   &&&&      (0.56)   \\
    Internet coverage&      -0.015   &&&&       1.770** \\
\smallskip
&      (0.06)   &&&&      (0.69)   \\
Distance from the capital $\times$ Internet coverage&     -0.466***&&&&       -2.950**         \\
\smallskip
            &      (0.10)   &&&&        (1.45)           \\

   Lightning strikes && -0.002*** && -0.000  \\
    \smallskip&& (0.00) && (0.00) \\
    Distance from the capital city $\times$ Lightning strikes&&-0.000 && -0.001***\\
    \medskip&& (0.00) && (0.00)\\

         \midrule
        SW \emph{F} - Lightning strikes &-&-& 13.74&- &-\\
        \smallskip
        SW \emph{F} - Distance $\times$ Lightning strikes &-&-& 8.62 &-&-\\
        \smallskip
        Individual \& regional controls  & Yes & Yes &&  Yes & Yes\\
        \smallskip
        Country X Round FE       & Yes & Yes&- & Yes & Yes \\
        \smallskip
        Observations       &       111,570    &113,243 & 113,243&       111,570  \\
        Adjusted-R$^2$    &       0.158     &-&-&-&  \\
                              \bottomrule
        \multicolumn{6}{p{25cm}}{\footnotesize \emph{Notes}: % Set font size to 8pt with 10pt line spacing
        \emph{Notes}: Robust standard errors clustered at the region x round level. The set of individual controls
        includes measures of: normalized distance from the largest non-capital city, age, age squared, sex,
        education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, newspaper consumption, radio consumption. The set of regional controls includes measures of: nighttime light, population density, region area, president birthplace dummy. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
    \end{tabular}}
        \end{table}

        \begin{table}[H]
            \sffamily
            \caption{{Political accountability}}
            \centering
            %\footnotesize
            \resizebox{12cm}{!}{
            \begin{tabular}{@{\extracolsep{5pt}} l c c}
            \\
            \toprule
            \toprule
            & \multicolumn{2}{c}{2SLS}\\
            \cmidrule(r){2-3}
            &\multicolumn{1}{c}{{Vote against ruling party}}&  \multicolumn{1}{c}{{Country performance}}  \\
            \cmidrule(r){2-2}
            \cmidrule(r){3-3}
            & \multicolumn{1}{c}{{Base sample}} & \multicolumn{1}{c}{{Base sample}}\\
                \cmidrule(r){2-2}
                \cmidrule(r){3-3}
                & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}\\
             \midrule

            Distance from the capital&      -0.932***&       1.874***\\
            \smallskip
            &      (0.28)   &      (0.65)   \\
            Internet coverage&       -1.163***&       2.089***\\
            \smallskip
            &        (0.41)   &      (0.77) \\
            Distance from the capital $\times$ Internet coverage&       2.104***&      -4.200** \\
            \medskip
            &      (0.74)   &      (1.69)   \\

             \midrule
             \smallskip
            Individual \& regional controls  & Yes & Yes\\
            \smallskip
            Country X Round FE       & Yes& Yes\\
            \smallskip
            Observations      &      74,959   &      111,696  \\            \bottomrule
            \multicolumn{3}{p{18cm}}{\footnotesize \emph{Notes}: Robust standard errors clustered at the region x round level are in parentheses. The set of individual controls
            includes measures of: normalized distance from the largest non-capital city, age, age squared, sex,
            education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, newspaper consumption, radio consumption. The set of regional controls includes measures of: nighttime light, population density, region area, president birthplace dummy. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
            \end{tabular}
            }
            \end{table}

            \begin{table}[H]
                \sffamily
                \caption{{Media and institutions freedom}}
                \centering
                %\footnotesize
                \resizebox{10cm}{!}{
                \begin{tabular}{@{\extracolsep{5pt}} l c c c c}
                \\
                \toprule
                \toprule
                & \multicolumn{4}{c}{2SLS: Political trust}\\
                \cmidrule(r){2-5}
                & \multicolumn{4}{c}{{Base sample}}\\
                \cmidrule(r){2-5}
                &\multicolumn{2}{c}{{Media}}&  \multicolumn{2}{c}{{Institutions}}\\
                \cmidrule(r){2-3}
                    \cmidrule(r){4-5}
                    & \multicolumn{1}{c}{{Free}} & \multicolumn{1}{c}{{Captured}}& \multicolumn{1}{c}{{Free}} & \multicolumn{1}{c}{{Captured}}\\
                    \cmidrule(r){2-2}
                    \cmidrule(r){3-3}
                    \cmidrule(r){4-4}
                    \cmidrule(r){5-5}
                    & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}}\\
                 \midrule

              Distance from the capital&      -8.382   &       1.083** &       2.367   &       0.938***\\
              \smallskip
              &      (24.71)   &      (0.42)   &      (1.50)   &      (0.32)   \\
              Internet coverage&      -5.260   &       1.004   &       2.653*  &       0.762   \\
              \smallskip
              &      (19.11)   &      (0.74)   &      (1.49)   &      (0.69)   \\
              Distance from the capital $\times$ Internet coverage&      18.335   &      -2.149*  &      -5.132   &      -1.524**  \\
              \medskip
              &       (52.59)   &      (1.25)   &      (3.92)   &      (0.77)   \\

                 \midrule
                 \smallskip
                Individual \& regional controls  & Yes & Yes & Yes & Yes\\
                \smallskip
                Country X Round FE       & Yes& Yes & Yes& Yes\\
                \smallskip
                Observations          &       50,288   &       61,282   &       51,737   &       59,833  \\
                \bottomrule
                \multicolumn{5}{p{15cm}}{\footnotesize \emph{Notes}: Robust standard errors clustered at the region x round level are in parentheses. The set of individual controls
                includes measures of: normalized distance from the largest non-capital city, age, age squared, sex,
                education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, newspaper consumption, radio consumption. The set of regional controls includes measures of: nighttime light, population density, region area, president birthplace dummy. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
                \end{tabular}
                }
                \end{table}

                \begin{table}[H]
                    \sffamily
                    \caption{{Individual heterogeneity}}
                    \centering
                    %\footnotesize
                    \resizebox{10cm}{!}{
                    \begin{tabular}{@{\extracolsep{5pt}} l c c c c}
                    \\
                    \toprule
                    \toprule
                    & \multicolumn{4}{c}{2SLS: Political trust}\\
                    \cmidrule(r){2-5}
                    & \multicolumn{4}{c}{{Base sample}}\\
                    \cmidrule(r){2-5}
                    &\multicolumn{2}{c}{{Education}}&  \multicolumn{2}{c}{{Urban/Rural}}\\
                    \cmidrule(r){2-3}
                        \cmidrule(r){4-5}
                        & \multicolumn{1}{c}{{$<$ Secondary}} & \multicolumn{1}{c}{{$\geq$ Secondary}}& \multicolumn{1}{c}{{Urban}} & \multicolumn{1}{c}{{Rural}}\\
                        \cmidrule(r){2-2}
                        \cmidrule(r){3-3}
                        \cmidrule(r){4-4}
                        \cmidrule(r){5-5}
                        & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}}\\
                     \midrule

                       Distance from the capital&        1.467***&       1.440   &       1.074** &       2.666***\\
                       \smallskip
                       &      (0.45)   &      (1.44)   &      (0.54)   &      (0.89)  \\
                       Internet coverage&       1.613** &       1.632   &       0.525   &       4.471***\\
                       \smallskip
                       &     (0.63)   &      (1.34)   &      (0.62)   &      (1.48)    \\
                       Distance from the capital $\times$ Internet coverage&      -2.901** &      -2.688   &      -1.787   &      -5.898** \\
                       \medskip
                       &       (1.24)   &      (3.22)   &      (1.12)   &      (2.55)    \\

                     \midrule
                     \smallskip
                    Individual \& regional controls  & Yes & Yes & Yes & Yes\\
                    \smallskip
                    Country X Round FE       & Yes& Yes & Yes& Yes\\
                    \smallskip
                    Observations         &        79,394   &       32,176   &       42,509   &       69,061\\
                    \bottomrule
                    \multicolumn{5}{p{18cm}}{\footnotesize \emph{Notes}: Robust standard errors clustered at the region x round level are in parentheses. The set of individual controls
                    includes measures of: normalized distance from the largest non-capital city, age, age squared, sex,
                    education, employment status, rural/urban situation, personal economic conditions perception, ethnic discrimination, interest in politics, TV news consumption, newspaper consumption, radio consumption. The set of regional controls includes measures of: nighttime light, population density, region area, president birthplace dummy. Education, rural/urban, and age controls are omitted from columns (1-2), (3-4), and (5-8), respectively.*** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
                    \end{tabular}
                    }
                    \end{table}

\newpage

\centering{\large{\textbf{APPENDIX}}}

\begin{table}[H]
  \sffamily
  \caption{{Effect of internet coverage on internet news consumption and internet use}}
  \centering
  %\footnotesize
  \resizebox{8cm}{!}{
  \begin{tabular}{@{\extracolsep{5pt}} l c c c c}
  \\
  \toprule
  \toprule
  &  \multicolumn{4}{c}{{OLS}}  \\
  \cmidrule(r){2-5}
  & \multicolumn{2}{c}{{Internet news}} & \multicolumn{2}{c}{{Internet use}}\\
      \cmidrule(r){2-3}
      \cmidrule(r){4-5}
      & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}}\\
   \midrule

   Internet coverage &       0.900***&       1.127***&       0.948***&       1.217***\\
   \medskip&      (0.06)   &      (0.05)   &      (0.06)   &      (0.06)   \\

   \midrule
   \smallskip
  Individual \& regional controls  & Yes & Yes & Yes & Yes  \\
  \smallskip
  Country controls & Yes& No& Yes& No\\
  \smallskip
  Round FE & Yes & No& Yes & No\\
  \smallskip
  Country X Round FE       & No & Yes& No & Yes\\
  \smallskip
  Ethnic homeland FE & No & No & Yes& Yes\\
  \smallskip
  Observations          &    120,988   &      120,988   &      120,859   &      120,859    \\
  Adjusted-R$^2$         &       0.092   &       0.148   &       0.100   &       0.158 \\
                        \bottomrule
  \multicolumn{5}{p{12.5cm}}{\footnotesize \emph{Notes}: Robust standard errors clustered at the region x round level are in parentheses. The set of individual controls
  includes measures of: normalized distance from the capital normalized distance from the largest non-capital city, age, age squared, sex,
  education, employment status, rural/urban situation, personal economic conditions perception, interest in politics, TV news consumption, radio news consumption, newspaper consumption. Border samples regressions also include a measure of ethnic discrimination. The set of regional controls includes measures of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
  \end{tabular}
  }
  \end{table}

  \begin{table}[H]
    \sffamily
    \caption{{Effect of distance from the capital on trust in president and parliament}}
    \centering
    %\footnotesize
    \resizebox{12cm}{!}{
    \begin{tabular}{@{\extracolsep{5pt}} l c c c c c c c c}
    \\
    \toprule
    \toprule
    &  \multicolumn{8}{c}{{OLS}}  \\
    \cmidrule(r){2-9}
    &  \multicolumn{4}{c}{{Trust in president}}&  \multicolumn{4}{c}{{Trust in parliament}}  \\
    \cmidrule(r){2-5}
    \cmidrule(r){6-9}
    & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}} & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}}\\
        \cmidrule(r){2-3}
        \cmidrule(r){4-5}
        \cmidrule(r){6-7}
        \cmidrule(r){8-9}
        & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}} &  \multicolumn{1}{c}{{(5)}} &  \multicolumn{1}{c}{{(6)}} &  \multicolumn{1}{c}{{(7)}} &  \multicolumn{1}{c}{{(8)}}\\
     \midrule

     Distance from the capital&       0.316***&       0.292***&       0.733***&       0.259   &       0.285***&       0.305***&       0.531***&       0.470***\\
     \medskip&      (0.05)   &      (0.04)   &      (0.18)   &      (0.18)  &      (0.04)   &      (0.03)   &      (0.14)   &      (0.15) \\

\midrule
     \smallskip
    Individual \& regional controls  & Yes & Yes & Yes & Yes  & Yes & Yes & Yes & Yes \\
    \smallskip
    Country controls & Yes& No& Yes& No& Yes& No& Yes& No\\
    \smallskip
    Round FE & Yes & No& Yes & No & Yes & No& Yes & No\\
    \smallskip
    Country X Round FE       & No & Yes& No & Yes& No & Yes& No & Yes\\
    \smallskip
    Ethnic homeland FE & No & No & Yes& Yes& No & No & Yes& Yes\\
    \smallskip
    Observations           &      105,993   &      110,439   &       10,642   &       11,772  &      104,535   &      108,982   &       10,534   &       11,667    \\
    Adjusted-R$^2$         &      0.084   &       0.138   &       0.173   &       0.200   &       0.095   &       0.134   &       0.086   &       0.120 \\
                          \bottomrule
    \multicolumn{9}{p{21cm}}{\footnotesize \emph{Notes}: The border sample includes individuals residing within a 40-kilometer buffer around a country border that overlaps with a historical ethnic homeland, as defined by Murdock (1959). Robust standard errors clustered at the region x round level  for the base sample and ethnic homeland x region x round level for the border sample are in parentheses. The set of individual controls
    includes measures of: normalized distance from the largest non-capital city, age, age squared, sex,
    education, employment status, rural/urban situation, personal economic conditions perception, interest in politics, TV news consumption, radio news consumption, newspaper consumption. Border samples regressions also include a measure of ethnic discrimination. The set of regional controls includes measures of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
    \end{tabular}
    }
    \end{table}

    \begin{table}[H]
      \sffamily
      \caption{{Effect of distance from the capital on political trust - Different distance measures}}
      \centering
      %\footnotesize
      \resizebox{18cm}{!}{
      \begin{tabular}{@{\extracolsep{5pt}} l c c c c c c c c c c c c}
      \\
      \toprule
      \toprule
      &  \multicolumn{12}{c}{{OLS}}  \\
      \cmidrule(r){2-13}
      &  \multicolumn{12}{c}{{Political trust}}  \\
      \cmidrule(r){2-13}
      & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}} & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}} & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}}\\
          \cmidrule(r){2-3}
          \cmidrule(r){4-5}
          \cmidrule(r){6-7}
          \cmidrule(r){8-9}
          \cmidrule(r){10-11}
          \cmidrule(r){12-13}
          & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}} &  \multicolumn{1}{c}{{(5)}} &  \multicolumn{1}{c}{{(6)}} &  \multicolumn{1}{c}{{(7)}} &  \multicolumn{1}{c}{{(8)}} &  \multicolumn{1}{c}{{(9)}} &  \multicolumn{1}{c}{{(10)}} &  \multicolumn{1}{c}{{(11)}} &  \multicolumn{1}{c}{{(12)}}\\
       \midrule

       Distance from the capital (in 10km) &       0.004***&       0.003***&       0.009***&       0.003   \\
       \smallskip
       &      (0.00)   &      (0.00)   &      (0.00)   &      (0.00)   \\
       log(Distance from the capital)&&&&&       0.039***&       0.042***&       0.158***&       0.088** \\
       \smallskip
       &&&&&      (0.01)   &      (0.01)   &      (0.04)   &      (0.04)   \\
        Distance from the capital (mean normalization) &&&&&&&&&       0.107***&       0.101***&       0.234***&       0.107** \\
       \medskip     &&&&&&&&&      (0.02)   &      (0.01)   &      (0.05)   &      (0.05)   \\

  \midrule
       \smallskip
      Individual \& regional controls  & Yes & Yes & Yes & Yes  & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\
      \smallskip
      Country controls & Yes& No& Yes& No& Yes& No& Yes& No& Yes& No& Yes& No\\
      \smallskip
      Round FE & Yes & No& Yes & No & Yes & No& Yes & No & Yes & No& Yes & No\\
      \smallskip
      Country X Round FE       & No & Yes& No & Yes& No & Yes& No & Yes& No & Yes& No & Yes\\
      \smallskip
      Ethnic homeland FE & No & No & Yes& Yes& No & No & Yes& Yes& No & No & Yes& Yes\\
      \smallskip
      Observations            &      107,117   &      111,570   &       10,790   &       11,924   &      107,117   &      111,570   &       10,790   &       11,924 &      107,117   &      111,570   &       10,790   &       11,924   \\
      Adjusted-R$^2$         &       0.107   &       0.156   &       0.142   &       0.172 &       0.103   &       0.154   &       0.139   &       0.173  &       0.105   &       0.156   &       0.141   &       0.172\\
                            \bottomrule
      \multicolumn{13}{p{32cm}}{\footnotesize \emph{Notes}: The border sample includes individuals residing within a 40-kilometer buffer around a country border that overlaps with a historical ethnic homeland, as defined by Murdock (1959). Robust standard errors clustered at the region x round level  for the base sample and ethnic homeland x region x round level for the border sample are in parentheses. The set of individual controls
      includes measures of: normalized distance from the largest non-capital city, age, age squared, sex,
      education, employment status, rural/urban situation, personal economic conditions perception, interest in politics, TV news consumption, radio news consumption, newspaper consumption. Border samples regressions also include a measure of ethnic discrimination. The set of regional controls includes measures of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
      \end{tabular}
      }
      \end{table}

      \begin{table}[H]
        \sffamily
        \caption{{Effect of distance from the capital on political trust - Non-linearities}}
        \centering
        %\footnotesize
        \resizebox{18cm}{!}{
        \begin{tabular}{@{\extracolsep{5pt}} l c c c c c c c c c c c c}
        \\
        \toprule
        \toprule
        &  \multicolumn{12}{c}{{OLS}}  \\
        \cmidrule(r){2-13}
        &  \multicolumn{12}{c}{{Political trust}}  \\
        \cmidrule(r){2-13}
        & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}} & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}} & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}}\\
            \cmidrule(r){2-3}
            \cmidrule(r){4-5}
            \cmidrule(r){6-7}
            \cmidrule(r){8-9}
            \cmidrule(r){10-11}
            \cmidrule(r){12-13}
            & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}} &  \multicolumn{1}{c}{{(5)}} &  \multicolumn{1}{c}{{(6)}} &  \multicolumn{1}{c}{{(7)}} &  \multicolumn{1}{c}{{(8)}} &  \multicolumn{1}{c}{{(9)}} &  \multicolumn{1}{c}{{(10)}} &  \multicolumn{1}{c}{{(11)}} &  \multicolumn{1}{c}{{(12)}}\\
         \midrule
         Distance from the capital (Q2)&      -0.005   &      -0.005   &       0.132   &       0.156*  \\
         \smallskip
         &      (0.03)   &      (0.02)   &      (0.11)   &      (0.09)   \\
         Distance from the capital (Q3)&      -0.026   &      -0.021   &       0.269** &       0.193*  \\
         \smallskip&      (0.04)   &      (0.03)   &      (0.13)   &      (0.10)   \\
         Distance from the capital (Q4)&       0.102***&       0.102***&       0.371***&       0.214** \\
         \smallskip&      (0.03)   &      (0.02)   &      (0.14)   &      (0.11)   \\
         Distance from the capital (Q5)&       0.170***&       0.165***&       0.379***&       0.169   \\
         \smallskip&      (0.04)   &      (0.03)   &      (0.14)   &      (0.11)   \\
         Distance from the capital&&&&&  -0.438***&      -0.252** &       1.250***&       1.019** &      -0.981***&      -0.537** &       1.923** &       2.642***\\
         \smallskip&&&&&     (0.15)   &      (0.11)   &      (0.48)   &      (0.40)   &      (0.38)   &      (0.26)   &      (0.89)   &      (0.76)   \\
         Distance from the capital$^2$&&&&&      0.832***&       0.621***&      -0.619   &      -0.669** &       2.320** &       1.404** &      -2.227   &      -4.725***\\
         \smallskip&&&&&     (0.16)   &      (0.12)   &      (0.41)   &      (0.34)  &      (0.91)   &      (0.64)   &      (1.78)   &      (1.65)  \\
         Distance from the capital$^3$&&&&&&&&&      -1.077*  &      -0.567   &       1.043   &       2.658** \\
         \medskip
         &&&&&&&&&      (0.61)   &      (0.45)   &      (1.11)   &      (1.05)   \\

    \midrule
         \smallskip
        Individual \& regional controls  & Yes & Yes & Yes & Yes  & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\
        \smallskip
        Country controls & Yes& No& Yes& No& Yes& No& Yes& No& Yes& No& Yes& No\\
        \smallskip
        Round FE & Yes & No& Yes & No & Yes & No& Yes & No & Yes & No& Yes & No\\
        \smallskip
        Country X Round FE       & No & Yes& No & Yes& No & Yes& No & Yes& No & Yes& No & Yes\\
        \smallskip
        Ethnic homeland FE & No & No & Yes& Yes& No & No & Yes& Yes& No & No & Yes& Yes\\
        \smallskip
        Observations            &      107,117   &      111,570   &       10,790   &       11,924   & 107,117   &      111,570   &       10,790   &       11,924   &      107,117   &      111,570   &       10,790   &       11,924 \\
        Adjusted-R$^2$         &    0.106   &       0.157   &       0.138   &       0.173 & 0.107   &       0.157   &       0.141   &       0.173 &       0.108   &       0.157   &       0.141   &       0.174\\
                              \bottomrule
        \multicolumn{13}{p{29cm}}{\footnotesize \emph{Notes}: The border sample includes individuals residing within a 40-kilometer buffer around a country border that overlaps with a historical ethnic homeland, as defined by Murdock (1959). Robust standard errors clustered at the region x round level  for the base sample and ethnic homeland x region x round level for the border sample are in parentheses. The set of individual controls
        includes measures of: normalized distance from the largest non-capital city, age, age squared, sex,
        education, employment status, rural/urban situation, personal economic conditions perception, interest in politics, TV news consumption, radio news consumption, newspaper consumption. Border samples regressions also include a measure of ethnic discrimination. The set of regional controls includes measures of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
        \end{tabular}
        }
        \end{table}

        \begin{table}[H]
          \sffamily
          \caption{{Effect of distance from the capital on political trust - Different border samples}}
          \centering
          %\footnotesize
          \resizebox{8cm}{!}{
          \begin{tabular}{@{\extracolsep{5pt}} l c c c c}
          \\
          \toprule
          \toprule
          &  \multicolumn{4}{c}{{OLS}}  \\
          \cmidrule(r){2-5}
          &  \multicolumn{4}{c}{{Political trust}}  \\
          \cmidrule(r){2-5}
          & \multicolumn{2}{c}{{30-km buffer}} & \multicolumn{2}{c}{{50-km buffer}}\\
              \cmidrule(r){2-3}
              \cmidrule(r){4-5}
              & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}}\\
           \midrule

          Distance from the capital&       0.629***&       0.147   &       0.561***&       0.274*  \\
           \medskip
           &      (0.16)   &      (0.19)   &      (0.14)   &      (0.14)   \\
          \midrule
           \smallskip
          Individual \& regional controls  & Yes & Yes & Yes & Yes  \\
          \smallskip
          Country controls & Yes& No& Yes& No\\
          \smallskip
          Round FE & Yes & No& Yes & No\\
          \smallskip
          Country X Round FE       & No & Yes& No & Yes\\
          \smallskip
          Ethnic homeland FE & No & No & Yes& Yes\\
          \smallskip
          Observations           &       7,568   &        8,543   &       13,352   &       14,729    \\
          Adjusted-R$^2$         &        0.135   &       0.173   &       0.138   &       0.167   \\
                                \bottomrule
          \multicolumn{5}{p{12.5cm}}{\footnotesize \emph{Notes}: Robust standard errors clustered at the ethnic homeland x region x round level for the border samples are in parentheses. The set of individual controls
          includes measures of: normalized distance from the largest non-capital city, age, age squared, sex,
          education, employment status, rural/urban situation, personal economic conditions perception, interest in politics, TV news consumption, radio news consumption, newspaper consumption. Border samples regressions also include a measure of ethnic discrimination. The set of regional controls includes measures of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
          \end{tabular}
          }
          \end{table}

          \begin{table}[H]
            \sffamily
            \caption{{Effect of distance from the capital on political trust - Media consumption}}
            \centering
            %\footnotesize
            \resizebox{18cm}{!}{
            \begin{tabular}{@{\extracolsep{5pt}} l c c c c c c c c c c c c}
            \\
            \toprule
            \toprule
            &  \multicolumn{12}{c}{{OLS}}  \\
            \cmidrule(r){2-13}
            &  \multicolumn{4}{c}{{TV news}}&  \multicolumn{4}{c}{{Radio news}}&  \multicolumn{4}{c}{{Newspaper}}  \\
            \cmidrule(r){2-5}
            \cmidrule(r){6-9}
            \cmidrule(r){10-13}
            & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}} & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}} & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}}\\
                \cmidrule(r){2-3}
                \cmidrule(r){4-5}
                \cmidrule(r){6-7}
                \cmidrule(r){8-9}
                \cmidrule(r){10-11}
                \cmidrule(r){12-13}
                & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}} &  \multicolumn{1}{c}{{(5)}} &  \multicolumn{1}{c}{{(6)}} &  \multicolumn{1}{c}{{(7)}} &  \multicolumn{1}{c}{{(8)}} &  \multicolumn{1}{c}{{(9)}} &  \multicolumn{1}{c}{{(10)}} &  \multicolumn{1}{c}{{(11)}} &  \multicolumn{1}{c}{{(12)}}\\
             \midrule
             Distance from the capital&      -0.536***&      -0.531***&      -0.371** &      -0.361** &       0.057   &       0.007   &      -0.221   &      -0.306 &      -0.113***&      -0.081***&      -0.079   &       0.075  \\
             \medskip
             &      (0.04)   &      (0.04)   &      (0.15)   &      (0.18)   &      (0.05)   &      (0.04)   &      (0.16)   &      (0.20)   &      (0.04)   &      (0.03)   &      (0.10)   &      (0.12) \\

        \midrule
             \smallskip
            Individual \& regional controls  & Yes & Yes & Yes & Yes  & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\
            \smallskip
            Country controls & Yes& No& Yes& No& Yes& No& Yes& No& Yes& No& Yes& No\\
            \smallskip
            Round FE & Yes & No& Yes & No & Yes & No& Yes & No & Yes & No& Yes & No\\
            \smallskip
            Country X Round FE       & No & Yes& No & Yes& No & Yes& No & Yes& No & Yes& No & Yes\\
            \smallskip
            Ethnic homeland FE & No & No & Yes& Yes& No & No & Yes& Yes& No & No & Yes& Yes\\
            \smallskip
            Observations            &      108,766   &      113,243   &       10,976   &       121,15  &  108,766   &      113,243   &       10,976   &       12,115  &      108,766   &      113,243   &       10,976   &       12,115\\
            Adjusted-R$^2$         &   0.458   &       0.469   &       0.436   &       0.445 &       0.151   &       0.181   &       0.178   &       0.196 &       0.360   &       0.378   &       0.342   &       0.349  \\
                                  \bottomrule
            \multicolumn{13}{p{29cm}}{\footnotesize \emph{Notes}: The border sample includes individuals residing within a 40-kilometer buffer around a country border that overlaps with a historical ethnic homeland, as defined by Murdock (1959). Robust standard errors clustered at the region x round level  for the base sample and ethnic homeland x region x round level for the border sample are in parentheses. The set of individual controls
            includes measures of: normalized distance from the largest non-capital city, age, age squared, sex,
            education, employment status, rural/urban situation, personal economic conditions perception, interest in politics. Each media consumption measure (TV, radio, newspaper) is excluded from controls when it is the dependent variable. Border samples regressions also include a measure of ethnic discrimination. The set of regional controls includes measures of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
            \end{tabular}
            }
            \end{table}

  \begin{table}[H]
    \sffamily
    \caption{{Effect of distance from the capital on internet news}}
    \centering
    %\footnotesize
    \resizebox{8cm}{!}{
    \begin{tabular}{@{\extracolsep{5pt}} l c c c c}
    \\
    \toprule
    \toprule
    &  \multicolumn{4}{c}{{OLS}}  \\
    \cmidrule(r){2-5}
    &  \multicolumn{4}{c}{{Internet news}}  \\
    \cmidrule(r){2-5}
    & \multicolumn{2}{c}{{Base sample}} & \multicolumn{2}{c}{{Border sample}}\\
        \cmidrule(r){2-3}
        \cmidrule(r){4-5}
        & \multicolumn{1}{c}{{(1)}} &  \multicolumn{1}{c}{{(2)}}  & \multicolumn{1}{c}{{(3)}} &  \multicolumn{1}{c}{{(4)}}\\
     \midrule

     Distance from the capital&      -0.080***&      -0.065** &       0.049   &       0.074 \\
     \medskip
     &      (0.03)   &      (0.03)   &      (0.09)   &      (0.09)\\

\midrule
     \smallskip
    Individual \& regional controls  & Yes & Yes & Yes & Yes\\
    \smallskip
    Country controls & Yes& No& Yes& No\\
    \smallskip
    Round FE & Yes & No& Yes & No\\
    \smallskip
    Country X Round FE       & No & Yes& No & Yes\\
    \smallskip
    Ethnic homeland FE & No & No & Yes& Yes\\
    \smallskip
    Observations           &      107,592   &      112,051   &       108,86   &       12,019\\
    Adjusted-R$^2$          &       0.422   &       0.432   &       0.387   &       0.386\\
                          \bottomrule
    \multicolumn{5}{p{12cm}}{\footnotesize \emph{Notes}: The border sample includes individuals residing within a 40-kilometer buffer around a country border that overlaps with a historical ethnic homeland, as defined by Murdock (1959). Robust standard errors clustered at the region x round level  for the base sample and ethnic homeland x region x round level for the border sample are in parentheses. The set of individual controls
    includes measures of: normalized distance from the largest non-capital city, age, age squared, sex,
    education, employment status, rural/urban situation, personal economic conditions perception, interest in politics, TV news consumption, radio news consumption, newspaper consumption. Border samples regressions also include a measure of ethnic discrimination. The set of regional controls includes measures of: nighttime light, population density, region area, president birthplace dummy. The set of country controls includes: log(GDP.p.c.), log(area), V-Dem Polyarchy index, World Bank corruption index, political regime type, colonial origin. *** / ** / * represent significance at the 0.01 / 0.05 / 0.10 levels, respectively.}
    \end{tabular}
    }
    \end{table}
\end{document}