### 2025-02-24 writing/presentation/Untitled-1.tex

\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{positioning} % Load the positioning library

\begin{document}

\centering    
\begin{tikzpicture}
    \tikzset{
      node distance=5mm and 0mm,
      boite/.style={fill=blue!50!cyan!70,draw,align=center,
        font=\footnotesize\bfseries,text=white},
      boite coins ronds/.style={boite,rounded corners=3pt},
      boite circulaire/.style={boite,circle},
      fleche/.style={
        line cap=round,-latex,line width=0.25mm,
        draw=blue!50!cyan!30},
      grosse fleche/.style={fleche,line width=1mm},
    }

    \node[boite coins ronds]  (one) {a\\Texte vraiment très long};    
    \node[boite coins round ,below=of one]  (a) {a\\Texte vraiment très long}; % Changed to use orange boite style
    \node[boite coins ronds,below left=of a] (b) {b\\Texte vraiment\\très long};
    \node[boite coins ronds,below right=of a] (c) {c\\Texte vraiment\\très long};
    \node[boite coins ronds,below=of c] (d) {d\\Texte vraiment\\très long};
    \node[boite coins ronds,below =of b] (e) {c\\Texte vraiment\\très long};
    \node[draw=none, fill=none] (dummy) at (0,-5) {}; % Create a dummy node at a specific position, not visible
    \node[boite coins ronds, below=of dummy] (trois) {3\\Texte vraiment\\très long};

    \begin{scope}[on background layer]
      \draw[fleche,shorten >=2mm]
      (a.west) -- (b.north) node[midway,left=4mm]{};
      \draw[fleche,shorten >=2mm]
      (a.east) -- (c.north) node[midway,right=4mm]{bla bla bla};
    \end{scope}
\end{tikzpicture}

\end{document}