### 2025-02-24 script/r/draft.Rmd

```{r}
library(ggplot2)
library(dplyr)
library(maps)

# Mapping of ISO codes to country names
iso_to_country <- data.frame(
  iso = c("AGO", "BDI", "BEN", "BFA", "BWA", "CIV", "CMR", "CPV", "DZA", "EGY", 
          "ETH", "GAB", "GHA", "GIN", "GMB", "KEN", "LBR", "LSO", "MAR", "MDG", 
          "MLI", "MOZ", "MUS", "MWI", "NAM", "NER", "NGA", "SDN", "SEN", "SLE", 
          "STP", "SWZ", "TGO", "TUN", "TZA", "UGA", "ZAF", "ZMB", "ZWE"),
  country = c("Angola", "Burundi", "Benin", "Burkina Faso", "Botswana", "Ivory Coast", 
              "Cameroon", "Cape Verde", "Algeria", "Egypt", "Ethiopia", "Gabon", 
              "Ghana", "Guinea", "Gambia", "Kenya", "Liberia", "Lesotho", "Morocco", 
              "Madagascar", "Mali", "Mozambique", "Mauritius", "Malawi", "Namibia", 
              "Niger", "Nigeria", "Sudan", "Senegal", "Sierra Leone", "Sao Tome and Principe", 
              "Eswatini", "Togo", "Tunisia", "Tanzania", "Uganda", "South Africa", 
              "Zambia", "Zimbabwe")
)

# Count observations for each country based on ISO codes
country_counts <- data_final_adm %>%
  group_by(iso) %>%
  summarize(observation_count = n())

# Filter countries with at least 1000 observations and match with ISO codes
highlight_countries <- country_counts %>%
  filter(observation_count >= 1000, iso %in% iso_to_country$iso) %>%
  pull(iso)

cty <- iso_to_country%>%
  pull(country)

# Get the world map data
africa_map <- map_data("world")%>%
  filter(region%in%cty)

# Add a column for highlighting based on country names
africa_map <- africa_map %>%
  left_join(iso_to_country, by = c("region" = "country")) %>%
  mutate(highlight = ifelse(iso %in% highlight_countries, "Yes", "No"))

# Plot
ggplot() +
  geom_polygon(data = africa_map, aes(x = long, y = lat, group = group, fill = highlight), 
               color = "black") +
  scale_fill_manual(values = c("Yes" = "lightblue", "No" = "gray"), guide = "none") +
  labs(title = "Map of Africa: Countries with At Least 1000 Observations Highlighted") +
  theme_minimal()

```

```{r}

afro_diff <- anti_join(afro_stack, data_final)

afro_diff%>%
  distinct(iso)

afro_diff %>%
  select(latitude, longitude)%>%
  View()

data_final%>%
  select(dist_cap_log, dist_ncap_log, dist_sndncap_log)%>%
  summary()

```

```{r}
gov_index <- read_excel("data/data_raw/gov_index.xlsx")%>%
  rename(iso=ISO)
polstab_index <- read_excel("data/data_raw/polstab_index.xlsx")%>%
  rename(iso=ISO)
reg_index <- read_excel("data/data_raw/reg_index.xlsx")%>%
  rename(iso=ISO)
rule_index <- read_excel("data/data_raw/rule_index.xlsx")%>%
  rename(iso=ISO)
voiceac_index <- read_excel("data/data_raw/voiceac_index.xlsx")%>%
  rename(iso=ISO)

vdem <- readRDS("data/data_raw/V-Dem-CY-Full+Others-v13.rds")

```

```{r}
data_final_adm%>% filter(iso=="BEN") %>% select(pop_adm1, pop_adm2, pop_adm3)%>%summary()

data_final_adm %>% select(gg_adm1_pot, round)%>%filter(round==6)%>%summary()

data_final_adm%>%filter(iso=="BEN")%>%select(iso, round, gg_adm1_pot, NAME_1)%>%View()
```

```{r}
ggg_stack <- readRDS("data/data_rds/ggg_stack.rds")

ggg_cty<-ggg_stack%>%distinct(iso)%>%pull(iso)

all_rounds <- seq(min(ggg_stack$round, na.rm = TRUE), max(ggg_stack$round, na.rm = TRUE), by = 1)

for (country in ggg_cty) {
  ggg_stack %>%
    filter(iso == country) %>%
    group_by(round) %>%
    summarize(
      ggg_adm1_mean = mean(ggg_adm1_pot, na.rm = TRUE),
      ggg_adm2_mean = mean(ggg_adm2_pot, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    complete(round = all_rounds, fill = list(ggg_adm1_mean = NA, ggg_adm2_mean = NA)) %>%  # Fill missing rounds with NA
    ggplot(aes(x = round)) +
    geom_line(aes(y = ggg_adm1_mean), size = 1, color = "blue") +
    geom_point(aes(y = ggg_adm1_mean), size = 2, color = "blue") +
    geom_line(aes(y = ggg_adm2_mean), size = 1, color = "red") +
    geom_point(aes(y = ggg_adm2_mean), size = 2, color = "red") +
    labs(
      title = paste("Evolution of ggg_adm1_pot and ggg_adm2_pot Across Rounds for", country),
      x = "Round",
      y = "Mean Values"
    ) +
    scale_y_continuous(limits = c(0, 1)) +  # Fixed y-axis scale between 0 and 1
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 16),
      axis.title = element_text(size = 14),
      legend.position = "none"
    ) -> p
  print(p)
}
```

```{r}
setwd("~/working_paper/trust/")
getwd()
```

```{r packages, echo=FALSE, }
library(haven)
library(dplyr)
select <- dplyr::select
library(magrittr)
library(sf)
```

```{r}
countries <- c("BEN", "BFA", "BWA", "CIV", "CMR", "GIN", "LBR", "MLI", "NAM", "NER", "SLE", "GHA", "KEN", "MOZ", "MWI", "NGA", "TZA", "UGA", "ZMB", "ZWE")

data_final <- readRDS("data/data_rds/data_final.rds")

data_final_sf <- subset(data_final, select = c("id_unique","iso","latitude","longitude"))

data_final_sf <- st_as_sf(data_final_sf, coords = c("longitude", "latitude"), crs = 4326)

result <- list()

for (country in countries) {
  shapefile <- file.path("C:/Users/Redha CHABA/Documents/working_paper/trust/data/data_raw/road_shp", paste0("routes_", country, ".shp"))

  if (file.exists(shapefile)) {
    roads_sf <- st_read(shapefile)
    roads_sf <- st_transform(roads_sf, st_crs(data_final_sf))

    data_final_sf$distance_to_road <- st_distance(data_final_sf, roads_sf) %>% apply(1, min)

    country_data_result <- as.data.frame(data_final_sf)
    country_data_result <- country_data_result %>%
      select(id_unique, distance_to_road)

    result[[country]] <- country_data_result
  } else {
    warning(paste("Shapefile not found for country:", country))
  }
}

# Combine results into a single data frame
result_combined <- do.call(rbind, lapply(names(result), function(country) {
  cbind(result[[country]], country = country)
}))

final_result_road <- do.call(rbind, result)

final_df6 <- merge(final_df6,final_result_road, by="id_unique")

remove(bdd, country_data, country_data_result, country_data_sf, result, roads_sf, countries, ISO, shapefile)

```

```{r}
tiff_file <- "C:/Users/Redha CHABA/Desktop/temp/desert.tif"

desert_wgs84 <- rast(tiff_file)

print(desert_wgs84)

binary_raster <- ifel(!is.na(desert_wgs84), 1, 0)

print(binary_raster)

freq(binary_raster)

plot(binary_raster, main = "Desert Map", col = terrain.colors(100))

desert_r5_adm1 <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(binary_raster), crs(country_shp))) {
    country_shp <- project(country_shp, crs(binary_raster))
  }

  binary_raster_cropped <- crop(binary_raster, country_shp)
  binary_raster_masked <- mask(binary_raster_cropped, country_shp)

  desert_adm1 <- exactextractr::exact_extract(binary_raster_masked, regions, 'mean')

  regions$desert_adm1 <- desert_adm1
  regions_df <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, desert_adm1) %>%
    mutate(round = 5)

  desert_r5_adm1 <- rbind(desert_r5_adm1, regions_df)
}

desert_r5_adm1 %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(desert_r5_adm1))

saveRDS(desert_r5_adm1, file="data/data_rds/geo_ctrl/desert_r5_adm1.rds")

desert_r5_adm2 <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(binary_raster), crs(country_shp))) {
    country_shp <- project(country_shp, crs(binary_raster))
  }

  binary_raster_cropped <- crop(binary_raster, country_shp)
  binary_raster_masked <- mask(binary_raster_cropped, country_shp)

  desert_adm2 <- exactextractr::exact_extract(binary_raster_masked, regions, 'mean')

  regions$desert_adm2 <- desert_adm2
  regions_df <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_2, desert_adm2) %>%
    mutate(round = 5)

  desert_r5_adm2 <- rbind(desert_r5_adm2, regions_df)
}

desert_r5_adm2 %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(desert_r5_adm2))

saveRDS(desert_r5_adm2, file="data/data_rds/geo_ctrl/desert_r5_adm2.rds")
```

```{r}
library(ggplot2)

ggplot(data_final_nlp_bis2%>%filter(!is.na(neig_eth_a)), aes(x = factor(neig_eth_a))) +
  geom_bar(fill = "steelblue") +
  labs(title = "Willingness to Have Neighbors of Different Ethnicities - Waves 6-8",
       y = "Count",
       x=NULL) +
  scale_x_discrete(labels = c("Strongly dislike", "Somewhat dislike","Would not care", "Somewhat like", "Strongly like")) +
  theme_minimal()

ggplot(data_final_nlp_bis2%>%filter(!is.na(neig_eth_a), spl==1), aes(x = factor(neig_eth_a))) +
  geom_bar(fill = "steelblue") +
  labs(title = "Willingness to Have Neighbors of Different Ethnicities - Waves 6-8",
       y = "Count",
       x=NULL) +
  scale_x_discrete(labels = c("Strongly dislike", "Somewhat dislike","Would not care", "Somewhat like", "Strongly like")) +
  theme_minimal()
```