### 2025-02-27 script/r/draft.Rmd
merge geo controls

```{r}
setwd("~/working_paper/trust/")
getwd()
```

```{r packages, echo=FALSE, }
library(haven)
library(dplyr)
select <- dplyr::select
library(magrittr)
library(stringr)
library(geosphere)
library(purrr)
library(kableExtra)
library(tidyr)
library(readxl)
library(sf)
library(exactextractr)
library(raster)
library(rgdal)
library(ncdf4)
library(terra)
```

#Levels

```{r}
folder_path <- "data/data_raw/gadm"

files <- list.files(folder_path, full.names = TRUE)

levels <- data.frame(file = files) %>%
  mutate(country = str_extract(file, "(?<=_)[A-Z]{3}(?=_\\d)"),
         level = as.integer(str_extract(file, "(?<=_)[0-9]+(?=\\.shp$)"))) %>%
  arrange(country, desc(level)) %>%
  distinct(country, .keep_all = TRUE)

levels %<>%
  group_by(country) %>%
  summarise(levels_nb = first(level))

print(levels, n=40)

cty_adm1 <- levels%>%
  filter(levels_nb>=1)%>%
  pull(country)

cty_adm2 <- levels%>%
  filter(levels_nb>=2)%>%
  pull(country)

cty_adm3 <- levels%>%
  filter(levels_nb>=3)%>%
  pull(country)

cty_adm4 <- levels%>%
  filter(levels_nb>=4)%>%
  pull(country)
```

```{r}
library(ggplot2)
library(dplyr)
library(maps)

# Mapping of ISO codes to country names
iso_to_country <- data.frame(
  iso = c("AGO", "BDI", "BEN", "BFA", "BWA", "CIV", "CMR", "CPV", "DZA", "EGY", 
          "ETH", "GAB", "GHA", "GIN", "GMB", "KEN", "LBR", "LSO", "MAR", "MDG", 
          "MLI", "MOZ", "MUS", "MWI", "NAM", "NER", "NGA", "SDN", "SEN", "SLE", 
          "STP", "SWZ", "TGO", "TUN", "TZA", "UGA", "ZAF", "ZMB", "ZWE"),
  country = c("Angola", "Burundi", "Benin", "Burkina Faso", "Botswana", "Ivory Coast", 
              "Cameroon", "Cape Verde", "Algeria", "Egypt", "Ethiopia", "Gabon", 
              "Ghana", "Guinea", "Gambia", "Kenya", "Liberia", "Lesotho", "Morocco", 
              "Madagascar", "Mali", "Mozambique", "Mauritius", "Malawi", "Namibia", 
              "Niger", "Nigeria", "Sudan", "Senegal", "Sierra Leone", "Sao Tome and Principe", 
              "Eswatini", "Togo", "Tunisia", "Tanzania", "Uganda", "South Africa", 
              "Zambia", "Zimbabwe")
)

# Count observations for each country based on ISO codes
country_counts <- data_final_adm %>%
  group_by(iso) %>%
  summarize(observation_count = n())

# Filter countries with at least 1000 observations and match with ISO codes
highlight_countries <- country_counts %>%
  filter(observation_count >= 1000, iso %in% iso_to_country$iso) %>%
  pull(iso)

cty <- iso_to_country%>%
  pull(country)

# Get the world map data
africa_map <- map_data("world")%>%
  filter(region%in%cty)

# Add a column for highlighting based on country names
africa_map <- africa_map %>%
  left_join(iso_to_country, by = c("region" = "country")) %>%
  mutate(highlight = ifelse(iso %in% highlight_countries, "Yes", "No"))

# Plot
ggplot() +
  geom_polygon(data = africa_map, aes(x = long, y = lat, group = group, fill = highlight), 
               color = "black") +
  scale_fill_manual(values = c("Yes" = "lightblue", "No" = "gray"), guide = "none") +
  labs(title = "Map of Africa: Countries with At Least 1000 Observations Highlighted") +
  theme_minimal()

```

```{r}

afro_diff <- anti_join(afro_stack, data_final)

afro_diff%>%
  distinct(iso)

afro_diff %>%
  select(latitude, longitude)%>%
  View()

data_final%>%
  select(dist_cap_log, dist_ncap_log, dist_sndncap_log)%>%
  summary()

```

```{r}
gov_index <- read_excel("data/data_raw/gov_index.xlsx")%>%
  rename(iso=ISO)
polstab_index <- read_excel("data/data_raw/polstab_index.xlsx")%>%
  rename(iso=ISO)
reg_index <- read_excel("data/data_raw/reg_index.xlsx")%>%
  rename(iso=ISO)
rule_index <- read_excel("data/data_raw/rule_index.xlsx")%>%
  rename(iso=ISO)
voiceac_index <- read_excel("data/data_raw/voiceac_index.xlsx")%>%
  rename(iso=ISO)

vdem <- readRDS("data/data_raw/V-Dem-CY-Full+Others-v13.rds")

```

```{r}
data_final_adm%>% filter(iso=="BEN") %>% select(pop_adm1, pop_adm2, pop_adm3)%>%summary()

data_final_adm %>% select(gg_adm1_pot, round)%>%filter(round==6)%>%summary()

data_final_adm%>%filter(iso=="BEN")%>%select(iso, round, gg_adm1_pot, NAME_1)%>%View()
```

```{r}
ggg_stack <- readRDS("data/data_rds/ggg_stack.rds")

ggg_cty<-ggg_stack%>%distinct(iso)%>%pull(iso)

all_rounds <- seq(min(ggg_stack$round, na.rm = TRUE), max(ggg_stack$round, na.rm = TRUE), by = 1)

for (country in ggg_cty) {
  ggg_stack %>%
    filter(iso == country) %>%
    group_by(round) %>%
    summarize(
      ggg_adm1_mean = mean(ggg_adm1_pot, na.rm = TRUE),
      ggg_adm2_mean = mean(ggg_adm2_pot, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    complete(round = all_rounds, fill = list(ggg_adm1_mean = NA, ggg_adm2_mean = NA)) %>%  # Fill missing rounds with NA
    ggplot(aes(x = round)) +
    geom_line(aes(y = ggg_adm1_mean), size = 1, color = "blue") +
    geom_point(aes(y = ggg_adm1_mean), size = 2, color = "blue") +
    geom_line(aes(y = ggg_adm2_mean), size = 1, color = "red") +
    geom_point(aes(y = ggg_adm2_mean), size = 2, color = "red") +
    labs(
      title = paste("Evolution of ggg_adm1_pot and ggg_adm2_pot Across Rounds for", country),
      x = "Round",
      y = "Mean Values"
    ) +
    scale_y_continuous(limits = c(0, 1)) +  # Fixed y-axis scale between 0 and 1
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 16),
      axis.title = element_text(size = 14),
      legend.position = "none"
    ) -> p
  print(p)
}
```

```{r}
setwd("~/working_paper/trust/")
getwd()
```

```{r packages, echo=FALSE, }
library(haven)
library(dplyr)
select <- dplyr::select
library(magrittr)
library(sf)
```

```{r}
countries <- c("BEN", "BFA", "BWA", "CIV", "CMR", "GIN", "LBR", "MLI", "NAM", "NER", "SLE", "GHA", "KEN", "MOZ", "MWI", "NGA", "TZA", "UGA", "ZMB", "ZWE")

data_final <- readRDS("data/data_rds/data_final.rds")

data_final_sf <- subset(data_final, select = c("id_unique","iso","latitude","longitude"))

data_final_sf <- st_as_sf(data_final_sf, coords = c("longitude", "latitude"), crs = 4326)

result <- list()

for (country in countries) {
  shapefile <- file.path("C:/Users/Redha CHABA/Documents/working_paper/trust/data/data_raw/road_shp", paste0("routes_", country, ".shp"))

  if (file.exists(shapefile)) {
    roads_sf <- st_read(shapefile)
    roads_sf <- st_transform(roads_sf, st_crs(data_final_sf))

    data_final_sf$distance_to_road <- st_distance(data_final_sf, roads_sf) %>% apply(1, min)

    country_data_result <- as.data.frame(data_final_sf)
    country_data_result <- country_data_result %>%
      select(id_unique, distance_to_road)

    result[[country]] <- country_data_result
  } else {
    warning(paste("Shapefile not found for country:", country))
  }
}

# Combine results into a single data frame
result_combined <- do.call(rbind, lapply(names(result), function(country) {
  cbind(result[[country]], country = country)
}))

final_result_road <- do.call(rbind, result)

final_df6 <- merge(final_df6,final_result_road, by="id_unique")

remove(bdd, country_data, country_data_result, country_data_sf, result, roads_sf, countries, ISO, shapefile)

```

```{r}
tiff_file <- "data/data_raw/geography/desert/desert.tif"

desert_wgs84 <- rast(tiff_file)

print(desert_wgs84)

africa_extent <- ext(-30, 60, -40, 40)

desert_africa <- crop(desert_wgs84, africa_extent)

binary_raster <- ifel(!is.na(desert_africa), 1, 0)

print(binary_raster)

plot(binary_raster, main = "Desert Map", col = rev(heat.colors(100)))

desert_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(binary_raster), crs(country_shp))) {
    country_shp <- project(country_shp, crs(binary_raster))
  }

  binary_raster_cropped <- crop(binary_raster, country_shp)
  binary_raster_masked <- mask(binary_raster_cropped, country_shp)

  desert_adm1 <- exactextractr::exact_extract(binary_raster_masked, regions, 'mean')

  regions$desert_adm1 <- desert_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, desert_adm1)

  desert_adm1_df <- rbind(desert_adm1_df, country_results)
}

desert_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(desert_adm1_df))

saveRDS(desert_adm1_df, file="data/data_rds/geo_ctrl/desert_adm1.rds")

desert_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(binary_raster), crs(country_shp))) {
    country_shp <- project(country_shp, crs(binary_raster))
  }

  binary_raster_cropped <- crop(binary_raster, country_shp)
  binary_raster_masked <- mask(binary_raster_cropped, country_shp)

  desert_adm2 <- exactextractr::exact_extract(binary_raster_masked, regions, 'mean')

  regions$desert_adm2 <- desert_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, desert_adm2)

  desert_adm2_df <- rbind(desert_adm2_df, country_results)
}

desert_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(desert_adm2_df))

saveRDS(desert_adm2_df, file="data/data_rds/geo_ctrl/desert_adm2.rds")
```

```{r}
tiff_file <- "data/data_raw/geography/elevation/elevation.tif"

elevation_wgs84 <- rast(tiff_file)

print(elevation_wgs84)

plot(elevation_wgs84, main = "Elevation Map", col = terrain.colors(100), breaks = seq(-1000, 3000, by = 100))

elevation_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(elevation_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(elevation_wgs84))
  }

  elevation_wgs84_cropped <- crop(elevation_wgs84, country_shp)
  elevation_wgs84_masked <- mask(elevation_wgs84_cropped, country_shp)

  elevation_adm1 <- exactextractr::exact_extract(elevation_wgs84_masked, regions, 'mean')

  regions$elevation_adm1 <- elevation_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, elevation_adm1)

  elevation_adm1_df <- rbind(elevation_adm1_df, country_results)
}

elevation_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(elevation_adm1_df))

saveRDS(elevation_adm1_df, file="data/data_rds/geo_ctrl/elevation_adm1.rds")

elevation_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(elevation_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(elevation_wgs84))
  }

  elevation_wgs84_cropped <- crop(elevation_wgs84, country_shp)
  elevation_wgs84_masked <- mask(elevation_wgs84_cropped, country_shp)

  elevation_adm2 <- exactextractr::exact_extract(elevation_wgs84_masked, regions, 'mean')

  regions$elevation_adm2 <- elevation_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, elevation_adm2)

  elevation_adm2_df <- rbind(elevation_adm2_df, country_results)
}

elevation_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(elevation_adm2_df))

saveRDS(elevation_adm2_df, file="data/data_rds/geo_ctrl/elevation_adm2.rds")
```

```{r}
library(viridis)

tiff_file <- "data/data_raw/geography/precipitation/CHELSA_bio12_1981-2010_V.2.1.tif"

precipitation_wgs84 <- rast(tiff_file)

print(precipitation_wgs84)

africa_extent <- ext(-30, 60, -40, 40)

precipitation_africa <- crop(precipitation_wgs84, africa_extent)

plot(precipitation_africa, main = "Precipitation Map of Africa", col = rev(viridis(100)))

precipitation_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(precipitation_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(precipitation_wgs84))
  }

  precipitation_wgs84_cropped <- crop(precipitation_wgs84, country_shp)
  precipitation_wgs84_masked <- mask(precipitation_wgs84_cropped, country_shp)

  precipitation_adm1 <- exactextractr::exact_extract(precipitation_wgs84_masked, regions, 'mean')

  regions$precipitation_adm1 <- precipitation_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, precipitation_adm1)

  precipitation_adm1_df <- rbind(precipitation_adm1_df, country_results)
}

precipitation_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(precipitation_adm1_df))

saveRDS(precipitation_adm1_df, file="data/data_rds/geo_ctrl/precipitation_adm1.rds")

precipitation_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(precipitation_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(precipitation_wgs84))
  }

  precipitation_wgs84_cropped <- crop(precipitation_wgs84, country_shp)
  precipitation_wgs84_masked <- mask(precipitation_wgs84_cropped, country_shp)

  precipitation_adm2 <- exactextractr::exact_extract(precipitation_wgs84_masked, regions, 'mean')

  regions$precipitation_adm2 <- precipitation_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, precipitation_adm2)

  precipitation_adm2_df <- rbind(precipitation_adm2_df, country_results)
}

precipitation_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(precipitation_adm2_df))

saveRDS(precipitation_adm2_df, file="data/data_rds/geo_ctrl/precipitation_adm2.rds")
```

```{r}
shapefile_directory <- "data/data_raw/gadm/"

shapefile_names <- list.files(shapefile_directory, pattern = "*_2.shp", full.names = TRUE)

adm2_shapefiles <- lapply(shapefile_names, st_read)
adm2_combined <- do.call(rbind, adm2_shapefiles)

adm2_combined_merged <- merge(adm2_combined, precipitation_adm2, by.x = "NAME_2", by.y = "NAME_2", all.x = TRUE)

library(ggplot2)

ggplot() +
  geom_sf(data = adm2_combined_merged, aes(fill = precipitation_adm2), color = "black", size = 0.25) +
  scale_fill_viridis(name = "Precipitation (mm)", option = "plasma", na.value = "grey") +
  labs(title = "Precipitation Map of Africa",
       subtitle = "Regional Precipitation Values") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    legend.position = "bottom"
  )

```

```{r}
tiff_file <- "data/data_raw/geography/temperature/CHELSA_bio1_1981-2010_V.2.1.tif"

temp_mean_wgs84 <- rast(tiff_file)

print(temp_mean_wgs84)

africa_extent <- ext(-30, 60, -40, 40)

temp_mean_africa <- crop(temp_mean_wgs84, africa_extent)

plot(temp_mean_africa, main = "temp_mean Map of Africa", col = rev(heat.colors(100)))

temp_mean_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(temp_mean_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_mean_wgs84))
  }

  temp_mean_wgs84_cropped <- crop(temp_mean_wgs84, country_shp)
  temp_mean_wgs84_masked <- mask(temp_mean_wgs84_cropped, country_shp)

  temp_mean_adm1 <- exactextractr::exact_extract(temp_mean_wgs84_masked, regions, 'mean')

  regions$temp_mean_adm1 <- temp_mean_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, temp_mean_adm1)

  temp_mean_adm1_df <- rbind(temp_mean_adm1_df, country_results)
}

temp_mean_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_mean_adm1_df))

saveRDS(temp_mean_adm1_df, file="data/data_rds/geo_ctrl/temp_mean_adm1.rds")

temp_mean_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(temp_mean_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_mean_wgs84))
  }

  temp_mean_wgs84_cropped <- crop(temp_mean_wgs84, country_shp)
  temp_mean_wgs84_masked <- mask(temp_mean_wgs84_cropped, country_shp)

  temp_mean_adm2 <- exactextractr::exact_extract(temp_mean_wgs84_masked, regions, 'mean')

  regions$temp_mean_adm2 <- temp_mean_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, temp_mean_adm2)

  temp_mean_adm2_df <- rbind(temp_mean_adm2_df, country_results)
}

temp_mean_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_mean_adm2_df))

saveRDS(temp_mean_adm2_df, file="data/data_rds/geo_ctrl/temp_mean_adm2.rds")
```

```{r}
tiff_file <- "data/data_raw/geography/temperature/CHELSA_bio5_1981-2010_V.2.1.tif"

temp_max_wgs84 <- rast(tiff_file)

print(temp_max_wgs84)

africa_extent <- ext(-30, 60, -40, 40)

temp_max_africa <- crop(temp_max_wgs84, africa_extent)

plot(temp_max_africa, main = "temp_max Map of Africa", col = rev(heat.colors(100)), breaks = seq(0, 50, by = 1))

temp_max_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(temp_max_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_max_wgs84))
  }

  temp_max_wgs84_cropped <- crop(temp_max_wgs84, country_shp)
  temp_max_wgs84_masked <- mask(temp_max_wgs84_cropped, country_shp)

  temp_max_adm1 <- exactextractr::exact_extract(temp_max_wgs84_masked, regions, 'mean')

  regions$temp_max_adm1 <- temp_max_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, temp_max_adm1)

  temp_max_adm1_df <- rbind(temp_max_adm1_df, country_results)
}

temp_max_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_max_adm1_df))

saveRDS(temp_max_adm1_df, file="data/data_rds/geo_ctrl/temp_max_adm1.rds")

temp_max_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(temp_max_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_max_wgs84))
  }

  temp_max_wgs84_cropped <- crop(temp_max_wgs84, country_shp)
  temp_max_wgs84_masked <- mask(temp_max_wgs84_cropped, country_shp)

  temp_max_adm2 <- exactextractr::exact_extract(temp_max_wgs84_masked, regions, 'mean')

  regions$temp_max_adm2 <- temp_max_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, temp_max_adm2)

  temp_max_adm2_df <- rbind(temp_max_adm2_df, country_results)
}

temp_max_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_max_adm2_df))

saveRDS(temp_max_adm2_df, file="data/data_rds/geo_ctrl/temp_max_adm2.rds")
```

```{r}
tiff_file <- "data/data_raw/geography/temperature/CHELSA_bio6_1981-2010_V.2.1.tif"

temp_min_wgs84 <- rast(tiff_file)

print(temp_min_wgs84)

africa_extent <- ext(-30, 60, -40, 40)

temp_min_africa <- crop(temp_min_wgs84, africa_extent)

plot(temp_min_africa, main = "temp_min Map of Africa", col = viridis(100), breaks = seq(-20, 30, by = 1))

temp_min_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(temp_min_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_min_wgs84))
  }

  temp_min_wgs84_cropped <- crop(temp_min_wgs84, country_shp)
  temp_min_wgs84_masked <- mask(temp_min_wgs84_cropped, country_shp)

  temp_min_adm1 <- exactextractr::exact_extract(temp_min_wgs84_masked, regions, 'mean')

  regions$temp_min_adm1 <- temp_min_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, temp_min_adm1)

  temp_min_adm1_df <- rbind(temp_min_adm1_df, country_results)
}

temp_min_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_min_adm1_df))

saveRDS(temp_min_adm1_df, file="data/data_rds/geo_ctrl/temp_min_adm1.rds")

temp_min_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(temp_min_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_min_wgs84))
  }

  temp_min_wgs84_cropped <- crop(temp_min_wgs84, country_shp)
  temp_min_wgs84_masked <- mask(temp_min_wgs84_cropped, country_shp)

  temp_min_adm2 <- exactextractr::exact_extract(temp_min_wgs84_masked, regions, 'mean')

  regions$temp_min_adm2 <- temp_min_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, temp_min_adm2)

  temp_min_adm2_df <- rbind(temp_min_adm2_df, country_results)
}

temp_min_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_min_adm2_df))

saveRDS(temp_min_adm2_df, file="data/data_rds/geo_ctrl/temp_min_adm2.rds")
```

```{r}
tiff_file <- "data/data_raw/geography/mountain/k3binary.tif"

mountain_wgs84 <- rast(tiff_file)

print(mountain_wgs84)

africa_extent <- ext(-30, 60, -40, 40)

mountain_africa <- crop(mountain_wgs84, africa_extent)

binary_raster <- ifel(!is.na(mountain_africa), 1, 0)

plot(binary_raster, main = "mountain Map of Africa", col = viridis(100))

mountain_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(binary_raster), crs(country_shp))) {
    country_shp <- project(country_shp, crs(binary_raster))
  }

  binary_raster_cropped <- crop(binary_raster, country_shp)
  binary_raster_masked <- mask(binary_raster_cropped, country_shp)

  mountain_adm1 <- exactextractr::exact_extract(binary_raster_masked, regions, 'mean')

  regions$mountain_adm1 <- mountain_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, mountain_adm1)

  mountain_adm1_df <- rbind(mountain_adm1_df, country_results)
}

mountain_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(mountain_adm1_df))

saveRDS(mountain_adm1_df, file="data/data_rds/geo_ctrl/mountain_adm1.rds")

mountain_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(binary_raster), crs(country_shp))) {
    country_shp <- project(country_shp, crs(binary_raster))
  }

  binary_raster_cropped <- crop(binary_raster, country_shp)
  binary_raster_masked <- mask(binary_raster_cropped, country_shp)

  mountain_adm2 <- exactextractr::exact_extract(binary_raster_masked, regions, 'mean')

  regions$mountain_adm2 <- mountain_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, mountain_adm2)

  mountain_adm2_df <- rbind(mountain_adm2_df, country_results)
}

mountain_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(mountain_adm2_df))

saveRDS(mountain_adm2_df, file="data/data_rds/geo_ctrl/mountain_adm2.rds")
```

```{r}
data_final_r <- readRDS("data/data_rds/data_final_r.rds")

desert_adm1 <- readRDS("data/data_rds/geo_ctrl/desert_adm1.rds")
desert_adm2 <- readRDS("data/data_rds/geo_ctrl/desert_adm2.rds")

desert <- merge(desert_adm1, desert_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(desert_adm1, desert_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(desert)))

expanded_data <- desert[rep(1:nrow(desert), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo <- expanded_data

rm(desert)

elevation_adm1 <- readRDS("data/data_rds/geo_ctrl/elevation_adm1.rds")
elevation_adm2 <- readRDS("data/data_rds/geo_ctrl/elevation_adm2.rds")

elevation <- merge(elevation_adm1, elevation_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(elevation_adm1, elevation_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(elevation)))

expanded_data <- elevation[rep(1:nrow(elevation), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo %<>% left_join(expanded_data, by=c("iso", "NAME_1", "NAME_2", "round"))

rm(elevation)

mountain_adm1 <- readRDS("data/data_rds/geo_ctrl/mountain_adm1.rds")
mountain_adm2 <- readRDS("data/data_rds/geo_ctrl/mountain_adm2.rds")

mountain <- merge(mountain_adm1, mountain_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(mountain_adm1, mountain_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(mountain)))

expanded_data <- mountain[rep(1:nrow(mountain), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo %<>% left_join(expanded_data, by=c("iso", "NAME_1", "NAME_2", "round"))

rm(mountain)

precipitation_adm1 <- readRDS("data/data_rds/geo_ctrl/precipitation_adm1.rds")
precipitation_adm2 <- readRDS("data/data_rds/geo_ctrl/precipitation_adm2.rds")

precipitation <- merge(precipitation_adm1, precipitation_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(precipitation_adm1, precipitation_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(precipitation)))

expanded_data <- precipitation[rep(1:nrow(precipitation), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo %<>% left_join(expanded_data, by=c("iso", "NAME_1", "NAME_2", "round"))

rm(precipitation)

temp_max_adm1 <- readRDS("data/data_rds/geo_ctrl/temp_max_adm1.rds")
temp_max_adm2 <- readRDS("data/data_rds/geo_ctrl/temp_max_adm2.rds")

temp_max <- merge(temp_max_adm1, temp_max_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(temp_max_adm1, temp_max_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(temp_max)))

expanded_data <- temp_max[rep(1:nrow(temp_max), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo %<>% left_join(expanded_data, by=c("iso", "NAME_1", "NAME_2", "round"))

rm(temp_max)

temp_min_adm1 <- readRDS("data/data_rds/geo_ctrl/temp_min_adm1.rds")
temp_min_adm2 <- readRDS("data/data_rds/geo_ctrl/temp_min_adm2.rds")

temp_min <- merge(temp_min_adm1, temp_min_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(temp_min_adm1, temp_min_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(temp_min)))

expanded_data <- temp_min[rep(1:nrow(temp_min), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo %<>% left_join(expanded_data, by=c("iso", "NAME_1", "NAME_2", "round"))

rm(temp_min)

temp_mean_adm1 <- readRDS("data/data_rds/geo_ctrl/temp_mean_adm1.rds")
temp_mean_adm2 <- readRDS("data/data_rds/geo_ctrl/temp_mean_adm2.rds")

temp_mean <- merge(temp_mean_adm1, temp_mean_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(temp_mean_adm1, temp_mean_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(temp_mean)))

expanded_data <- temp_mean[rep(1:nrow(temp_mean), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo %<>% left_join(expanded_data, by=c("iso", "NAME_1", "NAME_2", "round"))

rm(temp_mean)

data_final_r %<>% left_join(data_geo, by=c("iso", "NAME_1", "NAME_2", "round"))

saveRDS(data_final_r, file="data/data_rds/data_27_05_25.rds")

write_dta(data_final_r, "data/data_dta/data_27_05_25.dta")
```