### 2025-03-14 script/r/discontinuity.Rmd
clean and maj

```{r}
setwd("~/working_paper/trust/")
getwd()

library(dplyr)
library(readr)
library(magrittr)
library(ggplot2)
library(sf)
library(rnaturalearth)
library(murdock)
library(haven)
```

```{r}
murdock <- murdock::murdock
murdock_sf <- st_as_sf(murdock)

countries <- ne_countries(country = c("Benin","Botswana","Burkina Faso","Cameroon","Ivory Coast",
                                    "Guinea","Kenya","Liberia","Malawi","Mali",
                                    "Mozambique","Namibia","Niger","Nigeria","Sierra Leone",
                                    "United Republic of Tanzania","Uganda","Zambia","Zimbabwe",
                                    "Ghana"))

countries_sf <- st_as_sf(countries)

boundaries <- st_boundary(countries_sf)

ggplot() +
  geom_sf(data = murdock_sf, aes(color = "Murdock"), fill = "transparent", size = 1) +
  geom_sf(data = boundaries, aes(color = "Country Borders"), lwd = 1) +
  scale_color_manual(
    values = c("Murdock" = "black", "Country Borders" = "red"),
    labels = c("Contemporary Country Borders", "Historical Ethnic Areas"),
    name = NULL 
  ) +
  theme_minimal(base_size = 15) +
  theme(
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),  
    panel.background = element_blank(),   
    axis.text = element_blank(),         
    axis.ticks = element_blank(),         
    axis.title = element_blank(),         
    legend.key = element_rect(fill = "transparent", color = NA),
    legend.text = element_text(size = 8),
    legend.position = c(0.25, 0.35)
  )

ggsave("C:/Users/Redha CHABA/Documents/wp_git/rbci/plots/discontinuity/murdock_map.jpg", width = 10, height = 6, dpi = 300)

buffer <- st_buffer(boundaries, dist = 25000)

ggplot() +
  geom_sf(data = murdock_sf, aes (color="Murdock"), fill = "transparent", size = 1) +
  geom_sf(data = buffer, aes(color = "boundaries"), lwd = 1) +
  scale_color_manual(
    values = c("Murdock" = "black", "boundaries" = "red"),
    labels = c("25km boundaries", "Historical Ethnic Areas"),
    name = NULL 
  ) +
  theme_minimal(base_size = 15) +
  theme(
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),  
    panel.background = element_blank(),   
    axis.text = element_blank(),         
    axis.ticks = element_blank(),         
    axis.title = element_blank(),         
    legend.key = element_rect(fill = "transparent", color = NA),
    legend.text = element_text(size = 8),
    legend.position = c(0.25, 0.35)
  )

ggsave("C:/Users/Redha CHABA/Documents/wp_git/rbci/plots/discontinuity/murdock_boundaries.jpg", width = 10, height = 6, dpi = 300)
```

```{r}
countries <- c("Benin","Botswana","Burkina Faso","Cameroon","Ivory Coast",
                                    "Guinea","Kenya","Liberia","Malawi","Mali",
                                    "Mozambique","Namibia","Niger","Nigeria","Sierra Leone",
                                    "United Republic of Tanzania","Uganda","Zambia","Zimbabwe",
                                    "Ghana")

buffer_distances <- c(50000, 40000, 30000, 25000, 20000, 15000, 10000)

# Loop over each country and buffer distance
for (country in countries) {

  # Select the countries surrounding the target country (excluding the current one)
  surrounding_countries <- countries[countries != country]
  surrounding_countries_sf <- st_as_sf(ne_countries(country = surrounding_countries))

  # Loop over each buffer distance
  for (dist in buffer_distances) {

    # Create buffer for the target country
    buffer_country <- st_buffer(subset(countries_sf, countries_sf$sovereignt == country), dist = dist)

    # Find the intersection with the surrounding countries
    inter_country <- st_intersection(buffer_country, surrounding_countries_sf)

    # Plot the intersection geometry
    plot(inter_country$geometry, main = paste(country, "-", dist / 1000, "km buffer"))
  }
}
```

```{r}
# Storage lists
intersections_list <- list()

# Loop through buffer distances and countries to store intersections
for (dist in buffer_distances) {
  intersections <- lapply(countries, function(country) {
    # Create buffer for each country at the specified distance
    buffer <- st_buffer(subset(countries_sf, countries_sf$sovereignt == country), dist = dist)
    # Get the surrounding countries excluding the current one
    surrounding_countries <- countries[countries != country]
    surrounding_countries_sf <- st_as_sf(ne_countries(country = surrounding_countries))
    # Perform intersection and return it
    st_intersection(buffer, surrounding_countries_sf)
  })

  # Combine all intersections for the current buffer distance
  intersections_list[[as.character(dist)]] <- do.call(rbind, intersections)
}

# Perform union on the combined intersections for each distance
unions_list <- lapply(intersections_list, function(intersections) {
  union <- st_union(intersections)
  st_as_sf(union)
})

# Example of plotting the 25km buffer result
ggplot() +
  geom_sf(data = murdock_sf, fill = "transparent", aes(color = "Murdock"), size = 1) +
  geom_sf(data = unions_list[["25000"]], aes(color = "boundaries"), lwd=1, fill="transparent") +
  scale_color_manual(
    values = c("Murdock" = "black", "boundaries" = "red"),
    labels = c("25km boundaries", "Historical Ethnic Areas"),
    name = NULL 
  ) +
  theme_minimal(base_size = 15) +
  theme(
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),  
    panel.background = element_blank(),   
    axis.text = element_blank(),         
    axis.ticks = element_blank(),         
    axis.title = element_blank(),         
    legend.key = element_rect(fill = "transparent", color = NA),
    legend.text = element_text(size = 8),
    legend.position = c(0.25, 0.35)
  )

ggsave("C:/Users/Redha CHABA/Documents/wp_git/rbci/plots/discontinuity/murdock_boundaries_inter.jpg", width = 10, height = 6, dpi = 300)
```

```{r}
data <- readRDS("data/data_rds/data_27_02_25.rds")
data%<>%
  ungroup()

bdd_geo <- read_csv("data/data_raw/bdd_geo_w8.csv")

data %<>%
  filter(spl==1)%>%
  left_join(bdd_geo, by = "id_unique") %>%
  select(-...1) %>%
  mutate(
    distance_border_km = distance_border / 1000,
    distance_border_km2 = distance_border_km^2,
    distance_border_km3 = distance_border_km^3
  )

border_sample_50_union <- unions_list[["50000"]]
border_sample_40_union <- unions_list[["40000"]]
border_sample_30_union <- unions_list[["30000"]]
border_sample_25_union <- unions_list[["25000"]]
border_sample_20_union <- unions_list[["20000"]]
border_sample_15_union <- unions_list[["15000"]]
border_sample_10_union <- unions_list[["10000"]]

data_sf <- st_as_sf(data, coords = c("longitude", "latitude"), crs = st_crs(border_sample_50_union))

# Define a function to filter data based on border polygons
filter_by_border <- function(data, borders) {
  lapply(borders, function(border) {
    border_valid <- st_make_valid(border)
    st_filter(data, border_valid)
  })
}

# List of border polygons
borders <- list(
  border_sample_50_union,
  border_sample_40_union,
  border_sample_30_union,
  border_sample_25_union,
  border_sample_20_union,
  border_sample_15_union,
  border_sample_10_union
)

# Filter data
filtered_data_list <- filter_by_border(data_sf, borders)

names(filtered_data_list) <- c("50", "40", "30", "25", "20", "15", "10")
sapply(filtered_data_list, nrow)

# Prepare Murdock map
murdock <- murdock_sf %>%
  select(NAME, geometry) %>%
  rename(murdock_names = NAME)

# Join filtered data with Murdock map
df_list <- lapply(filtered_data_list, function(data) {
  st_join(data, murdock)
})

# Plot
ggplot() +
  geom_sf(data = murdock_sf, fill = "transparent", color = "black", size = 1) +
  geom_sf(data = df_list$`20`, size = 0.75, shape = 19, color = "red") +
  ggtitle("Individuals living less than 50 km from a border and Murdock's (1956) historical ethnic homeland") +
  theme_void()

ggsave("C:/Users/Redha CHABA/Documents/wp_git/rbci/plots/discontinuity/murdock_boundaries_inter_spl.jpg", width = 10, height = 6, dpi = 300)
```

```{r}
# Load country boundaries
countries <- ne_countries(scale = "medium", continent = "Africa", returnclass = "sf")

# Prepare data
individuals <- data %>%
  select(longitude, latitude, iso) %>%
  na.omit()

capitals <- data %>%
  distinct(caplong, caplat, iso)%>%
  na.omit() %>%
  rename(longitude = caplong, latitude = caplat)

# Combine individuals and capitals into a single data frame for plotting
combined_points <- bind_rows(
  individuals %>% mutate(category = "observations"),
  capitals %>% mutate(category = "capitals")
)

selected_countries <- countries%>%
  filter(name%in%c("Benin","Botswana","Burkina Faso","Cameroon","Côte d'Ivoire",
                                    "Guinea","Kenya","Liberia","Malawi","Mali",
                                    "Mozambique","Namibia","Niger","Nigeria","Sierra Leone",
                                    "Tanzania","Uganda","Zambia","Zimbabwe",
                                    "Ghana"))

selected_capitals <- capitals %>%
  filter(iso%in% c("BEN", "BFA", "BWA", "CMR", "CIV", "GHA", "GIN", "KEN", "LBR", "MLI", "MWI", "MOZ", "NAM", "NER", "NGA", "SLE", "TZA", "UGA", "ZMB", "ZWE"))

# Plot
ggplot() +
  geom_sf(data = countries, fill = "white", color = "black") +
  theme_void() +
  geom_sf(data = selected_countries, fill = "#E94849", color = "black") +
    geom_point(data = selected_capitals, 
             aes(x = longitude, y = latitude), 
             size = 2, shape = 22, fill = "black", show.legend = FALSE)

ggsave("C:/Users/Redha CHABA/Documents/wp_git/rbci/plots/discontinuity/cty_spl.jpg", width = 10, height = 6, dpi = 300)
```

```{r}
selected_countries <- countries %>% 
 filter(iso_a3 %in% c("BEN", "NER"))

selected_capitals <- capitals %>%
 filter(iso%in% c("BEN", "NER"))

selected_murdock <- murdock_sf%>%
  filter(NAME=="DENDI")

ggplot() +
   geom_sf(data = selected_countries, fill = "white", color = "black") +
geom_sf(data = selected_murdock, fill = "green", color = "darkgreen", size = 1, alpha=0.5) +
 theme_void() + 
geom_point(data = selected_capitals, 
          aes(x = longitude, y = latitude, color = "Capitals"), 
         size = 2, shape = 22, fill = "red", show.legend = FALSE) +

geom_label(aes(x = 2.5, y = 10, label = "Benin"), color = "black", size = 6, fill = "lightgrey", label.size = 0.5) +
     geom_label(aes(x = 4, y = 14.5, label = "Niger"), color = "black", size = 6, fill = "lightgrey", label.size = 0.5) +

geom_text(aes(x = 2.4, y = 13.75, label = "Niamey"), color = "black", size = 5) +
geom_text(aes(x = 2.2, y = 6.6, label = "Porto-Novo"), color = "black", size = 5) +

  labs(color = NULL) +
   coord_sf(xlim = c(0, 5), ylim = c(6, 18), expand = FALSE)

ggsave("C:/Users/Redha CHABA/Documents/wp_git/rbci/plots/discontinuity/benin_niger.jpg", width = 10, height = 10, dpi = 300)
```

```{r}
process_df <- function(df, threshold) {
  df_filtered <- subset(df, distance_border < threshold)
  df_filtered$cluster_frontiere <- paste(df_filtered$murdock_names, df_filtered$round, sep = "_")
  return(df_filtered)
}

df_list <- lapply(names(df_list), function(name) {
  df <- df_list[[name]]
  threshold <- case_when(
  name == "50" ~ 50000,
  name == "40" ~ 40000,
  name == "30" ~ 30000,
  name == "25" ~ 25000,
  name == "20" ~ 20000,
  name == "15" ~ 15000,
  name == "10" ~ 10000,
  TRUE ~ NA_real_
)
  df_filtered <- process_df(df, threshold)
  return(df_filtered)
})

names(df_list) <- c("df_50", "df_40", "df_30", "df_25", "df_20", "df_15", "df_10")
```

```{r}
convert_labelled_to_factors <- function(df) {
  df %>%
    mutate(across(where(haven::is.labelled), haven::as_factor))
}

df_list <- lapply(df_list, convert_labelled_to_factors)

names(df_list) <- c("df_50", "df_40", "df_30", "df_25", "df_20", "df_15", "df_10")
```

```{r}
df_50 <- df_list$df_50

list_columns <- sapply(df_50, is.list)

df_50[list_columns] <- lapply(df_50[list_columns], function(x) sapply(x, toString))

df_50%<>%as.data.frame()

df_50$murdock_round <- paste0(df_50$murdock_names, df_50$round, sep="_")

df_50 %<>%
  group_by(murdock_names) %>%
  filter(n_distinct(iso) >= 2)%>%
  ungroup()

df_50 %<>%
  mutate(cluster_bdd = as.numeric(factor(paste(region_time, murdock_names, sep = "_"))))

saveRDS(df_50, file="data/data_rds/df_50.rds")

write_dta(df_50, "data/data_dta/df_50.dta")
```

```{r}
df_40 <- df_list$df_40

list_columns <- sapply(df_40, is.list)

df_40[list_columns] <- lapply(df_40[list_columns], function(x) sapply(x, toString))

df_40%<>%as.data.frame()

df_40$murdock_round <- paste0(df_40$murdock_names, df_40$round, sep="_")

df_40 %<>%
  group_by(murdock_names) %>%
  filter(n_distinct(iso) >= 2)%>%
  ungroup()

df_40 %<>%
  mutate(cluster_bdd = as.numeric(factor(paste(region_time, murdock_names, sep = "_"))))

saveRDS(df_40, file="data/data_rds/df_40.rds")

write_dta(df_40, "data/data_dta/df_40.dta")
```

```{r}
df_30 <- df_list$df_30

list_columns <- sapply(df_30, is.list)

df_30[list_columns] <- lapply(df_30[list_columns], function(x) sapply(x, toString))

df_30%<>%as.data.frame()

df_30$murdock_round <- paste0(df_30$murdock_names, df_30$round, sep="_")

df_30 %<>%
  group_by(murdock_names) %>%
  filter(n_distinct(iso) >= 2)%>%
  ungroup()

df_30 %<>%
  mutate(cluster_bdd = as.numeric(factor(paste(region_time, murdock_names, sep = "_"))))

saveRDS(df_30, file="data/data_rds/df_30.rds")
write_dta(df_30, "data/data_dta/df_30.dta")
```

```{r}
df_25 <- df_list$df_25

list_columns <- sapply(df_25, is.list)

df_25[list_columns] <- lapply(df_25[list_columns], function(x) sapply(x, toString))

df_25%<>%as.data.frame()

df_25$murdock_round <- paste0(df_25$murdock_names, df_25$round, sep="_")

df_25 %<>%
  group_by(murdock_names) %>%
  filter(n_distinct(iso) >= 2)%>%
  ungroup()

df_25 %<>%
  mutate(cluster_bdd = as.numeric(factor(paste(region_time, murdock_names, sep = "_"))))

saveRDS(df_25, file="data/data_rds/df_25.rds")

write_dta(df_25, "data/data_dta/df_25.dta")
```

```{r}
df_20 <- df_list$df_20

list_columns <- sapply(df_20, is.list)

df_20[list_columns] <- lapply(df_20[list_columns], function(x) sapply(x, toString))

df_20%<>%as.data.frame()

df_20$murdock_round <- paste0(df_20$murdock_names, df_20$round, sep="_")

df_20 %<>%
  group_by(murdock_names) %>%
  filter(n_distinct(iso) >= 2)%>%
  ungroup()

df_20 %<>%
  mutate(cluster_bdd = as.numeric(factor(paste(region_time, murdock_names, sep = "_"))))

saveRDS(df_20, file="data/data_rds/df_20.rds")

write_dta(df_20, "data/data_dta/df_20.dta")
```

```{r}
df_15 <- df_list$df_15

list_columns <- sapply(df_15, is.list)

df_15[list_columns] <- lapply(df_15[list_columns], function(x) sapply(x, toString))

df_15%<>%as.data.frame()

df_15$murdock_round <- paste0(df_15$murdock_names, df_15$round, sep="_")

df_15 %<>%
  group_by(murdock_names) %>%
  filter(n_distinct(iso) >= 2)%>%
  ungroup()

df_15 %<>%
  mutate(cluster_bdd = as.numeric(factor(paste(region_time, murdock_names, sep = "_"))))

saveRDS(df_15, file="data/data_rds/df_15.rds")

write_dta(df_15, "data/data_dta/df_15.dta")
```

```{r}
df_10 <- df_list$df_10

list_columns <- sapply(df_10, is.list)

df_10[list_columns] <- lapply(df_10[list_columns], function(x) sapply(x, toString))

df_10%<>%as.data.frame()

df_10$murdock_round <- paste0(df_10$murdock_names, df_10$round, sep="_")

df_10 %<>%
  group_by(murdock_names) %>%
  filter(n_distinct(iso) >= 2)%>%
  ungroup()

df_10 %<>%
  mutate(cluster_bdd = as.numeric(factor(paste(region_time, murdock_names, sep = "_"))))

saveRDS(df_10, file="data/data_rds/df_10.rds")

write_dta(df_10, "data/data_dta/df_10.dta")
```