### 2025-04-06 script/r/draft.Rmd
maj gitignore

```{r}
setwd("~/working_paper/rbci/")
getwd()
```

```{r packages, echo=FALSE, }
library(haven)
library(dplyr)
select <- dplyr::select
library(magrittr)
library(stringr)
library(geosphere)
library(purrr)
library(kableExtra)
library(tidyr)
library(readxl)
library(sf)
library(exactextractr)
library(raster)
library(rgdal)
library(ncdf4)
library(terra)
```

#Levels

```{r}
folder_path <- "data/data_raw/gadm"

files <- list.files(folder_path, full.names = TRUE)

levels <- data.frame(file = files) %>%
  mutate(country = str_extract(file, "(?<=_)[A-Z]{3}(?=_\\d)"),
         level = as.integer(str_extract(file, "(?<=_)[0-9]+(?=\\.shp$)"))) %>%
  arrange(country, desc(level)) %>%
  distinct(country, .keep_all = TRUE)

levels %<>%
  group_by(country) %>%
  summarise(levels_nb = first(level))

print(levels, n=40)

cty_adm1 <- levels%>%
  filter(levels_nb>=1)%>%
  pull(country)

cty_adm2 <- levels%>%
  filter(levels_nb>=2)%>%
  pull(country)

cty_adm3 <- levels%>%
  filter(levels_nb>=3)%>%
  pull(country)

cty_adm4 <- levels%>%
  filter(levels_nb>=4)%>%
  pull(country)
```

##Merge internet

```{r}
gg_stack <- readRDS("data/data_rds/gg_stack.rds")
ggg_stack <- readRDS("data/data_rds/ggg_stack.rds")

int_stack <- gg_stack %>%
  left_join(ggg_stack, by = c("iso", "NAME_1", "NAME_2", "NAME_3", "round"))

saveRDS(int_stack, file="data/data_rds/int_stack.rds")

int_stack <- readRDS("data/data_rds/int_stack.rds")

int_stack%<>%
  mutate(pop_adm1= coalesce(pop_adm1.x, pop_adm1.y))%>%
  mutate(pop_adm2 = coalesce(pop_adm2.x, pop_adm2.y))%>%
  mutate(pop_adm3 = coalesce(pop_adm3.x, pop_adm3.y))

int_stack%<>%
  select(-ends_with(".x"), -ends_with(".y"))

data_final <- readRDS("data/data_rds/data_final.rds")

data_final %<>%
  left_join(int_stack, by = c("iso", "NAME_1", "NAME_2", "NAME_3", "round"))

countries_adm1 <- c("GHA","KEN","MOZ", "MWI","NGA","TZA","UGA","ZMB", "ZWE")

countries_adm2 <- c("BEN", "BFA", "BWA", "CIV", "CMR", "GIN", "LBR", "MLI", "NAM", "NER", "SLE")

data_final%<>%
mutate(gg_region_pot = case_when(
    iso %in% countries_adm1 ~ gg_adm1_pot,
    iso %in%  countries_adm2 ~ gg_adm2_pot, 
    TRUE ~ NA_integer_
  ))%>%
mutate(ggg_region_pot = case_when(
    iso %in% countries_adm1 ~ ggg_adm1_pot,
    iso %in%  countries_adm2 ~ ggg_adm2_pot, 
    TRUE ~ NA_integer_))%>%
  mutate(cvg_region_pot=ifelse(is.na(ggg_region_pot), gg_region_pot, ggg_region_pot))

data_final%<>%
mutate(pop_region = case_when(
    iso %in% countries_adm1 ~ pop_adm1,
    iso %in% countries_adm2 ~ pop_adm2, 
    TRUE ~ NA_integer_
  ))%>%
  mutate(pop_region_log=log(pop_region+1))

data_final%<>%
  group_by(NAME_2) %>%
  mutate(name2_id = cur_group_id()) %>%
  ungroup() %>%
  group_by(NAME_1) %>%
  mutate(name1_id = cur_group_id())%>%
  ungroup()%>%
  mutate(adm1_time = case_when(
    iso %in% countries_adm1 ~ name1_id * 10000 + round,
    TRUE ~ NA_integer_
  ))%>%
mutate(adm1_time = as.integer(as.factor(adm1_time)))%>%
mutate(adm2_time = case_when(
    iso %in% countries_adm2 ~ name2_id * 10000 + round,
    TRUE ~ NA_integer_
  ))%>%
mutate(adm2_time = as.integer(as.factor(adm2_time)))

data_final%<>%
  mutate(country_round=as.numeric(paste0(round,ccode)))%>%
  mutate(region_time = case_when(
    iso %in% countries_adm1 ~ name1_id * 10000 + round, 
    iso %in% countries_adm2 ~ name2_id * 10000 + round,
    TRUE ~ NA_integer_
  ))%>%
  mutate(region_time = as.integer(as.factor(region_time)))

saveRDS(data_final, file="data/data_rds/data_final.rds")
write_dta(data_final, "data/data_dta/data_final.dta")
```

```{r}
library(ggplot2)
library(dplyr)
library(maps)

# Mapping of ISO codes to country names
iso_to_country <- data.frame(
  iso = c("AGO", "BDI", "BEN", "BFA", "BWA", "CIV", "CMR", "CPV", "DZA", "EGY", 
          "ETH", "GAB", "GHA", "GIN", "GMB", "KEN", "LBR", "LSO", "MAR", "MDG", 
          "MLI", "MOZ", "MUS", "MWI", "NAM", "NER", "NGA", "SDN", "SEN", "SLE", 
          "STP", "SWZ", "TGO", "TUN", "TZA", "UGA", "ZAF", "ZMB", "ZWE"),
  country = c("Angola", "Burundi", "Benin", "Burkina Faso", "Botswana", "Ivory Coast", 
              "Cameroon", "Cape Verde", "Algeria", "Egypt", "Ethiopia", "Gabon", 
              "Ghana", "Guinea", "Gambia", "Kenya", "Liberia", "Lesotho", "Morocco", 
              "Madagascar", "Mali", "Mozambique", "Mauritius", "Malawi", "Namibia", 
              "Niger", "Nigeria", "Sudan", "Senegal", "Sierra Leone", "Sao Tome and Principe", 
              "Eswatini", "Togo", "Tunisia", "Tanzania", "Uganda", "South Africa", 
              "Zambia", "Zimbabwe")
)

# Count observations for each country based on ISO codes
country_counts <- data_final_adm %>%
  group_by(iso) %>%
  summarize(observation_count = n())

# Filter countries with at least 1000 observations and match with ISO codes
highlight_countries <- country_counts %>%
  filter(observation_count >= 1000, iso %in% iso_to_country$iso) %>%
  pull(iso)

cty <- iso_to_country%>%
  pull(country)

# Get the world map data
africa_map <- map_data("world")%>%
  filter(region%in%cty)

# Add a column for highlighting based on country names
africa_map <- africa_map %>%
  left_join(iso_to_country, by = c("region" = "country")) %>%
  mutate(highlight = ifelse(iso %in% highlight_countries, "Yes", "No"))

# Plot
ggplot() +
  geom_polygon(data = africa_map, aes(x = long, y = lat, group = group, fill = highlight), 
               color = "black") +
  scale_fill_manual(values = c("Yes" = "lightblue", "No" = "gray"), guide = "none") +
  labs(title = "Map of Africa: Countries with At Least 1000 Observations Highlighted") +
  theme_minimal()

```

```{r}

afro_diff <- anti_join(afro_stack, data_final)

afro_diff%>%
  distinct(iso)

afro_diff %>%
  select(latitude, longitude)%>%
  View()

data_final%>%
  select(dist_cap_log, dist_ncap_log, dist_sndncap_log)%>%
  summary()

```

```{r}
gov_index <- read_excel("data/data_raw/gov_index.xlsx")%>%
  rename(iso=ISO)
polstab_index <- read_excel("data/data_raw/polstab_index.xlsx")%>%
  rename(iso=ISO)
reg_index <- read_excel("data/data_raw/reg_index.xlsx")%>%
  rename(iso=ISO)
rule_index <- read_excel("data/data_raw/rule_index.xlsx")%>%
  rename(iso=ISO)
voiceac_index <- read_excel("data/data_raw/voiceac_index.xlsx")%>%
  rename(iso=ISO)

vdem <- readRDS("data/data_raw/V-Dem-CY-Full+Others-v13.rds")

```

```{r}
data_final_adm%>% filter(iso=="BEN") %>% select(pop_adm1, pop_adm2, pop_adm3)%>%summary()

data_final_adm %>% select(gg_adm1_pot, round)%>%filter(round==6)%>%summary()

data_final_adm%>%filter(iso=="BEN")%>%select(iso, round, gg_adm1_pot, NAME_1)%>%View()
```

```{r}
ggg_stack <- readRDS("data/data_rds/ggg_stack.rds")

ggg_cty<-ggg_stack%>%distinct(iso)%>%pull(iso)

all_rounds <- seq(min(ggg_stack$round, na.rm = TRUE), max(ggg_stack$round, na.rm = TRUE), by = 1)

for (country in ggg_cty) {
  ggg_stack %>%
    filter(iso == country) %>%
    group_by(round) %>%
    summarize(
      ggg_adm1_mean = mean(ggg_adm1_pot, na.rm = TRUE),
      ggg_adm2_mean = mean(ggg_adm2_pot, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    complete(round = all_rounds, fill = list(ggg_adm1_mean = NA, ggg_adm2_mean = NA)) %>%  # Fill missing rounds with NA
    ggplot(aes(x = round)) +
    geom_line(aes(y = ggg_adm1_mean), size = 1, color = "blue") +
    geom_point(aes(y = ggg_adm1_mean), size = 2, color = "blue") +
    geom_line(aes(y = ggg_adm2_mean), size = 1, color = "red") +
    geom_point(aes(y = ggg_adm2_mean), size = 2, color = "red") +
    labs(
      title = paste("Evolution of ggg_adm1_pot and ggg_adm2_pot Across Rounds for", country),
      x = "Round",
      y = "Mean Values"
    ) +
    scale_y_continuous(limits = c(0, 1)) +  # Fixed y-axis scale between 0 and 1
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 16),
      axis.title = element_text(size = 14),
      legend.position = "none"
    ) -> p
  print(p)
}
```

```{r}
setwd("~/working_paper/trust/")
getwd()
```

```{r packages, echo=FALSE, }
library(haven)
library(dplyr)
select <- dplyr::select
library(magrittr)
library(sf)
```

```{r}
countries <- c("BEN", "BFA", "BWA", "CIV", "CMR", "GIN", "LBR", "MLI", "NAM", "NER", "SLE", "GHA", "KEN", "MOZ", "MWI", "NGA", "TZA", "UGA", "ZMB", "ZWE")

data_final <- readRDS("data/data_rds/data_final.rds")

data_final_sf <- subset(data_final, select = c("id_unique","iso","latitude","longitude"))

data_final_sf <- st_as_sf(data_final_sf, coords = c("longitude", "latitude"), crs = 4326)

result <- list()

for (country in countries) {
  shapefile <- file.path("C:/Users/Redha CHABA/Documents/working_paper/trust/data/data_raw/road_shp", paste0("routes_", country, ".shp"))

  if (file.exists(shapefile)) {
    roads_sf <- st_read(shapefile)
    roads_sf <- st_transform(roads_sf, st_crs(data_final_sf))

    data_final_sf$distance_to_road <- st_distance(data_final_sf, roads_sf) %>% apply(1, min)

    country_data_result <- as.data.frame(data_final_sf)
    country_data_result <- country_data_result %>%
      select(id_unique, distance_to_road)

    result[[country]] <- country_data_result
  } else {
    warning(paste("Shapefile not found for country:", country))
  }
}

# Combine results into a single data frame
result_combined <- do.call(rbind, lapply(names(result), function(country) {
  cbind(result[[country]], country = country)
}))

final_result_road <- do.call(rbind, result)

final_df6 <- merge(final_df6,final_result_road, by="id_unique")

remove(bdd, country_data, country_data_result, country_data_sf, result, roads_sf, countries, ISO, shapefile)

```

```{r}
tiff_file <- "data/data_raw/geography/desert/desert.tif"

desert_wgs84 <- rast(tiff_file)

print(desert_wgs84)

africa_extent <- ext(-30, 60, -40, 40)

desert_africa <- crop(desert_wgs84, africa_extent)

binary_raster <- ifel(!is.na(desert_africa), 1, 0)

print(binary_raster)

plot(binary_raster, main = "Desert Map", col = rev(heat.colors(100)))

desert_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(binary_raster), crs(country_shp))) {
    country_shp <- project(country_shp, crs(binary_raster))
  }

  binary_raster_cropped <- crop(binary_raster, country_shp)
  binary_raster_masked <- mask(binary_raster_cropped, country_shp)

  desert_adm1 <- exactextractr::exact_extract(binary_raster_masked, regions, 'mean')

  regions$desert_adm1 <- desert_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, desert_adm1)

  desert_adm1_df <- rbind(desert_adm1_df, country_results)
}

desert_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(desert_adm1_df))

saveRDS(desert_adm1_df, file="data/data_rds/geo_ctrl/desert_adm1.rds")

desert_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(binary_raster), crs(country_shp))) {
    country_shp <- project(country_shp, crs(binary_raster))
  }

  binary_raster_cropped <- crop(binary_raster, country_shp)
  binary_raster_masked <- mask(binary_raster_cropped, country_shp)

  desert_adm2 <- exactextractr::exact_extract(binary_raster_masked, regions, 'mean')

  regions$desert_adm2 <- desert_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, desert_adm2)

  desert_adm2_df <- rbind(desert_adm2_df, country_results)
}

desert_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(desert_adm2_df))

saveRDS(desert_adm2_df, file="data/data_rds/geo_ctrl/desert_adm2.rds")
```

```{r}
tiff_file <- "data/data_raw/geography/elevation/elevation.tif"

elevation_wgs84 <- rast(tiff_file)

print(elevation_wgs84)

plot(elevation_wgs84, main = "Elevation Map", col = terrain.colors(100), breaks = seq(-1000, 3000, by = 100))

elevation_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(elevation_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(elevation_wgs84))
  }

  elevation_wgs84_cropped <- crop(elevation_wgs84, country_shp)
  elevation_wgs84_masked <- mask(elevation_wgs84_cropped, country_shp)

  elevation_adm1 <- exactextractr::exact_extract(elevation_wgs84_masked, regions, 'mean')

  regions$elevation_adm1 <- elevation_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, elevation_adm1)

  elevation_adm1_df <- rbind(elevation_adm1_df, country_results)
}

elevation_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(elevation_adm1_df))

saveRDS(elevation_adm1_df, file="data/data_rds/geo_ctrl/elevation_adm1.rds")

elevation_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(elevation_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(elevation_wgs84))
  }

  elevation_wgs84_cropped <- crop(elevation_wgs84, country_shp)
  elevation_wgs84_masked <- mask(elevation_wgs84_cropped, country_shp)

  elevation_adm2 <- exactextractr::exact_extract(elevation_wgs84_masked, regions, 'mean')

  regions$elevation_adm2 <- elevation_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, elevation_adm2)

  elevation_adm2_df <- rbind(elevation_adm2_df, country_results)
}

elevation_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(elevation_adm2_df))

saveRDS(elevation_adm2_df, file="data/data_rds/geo_ctrl/elevation_adm2.rds")
```

```{r}
library(viridis)

tiff_file <- "data/data_raw/geography/precipitation/CHELSA_bio12_1981-2010_V.2.1.tif"

precipitation_wgs84 <- rast(tiff_file)

print(precipitation_wgs84)

africa_extent <- ext(-30, 60, -40, 40)

precipitation_africa <- crop(precipitation_wgs84, africa_extent)

plot(precipitation_africa, main = "Precipitation Map of Africa", col = rev(viridis(100)))

precipitation_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(precipitation_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(precipitation_wgs84))
  }

  precipitation_wgs84_cropped <- crop(precipitation_wgs84, country_shp)
  precipitation_wgs84_masked <- mask(precipitation_wgs84_cropped, country_shp)

  precipitation_adm1 <- exactextractr::exact_extract(precipitation_wgs84_masked, regions, 'mean')

  regions$precipitation_adm1 <- precipitation_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, precipitation_adm1)

  precipitation_adm1_df <- rbind(precipitation_adm1_df, country_results)
}

precipitation_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(precipitation_adm1_df))

saveRDS(precipitation_adm1_df, file="data/data_rds/geo_ctrl/precipitation_adm1.rds")

precipitation_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(precipitation_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(precipitation_wgs84))
  }

  precipitation_wgs84_cropped <- crop(precipitation_wgs84, country_shp)
  precipitation_wgs84_masked <- mask(precipitation_wgs84_cropped, country_shp)

  precipitation_adm2 <- exactextractr::exact_extract(precipitation_wgs84_masked, regions, 'mean')

  regions$precipitation_adm2 <- precipitation_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, precipitation_adm2)

  precipitation_adm2_df <- rbind(precipitation_adm2_df, country_results)
}

precipitation_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(precipitation_adm2_df))

saveRDS(precipitation_adm2_df, file="data/data_rds/geo_ctrl/precipitation_adm2.rds")
```

```{r}
shapefile_directory <- "data/data_raw/gadm/"

shapefile_names <- list.files(shapefile_directory, pattern = "*_2.shp", full.names = TRUE)

adm2_shapefiles <- lapply(shapefile_names, st_read)
adm2_combined <- do.call(rbind, adm2_shapefiles)

adm2_combined_merged <- merge(adm2_combined, precipitation_adm2, by.x = "NAME_2", by.y = "NAME_2", all.x = TRUE)

library(ggplot2)

ggplot() +
  geom_sf(data = adm2_combined_merged, aes(fill = precipitation_adm2), color = "black", size = 0.25) +
  scale_fill_viridis(name = "Precipitation (mm)", option = "plasma", na.value = "grey") +
  labs(title = "Precipitation Map of Africa",
       subtitle = "Regional Precipitation Values") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    legend.position = "bottom"
  )

```

```{r}
tiff_file <- "data/data_raw/geography/temperature/CHELSA_bio1_1981-2010_V.2.1.tif"

temp_mean_wgs84 <- rast(tiff_file)

print(temp_mean_wgs84)

africa_extent <- ext(-30, 60, -40, 40)

temp_mean_africa <- crop(temp_mean_wgs84, africa_extent)

plot(temp_mean_africa, main = "temp_mean Map of Africa", col = rev(heat.colors(100)))

temp_mean_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(temp_mean_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_mean_wgs84))
  }

  temp_mean_wgs84_cropped <- crop(temp_mean_wgs84, country_shp)
  temp_mean_wgs84_masked <- mask(temp_mean_wgs84_cropped, country_shp)

  temp_mean_adm1 <- exactextractr::exact_extract(temp_mean_wgs84_masked, regions, 'mean')

  regions$temp_mean_adm1 <- temp_mean_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, temp_mean_adm1)

  temp_mean_adm1_df <- rbind(temp_mean_adm1_df, country_results)
}

temp_mean_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_mean_adm1_df))

saveRDS(temp_mean_adm1_df, file="data/data_rds/geo_ctrl/temp_mean_adm1.rds")

temp_mean_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(temp_mean_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_mean_wgs84))
  }

  temp_mean_wgs84_cropped <- crop(temp_mean_wgs84, country_shp)
  temp_mean_wgs84_masked <- mask(temp_mean_wgs84_cropped, country_shp)

  temp_mean_adm2 <- exactextractr::exact_extract(temp_mean_wgs84_masked, regions, 'mean')

  regions$temp_mean_adm2 <- temp_mean_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, temp_mean_adm2)

  temp_mean_adm2_df <- rbind(temp_mean_adm2_df, country_results)
}

temp_mean_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_mean_adm2_df))

saveRDS(temp_mean_adm2_df, file="data/data_rds/geo_ctrl/temp_mean_adm2.rds")
```

```{r}
tiff_file <- "data/data_raw/geography/temperature/CHELSA_bio5_1981-2010_V.2.1.tif"

temp_max_wgs84 <- rast(tiff_file)

print(temp_max_wgs84)

africa_extent <- ext(-30, 60, -40, 40)

temp_max_africa <- crop(temp_max_wgs84, africa_extent)

plot(temp_max_africa, main = "temp_max Map of Africa", col = rev(heat.colors(100)), breaks = seq(0, 50, by = 1))

temp_max_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(temp_max_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_max_wgs84))
  }

  temp_max_wgs84_cropped <- crop(temp_max_wgs84, country_shp)
  temp_max_wgs84_masked <- mask(temp_max_wgs84_cropped, country_shp)

  temp_max_adm1 <- exactextractr::exact_extract(temp_max_wgs84_masked, regions, 'mean')

  regions$temp_max_adm1 <- temp_max_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, temp_max_adm1)

  temp_max_adm1_df <- rbind(temp_max_adm1_df, country_results)
}

temp_max_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_max_adm1_df))

saveRDS(temp_max_adm1_df, file="data/data_rds/geo_ctrl/temp_max_adm1.rds")

temp_max_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(temp_max_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_max_wgs84))
  }

  temp_max_wgs84_cropped <- crop(temp_max_wgs84, country_shp)
  temp_max_wgs84_masked <- mask(temp_max_wgs84_cropped, country_shp)

  temp_max_adm2 <- exactextractr::exact_extract(temp_max_wgs84_masked, regions, 'mean')

  regions$temp_max_adm2 <- temp_max_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, temp_max_adm2)

  temp_max_adm2_df <- rbind(temp_max_adm2_df, country_results)
}

temp_max_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_max_adm2_df))

saveRDS(temp_max_adm2_df, file="data/data_rds/geo_ctrl/temp_max_adm2.rds")
```

```{r}
tiff_file <- "data/data_raw/geography/temperature/CHELSA_bio6_1981-2010_V.2.1.tif"

temp_min_wgs84 <- rast(tiff_file)

print(temp_min_wgs84)

africa_extent <- ext(-30, 60, -40, 40)

temp_min_africa <- crop(temp_min_wgs84, africa_extent)

plot(temp_min_africa, main = "temp_min Map of Africa", col = viridis(100), breaks = seq(-20, 30, by = 1))

temp_min_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(temp_min_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_min_wgs84))
  }

  temp_min_wgs84_cropped <- crop(temp_min_wgs84, country_shp)
  temp_min_wgs84_masked <- mask(temp_min_wgs84_cropped, country_shp)

  temp_min_adm1 <- exactextractr::exact_extract(temp_min_wgs84_masked, regions, 'mean')

  regions$temp_min_adm1 <- temp_min_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, temp_min_adm1)

  temp_min_adm1_df <- rbind(temp_min_adm1_df, country_results)
}

temp_min_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_min_adm1_df))

saveRDS(temp_min_adm1_df, file="data/data_rds/geo_ctrl/temp_min_adm1.rds")

temp_min_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(temp_min_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_min_wgs84))
  }

  temp_min_wgs84_cropped <- crop(temp_min_wgs84, country_shp)
  temp_min_wgs84_masked <- mask(temp_min_wgs84_cropped, country_shp)

  temp_min_adm2 <- exactextractr::exact_extract(temp_min_wgs84_masked, regions, 'mean')

  regions$temp_min_adm2 <- temp_min_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, temp_min_adm2)

  temp_min_adm2_df <- rbind(temp_min_adm2_df, country_results)
}

temp_min_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_min_adm2_df))

saveRDS(temp_min_adm2_df, file="data/data_rds/geo_ctrl/temp_min_adm2.rds")
```

```{r}
tiff_file <- "data/data_raw/geography/mountain/k3binary.tif"

mountain_wgs84 <- rast(tiff_file)

print(mountain_wgs84)

africa_extent <- ext(-30, 60, -40, 40)

mountain_africa <- crop(mountain_wgs84, africa_extent)

binary_raster <- ifel(!is.na(mountain_africa), 1, 0)

plot(binary_raster, main = "mountain Map of Africa", col = viridis(100))

mountain_adm1_df <- data.frame()

for (country in cty_adm1) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_1.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(binary_raster), crs(country_shp))) {
    country_shp <- project(country_shp, crs(binary_raster))
  }

  binary_raster_cropped <- crop(binary_raster, country_shp)
  binary_raster_masked <- mask(binary_raster_cropped, country_shp)

  mountain_adm1 <- exactextractr::exact_extract(binary_raster_masked, regions, 'mean')

  regions$mountain_adm1 <- mountain_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, mountain_adm1)

  mountain_adm1_df <- rbind(mountain_adm1_df, country_results)
}

mountain_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(mountain_adm1_df))

saveRDS(mountain_adm1_df, file="data/data_rds/geo_ctrl/mountain_adm1.rds")

mountain_adm2_df <- data.frame()

for (country in cty_adm2) {

  shapefile <- paste0("data/data_raw/gadm/gadm41_", country, "_2.shp")
  regions <- st_read(shapefile) %>% st_make_valid()

  shapefile_path <- paste0("data/data_raw/gadm/gadm41_", country, "_0.shp")
  country_shp <- vect(shapefile_path)

  if (!identical(crs(binary_raster), crs(country_shp))) {
    country_shp <- project(country_shp, crs(binary_raster))
  }

  binary_raster_cropped <- crop(binary_raster, country_shp)
  binary_raster_masked <- mask(binary_raster_cropped, country_shp)

  mountain_adm2 <- exactextractr::exact_extract(binary_raster_masked, regions, 'mean')

  regions$mountain_adm2 <- mountain_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, NAME_2, mountain_adm2)

  mountain_adm2_df <- rbind(mountain_adm2_df, country_results)
}

mountain_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(mountain_adm2_df))

saveRDS(mountain_adm2_df, file="data/data_rds/geo_ctrl/mountain_adm2.rds")
```

```{r}
desert_adm1 <- readRDS("data/data_rds/geo_ctrl/desert_adm1.rds")
desert_adm2 <- readRDS("data/data_rds/geo_ctrl/desert_adm2.rds")

desert <- merge(desert_adm1, desert_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(desert_adm1, desert_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(desert)))

expanded_data <- desert[rep(1:nrow(desert), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo <- expanded_data

rm(desert)

elevation_adm1 <- readRDS("data/data_rds/geo_ctrl/elevation_adm1.rds")
elevation_adm2 <- readRDS("data/data_rds/geo_ctrl/elevation_adm2.rds")

elevation <- merge(elevation_adm1, elevation_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(elevation_adm1, elevation_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(elevation)))

expanded_data <- elevation[rep(1:nrow(elevation), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo %<>% left_join(expanded_data, by=c("iso", "NAME_1", "NAME_2", "round"))

rm(elevation)

mountain_adm1 <- readRDS("data/data_rds/geo_ctrl/mountain_adm1.rds")
mountain_adm2 <- readRDS("data/data_rds/geo_ctrl/mountain_adm2.rds")

mountain <- merge(mountain_adm1, mountain_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(mountain_adm1, mountain_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(mountain)))

expanded_data <- mountain[rep(1:nrow(mountain), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo %<>% left_join(expanded_data, by=c("iso", "NAME_1", "NAME_2", "round"))

rm(mountain)

precipitation_adm1 <- readRDS("data/data_rds/geo_ctrl/precipitation_adm1.rds")
precipitation_adm2 <- readRDS("data/data_rds/geo_ctrl/precipitation_adm2.rds")

precipitation <- merge(precipitation_adm1, precipitation_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(precipitation_adm1, precipitation_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(precipitation)))

expanded_data <- precipitation[rep(1:nrow(precipitation), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo %<>% left_join(expanded_data, by=c("iso", "NAME_1", "NAME_2", "round"))

rm(precipitation)

temp_max_adm1 <- readRDS("data/data_rds/geo_ctrl/temp_max_adm1.rds")
temp_max_adm2 <- readRDS("data/data_rds/geo_ctrl/temp_max_adm2.rds")

temp_max <- merge(temp_max_adm1, temp_max_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(temp_max_adm1, temp_max_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(temp_max)))

expanded_data <- temp_max[rep(1:nrow(temp_max), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo %<>% left_join(expanded_data, by=c("iso", "NAME_1", "NAME_2", "round"))

rm(temp_max)

temp_min_adm1 <- readRDS("data/data_rds/geo_ctrl/temp_min_adm1.rds")
temp_min_adm2 <- readRDS("data/data_rds/geo_ctrl/temp_min_adm2.rds")

temp_min <- merge(temp_min_adm1, temp_min_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(temp_min_adm1, temp_min_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(temp_min)))

expanded_data <- temp_min[rep(1:nrow(temp_min), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo %<>% left_join(expanded_data, by=c("iso", "NAME_1", "NAME_2", "round"))

rm(temp_min)

temp_mean_adm1 <- readRDS("data/data_rds/geo_ctrl/temp_mean_adm1.rds")
temp_mean_adm2 <- readRDS("data/data_rds/geo_ctrl/temp_mean_adm2.rds")

temp_mean <- merge(temp_mean_adm1, temp_mean_adm2, by = c("iso", "NAME_1"), all = TRUE)

rm(temp_mean_adm1, temp_mean_adm2)

rounds <- data.frame(round = rep(5:8, each = nrow(temp_mean)))

expanded_data <- temp_mean[rep(1:nrow(temp_mean), times = 4), ]

expanded_data$round <- rounds$round

expanded_data <- expanded_data %>% distinct(iso, NAME_1, NAME_2, round, .keep_all = TRUE)

data_geo %<>% left_join(expanded_data, by=c("iso", "NAME_1", "NAME_2", "round"))

rm(temp_mean)

data_final_r <- readRDS("data/data_rds/data_final_r.rds")

data_final_r %<>% left_join(data_geo, by=c("iso", "NAME_1", "NAME_2", "round"))

countries_adm1 <- c("GHA","KEN","MOZ", "MWI","NGA","TZA","UGA","ZMB", "ZWE")
countries_adm2 <- c("BEN", "BFA", "BWA", "CIV", "CMR", "GIN", "LBR", "MLI", "NAM", "NER", "SLE")

data_final_r %<>%
mutate(desert_region = case_when(
    iso %in% countries_adm1 ~ desert_adm1,
    iso %in% countries_adm2 ~ desert_adm2, 
    TRUE ~ NA_integer_
  ))%>%
mutate(elevation_region = case_when(
    iso %in% countries_adm1 ~ elevation_adm1,
    iso %in% countries_adm2 ~ elevation_adm2, 
    TRUE ~ NA_integer_
  ))%>%
mutate(mountain_region = case_when(
    iso %in% countries_adm1 ~ mountain_adm1,
    iso %in% countries_adm2 ~ mountain_adm2, 
    TRUE ~ NA_integer_
  ))%>%
mutate(precipitation_region = case_when(
    iso %in% countries_adm1 ~ precipitation_adm1,
    iso %in% countries_adm2 ~ precipitation_adm2, 
    TRUE ~ NA_integer_
  ))%>%
mutate(temp_max_region = case_when(
    iso %in% countries_adm1 ~ temp_max_adm1,
    iso %in% countries_adm2 ~ temp_max_adm2, 
    TRUE ~ NA_integer_
  ))%>%
mutate(temp_min_region = case_when(
    iso %in% countries_adm1 ~ temp_min_adm1,
    iso %in% countries_adm2 ~ temp_min_adm2, 
    TRUE ~ NA_integer_
  ))%>%
mutate(temp_mean_region = case_when(
    iso %in% countries_adm1 ~ temp_mean_adm1,
    iso %in% countries_adm2 ~ temp_mean_adm2, 
    TRUE ~ NA_integer_
  ))

data_final <- readRDS("data/data_rds/data_27_02_25.rds")

data_final%<>%
  ungroup()

data_final%<>%
  mutate(desert_region_tv=desert_region*tv,
         mountain_region_tv=mountain_region*tv,
         elevation_region_tv=elevation_region*tv)

saveRDS(data_final, file="data/data_rds/data_final.rds")

write_dta(data_final, "data/data_dta/data_final.dta")
```

```{r}
vdem <- readRDS("data/data_raw/V-Dem-CY-Full+Others-v13.rds")

vdem%<>%
  select(country_text_id, year, v2xedvd_inpt, v2x_polyarchy, v2xedvd_me_cent, v2medstateprint, v2medstatebroad, v2medpolstate, v2medpolnonstate, e_v2x_freexp_altinf_3C, e_v2x_freexp_altinf_4C, e_v2x_freexp_altinf_5C, v2x_freexp, v2xme_altinf, e_polity2)

vdem%<>%
  filter(country_text_id %in% countries)

vdem%<>%
  rename(iso=country_text_id)%>%
    mutate(round=case_when(between(year, 2011, 2013)~5,
                         between(year, 2014, 2015)~6,
                         between(year, 2016, 2018)~7,
                         between(year, 2019, 2021)~8,
         TRUE~NA_integer_))

vdem %<>%
  group_by(iso, round) %>%
  filter(!is.na(round)) %>%
  mutate(
    v2xedvd_inpt_mean2 = ifelse(all(is.na(v2xedvd_inpt)), NA_integer_, mean(v2xedvd_inpt, na.rm = TRUE)),
    v2x_polyarchy_mean2 = ifelse(all(is.na(v2x_polyarchy)), NA_integer_, mean(v2x_polyarchy, na.rm = TRUE)),
    v2xedvd_me_cent_mean2 = ifelse(all(is.na(v2xedvd_me_cent)), NA_integer_, mean(v2xedvd_me_cent, na.rm = TRUE)),
    v2medstateprint_mean2 = ifelse(all(is.na(v2medstateprint)), NA_integer_, mean(v2medstateprint, na.rm = TRUE)),
    v2medstatebroad_mean2 = ifelse(all(is.na(v2medstatebroad)), NA_integer_, mean(v2medstatebroad, na.rm = TRUE)),
    v2medpolstate_mean2 = ifelse(all(is.na(v2medpolstate)), NA_integer_, mean(v2medpolstate, na.rm = TRUE)),
    v2medpolnonstate_mean2 = ifelse(all(is.na(v2medpolnonstate)), NA_integer_, mean(v2medpolnonstate, na.rm = TRUE)),
    e_v2x_freexp_altinf_3C_mean2 = ifelse(all(is.na(e_v2x_freexp_altinf_3C)), NA_integer_, mean(e_v2x_freexp_altinf_3C, na.rm = TRUE)),
    e_v2x_freexp_altinf_4C_mean2 = ifelse(all(is.na(e_v2x_freexp_altinf_4C)), NA_integer_, mean(e_v2x_freexp_altinf_4C, na.rm = TRUE)),
    e_v2x_freexp_altinf_5C_mean2 = ifelse(all(is.na(e_v2x_freexp_altinf_5C)), NA_integer_, mean(e_v2x_freexp_altinf_5C, na.rm = TRUE)),
    v2x_freexp_mean2 = ifelse(all(is.na(v2x_freexp)), NA_integer_, mean(v2x_freexp, na.rm = TRUE)),
    v2xme_altinf_mean2 = ifelse(all(is.na(v2xme_altinf)), NA_integer_, mean(v2xme_altinf, na.rm = TRUE)),
    e_polity2_mean2 = ifelse(all(is.na(e_polity2)), NA_integer_, mean(e_polity2, na.rm = TRUE))
  ) %>%
  ungroup()

vdem%<>%
  mutate(e_polity2_med_all=ifelse(is.na(round), NA_integer_, median(e_polity2, na.rm=TRUE)))

vdem%<>%
  select(iso, round, v2xedvd_inpt_mean2, v2x_polyarchy_mean2, v2xedvd_me_cent_mean2, v2medstateprint_mean2, v2medstatebroad_mean2, v2medpolstate_mean2, v2medpolnonstate_mean2, e_v2x_freexp_altinf_3C_mean2, e_v2x_freexp_altinf_4C_mean2, e_v2x_freexp_altinf_5C_mean2, v2x_freexp_mean2, v2xme_altinf_mean2, e_polity2_mean2, e_polity2_med_all)%>%
    distinct(iso, round, .keep_all = TRUE)

data_final %<>% left_join(vdem, by=c("iso", "round"))

```

```{r}
library(readxl)
library(tidyr)

data_final <- readRDS("data/data_rds/data_final.rds")

countries <- data_final %>% pull(iso) %>% unique()

wb_gov <- read_excel("data/data_raw/wb_gov/wgidataset.xlsx")

wb_gov %<>%
  select(code, year, indicator, estimate) %>%
  mutate(estimate = na_if(estimate, ".."))

wb_gov %<>%
  pivot_wider(names_from = indicator, values_from = estimate)

wb_gov %<>%
  rename(
    ctrl_corr = cc,
    voice_acc = va,
    pol_stab = pv,
    gov_eff = ge,
    reg_qty = rq,
    rol = rl,
    iso = code
  ) %>%
  mutate(across(c(
    ctrl_corr, voice_acc, pol_stab, gov_eff, reg_qty, rol
  ), as.numeric))

wb_gov %<>%
  mutate(round = case_when(
    between(year, 2011, 2013) ~ 5,
    between(year, 2014, 2015) ~ 6,
    between(year, 2016, 2018) ~ 7,
    between(year, 2019, 2021) ~ 8,
    TRUE ~ NA_integer_
  ))

wb_gov %<>%
  group_by(iso, round) %>%
  filter(!is.na(round)) %>%
  mutate(
    ctrl_corr_mean = ifelse(all(is.na(ctrl_corr)), NA_integer_, mean(ctrl_corr, na.rm = TRUE)),
    voice_acc_mean = ifelse(all(is.na(voice_acc)), NA_integer_, mean(voice_acc, na.rm = TRUE)),
    pol_stab_mean = ifelse(all(is.na(pol_stab)), NA_integer_, mean(pol_stab, na.rm = TRUE)),
    gov_eff_mean = ifelse(all(is.na(gov_eff)), NA_integer_, mean(gov_eff, na.rm = TRUE)),
    reg_qty_mean = ifelse(all(is.na(reg_qty)), NA_integer_, mean(reg_qty, na.rm = TRUE)),
    rol_mean = ifelse(all(is.na(rol)), NA_integer_, mean(rol, na.rm = TRUE)),
  ) %>%
  ungroup()

wb_gov%<>%
  mutate(gov_cap = rowMeans(select(., gov_eff_mean, reg_qty_mean), na.rm = TRUE))

wb_gov %<>%
  select(
    iso,
    round,
    ctrl_corr_mean,
    voice_acc_mean,
    pol_stab_mean,
    gov_eff_mean,
    reg_qty_mean,
    rol_mean,
    gov_cap
  ) %>%
  distinct(iso, round, .keep_all = TRUE)

wb_gov %<>%
  filter(iso %in% countries)

data_final %<>% left_join(wb_gov, by = c("iso", "round"))

saveRDS(data_final, file = "data/data_rds/data_final.rds")

write_dta(data_final, "data/data_dta/data_final.dta")

```

```{r}
# Load dplyr package
library(dplyr)

# Assuming your data is in data_final and has columns ctrl_corr_mean and spl
percentile_below_neg_0_5 <- data_final %>%
  filter(spl == 1) %>%  # Filter for spl == 1
  summarise(
    total_count = n(),  # Total number of observations where spl == 1
    below_neg_0_5_count = sum(ctrl_corr_mean < -0.5)  # Count values less than -0.5
  ) %>%
  mutate(
    percentile = (below_neg_0_5_count / total_count) * 100  # Calculate the percentage
  )

# Display the result
print(percentile_below_neg_0_5)

```

```{r}
df <- readRDS("data/data_rds/data_final.rds")

df %<>%
  filter(spl==1, round==5)

radius<- 5000

df_sf_buff <- df[, .(longitude, latitude, iso, id_ea)] %>%
  vect(geom = c("longitude", "latitude"), crs = "EPSG:4326") %>%
  buffer(width = radius)

if (any(!is.valid(df_sf_buff))) {
  df_sf_buff <- make_valid(df_sf_buff)
}

cty_shp <- vect("data/data_raw/gadm/gadm41_BEN_2.shp")

cty_buff <- intersect(df_sf_buff, cty_shp)

cty_sf <- st_as_sf(cty_shp)
cty_buff_sf <- st_as_sf(cty_buff)

ggplot() +
  geom_sf(
    data = cty_sf,
    fill = "lightgray",
    color = "black",
    alpha = 0.5
  ) +

  geom_sf(
    data = cty_buff_sf,
    fill = "gold",
    color = "gold4",
    alpha = 0.3
  ) +

  theme_minimal() +
  labs(title = "Circles 10km radius")

```

```{r}
rc <- function(x) {
  ifelse(is.na(x), 0 , 1)
}

# Define TIFF file path
tiff_file <- "data/data_raw/collins_recode/Mobile Coverage Explorer v2012 - GeoTIFF/Data_MCE/ByCountry/MCE_2G/MCE_BEN2G_2012.tif"

# Filter circles based on ISO code
circles <- df_sf_buff[df_sf_buff$iso == "BEN"]

# Harmonize CRS to WGS84 using terra::project()
target_crs <- "EPSG:4326"  # WGS84 CRS

# Load and process raster using terra (faster than raster)
gg <- rast(tiff_file)
plot(gg)
gg_wgs84 <- project(gg, target_crs)
plot(gg_wgs84)
gg_wgs84_recode <- app(gg_wgs84, fun = rc)
plot(gg_wgs84_recode)

# Check the extent of circles to ensure it's valid
circles_extent <- ext(circles)

plot(circles)
plot(circles_extent)
# Crop and mask raster to the buffered areas
gg_wgs84_cropped <- crop(gg_wgs84_recode, circles_extent)
gg_wgs84_masked <- mask(gg_wgs84_cropped, circles)
plot(gg_wgs84_cropped)
plot(gg_wgs84_masked)

global(gg_wgs84_masked, mean, na.rm = TRUE)

circles_cropped <- crop(circles, gg_wgs84_masked)

plot(circles_cropped, add = TRUE, border = "red")

# Extract values from the raster based on the polygons
gg_values <- terra::extract(gg_wgs84_masked,
                            circles_cropped,
                            fun = mean,
                            na.rm = TRUE)

circles_cropped$gg_mean <- gg_values[, 2]
```

```{r}
# 1. Load region shapefile (convert to sf for use with dplyr/ggplot2)
adm2_shp <- vect("data/data_raw/gadm/gadm41_BEN_2.shp")
adm2_sf <- st_as_sf(adm2_shp)
adm2_sf <- st_make_valid(adm2_sf)

df <- readRDS("data/data_rds/data_final.rds")

df_round5 <- df %>% filter(round == 5)
df_round8 <- df %>% filter(round == 8)

# Get the unique region identifiers in both rounds (intersection)
common_regions <- intersect(df_round5$NAME_2, df_round8$NAME_2)

# Filter original data to keep only the regions that appear in both rounds
df_common <- df %>% filter(iso=="BEN", NAME_2 %in% common_regions) %>% select(-NAME_2)

df_common%<>%filter(round==5)

# Convert to sf object (if not already)
df_sf <- st_as_sf(df_common, coords = c("longitude", "latitude"), crs = st_crs(adm2_sf))

# Proceed with spatial join or further analysis
df_adm2_sf <- st_join(df_sf, adm2_sf)

# 5. Aggregate: calculate mean trust by region
trust_mean <- df_adm2_sf %>%
  group_by(NAME_2, round) %>%  # replace with actual region ID column
  summarise(mean_trust = mean(trust_parl_a, na.rm = TRUE))

trust_mean_df <- st_drop_geometry(trust_mean)

# 6. Join aggregated data back to region geometries
regions_map <- adm2_sf %>%
  left_join(trust_mean_df, by = "NAME_2")  # make sure column name matches

capitals <- df_common %>%
  distinct(caplong, caplat, iso)%>%
  na.omit() %>%
  rename(longitude = caplong, latitude = caplat)

# 7. Plot the map
ggplot(regions_map) +
  geom_sf(aes(fill = mean_trust), color = "white") +
  scale_fill_gradient(
    low = "red",   # Low values will be red
    high = "green", # High values will be green
    na.value = "lightgrey",  # Missing values will be light grey
    limits = c(0, 3)  # Fix scale between 0 and 3

  ) +
  theme_minimal() +
  labs(
    title = "Trust - Round 5",
    fill = "Trust"
  )+
   geom_point(data = capitals, 
             aes(x = longitude, y = latitude), 
             size = 2, shape = 22, fill = "black", show.legend = FALSE)

# 1. Load region shapefile (convert to sf for use with dplyr/ggplot2)
adm2_shp <- vect("data/data_raw/gadm/gadm41_BEN_2.shp")
adm2_sf <- st_as_sf(adm2_shp)
adm2_sf <- st_make_valid(adm2_sf)

df <- readRDS("data/data_rds/data_final.rds")

df_round5 <- df %>% filter(round == 5)
df_round8 <- df %>% filter(round == 8)

# Get the unique region identifiers in both rounds (intersection)
common_regions <- intersect(df_round5$NAME_2, df_round8$NAME_2)

# Filter original data to keep only the regions that appear in both rounds
df_common <- df %>% filter(iso=="BEN", NAME_2 %in% common_regions) %>% select(-NAME_2)

df_common%<>%filter(round==8)

# Convert to sf object (if not already)
df_sf <- st_as_sf(df_common, coords = c("longitude", "latitude"), crs = st_crs(adm2_sf))

# Proceed with spatial join or further analysis
df_adm2_sf <- st_join(df_sf, adm2_sf)

# 5. Aggregate: calculate mean trust by region
trust_mean <- df_adm2_sf %>%
  group_by(NAME_2, round) %>%  # replace with actual region ID column
  summarise(mean_trust = mean(trust_parl_a, na.rm = TRUE))

trust_mean_df <- st_drop_geometry(trust_mean)

# 6. Join aggregated data back to region geometries
regions_map <- adm2_sf %>%
  left_join(trust_mean_df, by = "NAME_2")  # make sure column name matches

capitals <- df_common %>%
  distinct(caplong, caplat, iso)%>%
  na.omit() %>%
  rename(longitude = caplong, latitude = caplat)

# 7. Plot the map
ggplot(regions_map) +
  geom_sf(aes(fill = mean_trust), color = "white") +
  scale_fill_gradient(
    low = "red",   # Low values will be red
    high = "green", # High values will be green
    na.value = "lightgrey",  # Missing values will be light grey
    limits = c(0, 3)  # Fix scale between 0 and 3

  ) +
  theme_minimal() +
  labs(
    title = "Trust - Round 8",
    fill = "Trust"
  )+
   geom_point(data = capitals, 
             aes(x = longitude, y = latitude), 
             size = 2, shape = 22, fill = "black", show.legend = FALSE)

# 1. Load region shapefile (convert to sf for use with dplyr/ggplot2)
adm2_shp <- vect("data/data_raw/gadm/gadm41_BEN_2.shp")
adm2_sf <- st_as_sf(adm2_shp)
adm2_sf <- st_make_valid(adm2_sf)

df <- readRDS("data/data_rds/data_final.rds")

df_round5 <- df %>% filter(round == 5)
df_round8 <- df %>% filter(round == 8)

# Get the unique region identifiers in both rounds (intersection)
common_regions <- intersect(df_round5$NAME_2, df_round8$NAME_2)

# Filter original data to keep only the regions that appear in both rounds
df_common <- df %>% filter(iso=="BEN", NAME_2 %in% common_regions)

# 3. Aggregate: calculate mean trust by region and round (for both rounds 5 and 8)
trust_mean_by_round <- df_common %>%
  group_by(NAME_2, round) %>%
  summarise(mean_trust = mean(trust_parl_a, na.rm = TRUE), .groups = "drop")

# 4. Spread data: Create a wide format to compare rounds 5 and 8
trust_mean_wide <- trust_mean_by_round %>%
  pivot_wider(names_from = round, values_from = mean_trust) %>%
  rename(mean_trust_5 = `5`, mean_trust_8 = `8`)

# 5. Calculate the variation in trust (difference between round 8 and round 5)
trust_mean_wide <- trust_mean_wide %>%
  mutate(variation = mean_trust_8 - mean_trust_5)

# 6. Join the aggregated data back to the region geometries
regions_map <- adm2_sf %>%
  left_join(trust_mean_wide, by = "NAME_2")

# 7. Plot the variation map
ggplot(regions_map) +
  geom_sf(aes(fill = variation), color = "white") +
  scale_fill_gradient(
    low = "red",   # Low values will be red (indicating decrease in trust)
    high = "green", # High values will be green (indicating increase in trust)
    na.value = "lightgrey",  # Missing values will be light grey
    limits = c(-3, 3)  # Fix scale between -3 and 3 for variation
  ) +
  theme_minimal() +
  labs(
    title = "Trust (Round 8 - Round 5)",
    fill = "Delta"
  ) +
   geom_point(data = capitals, 
             aes(x = longitude, y = latitude), 
             size = 2, shape = 22, fill = "black", show.legend = FALSE)

```

```{r}
library(terra)
library(dplyr)
library(magrittr)
library(ggplot2)
library(tidyr)
library(gt)
library(rnaturalearth)

# Load survey data
df <- readRDS("data/data_rds/data_final.rds")

df%<>%
  filter(spl==1)

# Unique ISO codes in data
country_codes <- unique(df$iso)

# Path to admin2 shapefiles
shapefile_path <- "data/data_raw/gadm/"

# Store each country’s region vector
all_regions_vect <- list()

for (iso in country_codes) {
  shp_file <- paste0(shapefile_path, "gadm41_", iso, "_2.shp")

  if (!file.exists(shp_file)) {
    message("⛔ Skipping ", iso, " - shapefile missing.")
    next
  }

  message("✅ Processing ", iso)

  # Load shapefile
  adm2_vect <- vect(shp_file)

  # Filter country data and only include rounds 5 and 8
  df_country <- df %>% filter(iso == iso, round %in% c(5, 8))

  if (nrow(df_country) == 0) {
    message("⚠️ No data for rounds 5 or 8 for ", iso)
    adm2_vect$variation <- NA
    all_regions_vect[[iso]] <- adm2_vect
    next
  }

  df5 <- df_country %>% filter(round == 5)
  df8 <- df_country %>% filter(round == 8)
  common_regions <- intersect(df5$NAME_2, df8$NAME_2)

  if (length(common_regions) == 0) {
    message("⚠️ No common regions for ", iso)
    adm2_vect$variation <- NA
    all_regions_vect[[iso]] <- adm2_vect
    next
  }

  df_common <- df_country %>% filter(NAME_2 %in% common_regions)

  # Aggregate and calculate mean trust values
  trust_wide <- df_common %>%
    group_by(NAME_2, round) %>%
    summarise(trust = mean(trust_parl_a, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = round, values_from = trust, names_prefix = "R") %>%
    mutate(variation = R8 - R5)

  # Merge trust data back with spatial data
  adm2_vect$NAME_2 <- as.character(adm2_vect$NAME_2)
  joined <- tryCatch({
    merge(adm2_vect, trust_wide, by = "NAME_2", all.x = TRUE)
  }, error = function(e) {
    message("❌ Merge failed for ", iso, ": ", e$message)
    NULL
  })

  all_regions_vect[[iso]] <- joined
}

combined <- Reduce(rbind, all_regions_vect)
combined_df <- as.data.frame(combined)

capitals <- df %>%
  filter(round %in% c(5, 8)) %>%
  distinct(caplong, caplat, iso) %>%
  na.omit() %>%
  rename(longitude = caplong, latitude = caplat)

africa_outline <- ne_countries(continent = "Africa", returnclass = "sf") %>%
  st_make_valid()

africa_sf <- st_as_sf(combined)

ggplot() +
  geom_sf(data = africa_sf, aes(fill = variation), color = "white", linewidth = 0.1) +
  geom_sf(data = africa_outline, fill = NA, color = "black", linewidth = 0.2) +
  geom_point(data = capitals, aes(x = longitude, y = latitude),
             shape = 21, size = 2, fill = "black") +
  scale_fill_gradient2(
    low = "red", mid = "white", high = "green",
    midpoint = 0, limits = c(-3, 3),
    na.value = "lightgrey",
    name = "Δ Trust\n(R8 - R5)"
  ) +
  theme_minimal() +
  labs(
    title = "Trust Change in Parliament (Round 8 - Round 5)",
    subtitle = "Across African Regions",
    caption = "Grey = No data"
  )
```

```{r}
# Define which countries use which admin level
countries_adm1 <- c("GHA","KEN","MOZ", "MWI","NGA","TZA","UGA","ZMB", "ZWE")
countries_adm2 <- c("BEN", "BFA", "BWA", "CIV", "CMR", "GIN", "LBR", "MLI", "NAM", "NER", "SLE")

# Load required libraries
library(magrittr)
library(dplyr)
library(tidyr)
library(terra)
library(sf) # For better spatial data handling

# Load data
df <- readRDS("data/data_rds/data_final.rds")
df %<>% filter(spl == 1)

# Unique ISO codes in data
country_codes <- unique(df$iso)

# Path to admin shapefiles
shapefile_path <- "data/data_raw/gadm/"

# Store each country's region vector - keeping as sf objects until the end
all_regions_sf <- list()

for (iso_code in country_codes) {
  # Skip if not in our specified lists
  if (!(iso_code %in% c(countries_adm1, countries_adm2))) {
    message("⛔ Skipping ", iso_code, " - not in specified admin level lists.")
    next
  }

  # Determine admin level for this country
  admin_level <- if(iso_code %in% countries_adm1) 1 else 2

  # Build shapefile path based on admin level
  shp_file <- paste0(shapefile_path, "gadm41_", iso_code, "_", admin_level, ".shp")

  if (!file.exists(shp_file)) {
    message("⛔ Skipping ", iso_code, " - shapefile missing at admin level ", admin_level)
    next
  }

  message("✅ Processing ", iso_code, " at admin level ", admin_level)

  # Load shapefile with sf for more reliable handling
  adm_sf <- sf::st_read(shp_file, quiet = TRUE)

  # Get the appropriate region name field based on admin level
  region_field <- paste0("NAME_", admin_level)

  # Make sure the field exists in the shapefile
  if(!(region_field %in% names(adm_sf))) {
    message("⚠️ Field ", region_field, " not found in shapefile for ", iso_code)
    next
  }

  # Filter country data and only include rounds 5 and 8
  df_country <- df %>% filter(iso == iso_code, round %in% c(5, 8))

  if (nrow(df_country) == 0) {
    message("⚠️ No data for rounds 5 or 8 for ", iso_code)
    adm_sf$variation <- NA_real_
    all_regions_sf[[iso_code]] <- adm_sf
    next
  }

  # Ensure we're using the correct region field in the data
  # Check if fields exist in df_country
  if(!(region_field %in% names(df_country))) {
    message("⚠️ Field ", region_field, " not found in data for ", iso_code)
    adm_sf$variation <- NA_real_
    all_regions_sf[[iso_code]] <- adm_sf
    next
  }

  # Split by round
  df5 <- df_country %>% filter(round == 5)
  df8 <- df_country %>% filter(round == 8)

  # Find common regions between rounds
  common_regions <- intersect(df5[[region_field]], df8[[region_field]])

  if (length(common_regions) == 0) {
    message("⚠️ No common regions for ", iso_code)
    adm_sf$variation <- NA_real_
    all_regions_sf[[iso_code]] <- adm_sf
    next
  }

  # Filter to common regions
  df_common <- df_country %>% filter(!!sym(region_field) %in% common_regions)

  # Debug: Print a sample of region names to verify structure
  message("Sample regions for ", iso_code, ": ", 
          paste(head(unique(df_common[[region_field]]), 3), collapse=", "))

  # Aggregate and calculate mean trust values
  trust_wide <- df_common %>%
    group_by(!!sym(region_field), round) %>%
    summarise(trust = mean(trust_parl_a, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = round, values_from = trust, names_prefix = "R") %>%
    mutate(variation = R8 - R5)

  # Debug: Check trust_wide structure
  message("Trust data structure: ", nrow(trust_wide), " rows with variation calculated")

  # Convert region names to character for reliable joining
  adm_sf[[region_field]] <- as.character(adm_sf[[region_field]])
  trust_wide[[region_field]] <- as.character(trust_wide[[region_field]])

  # Left join instead of merge for more reliable behavior
  joined <- tryCatch({
    sf::st_join(adm_sf, trust_wide, by = region_field, join = st_equals)
  }, error = function(e) {
    # Try a different join method if st_equals fails
    tryCatch({
      left_join(adm_sf, trust_wide, by = region_field)
    }, error = function(e2) {
      message("❌ Join failed for ", iso_code, ": ", e2$message)
      message("SHP fields: ", paste(head(names(adm_sf)), collapse=", "))
      message("Data fields: ", paste(head(names(trust_wide)), collapse=", "))
      NULL
    })
  })

  if(!is.null(joined)) {
    all_regions_sf[[iso_code]] <- joined
    # Debug: Check for NAs in variation
    na_count <- sum(is.na(joined$variation))
    total_count <- nrow(joined)
    message("Variation has ", na_count, " NAs out of ", total_count, " regions")
  }
}

# Only combine non-NULL elements
valid_regions <- all_regions_sf[!sapply(all_regions_sf, is.null)]

if(length(valid_regions) > 0) {
  # Check if we can rbind these objects
  if(length(valid_regions) > 1) {
    # Make sure all objects have compatible columns
    all_cols <- unique(unlist(lapply(valid_regions, names)))
    for(i in seq_along(valid_regions)) {
      missing_cols <- setdiff(all_cols, names(valid_regions[[i]]))
      for(col in missing_cols) {
        valid_regions[[i]][[col]] <- NA
      }
    }
  }

  # Now bind them safely
  combined <- do.call(rbind, valid_regions)

  # Convert back to terra format if needed
  combined_vect <- terra::vect(combined)
  combined_df <- as.data.frame(combined_vect)

  # Save or continue processing
} else {
  message("No valid data to combine")
}

capitals <- df %>%
  filter(round %in% c(5, 8)) %>%
  distinct(caplong, caplat, iso) %>%
  na.omit() %>%
  rename(longitude = caplong, latitude = caplat)

africa_outline <- ne_countries(continent = "Africa", returnclass = "sf") %>%
  st_make_valid()

africa_sf <- st_as_sf(combined)

ggplot() +
  geom_sf(data = africa_sf, aes(fill = variation), color = "white", linewidth = 0.1) +
  geom_sf(data = africa_outline, fill = NA, color = "black", linewidth = 0.2) +
  geom_point(data = capitals, aes(x = longitude, y = latitude),
             shape = 21, size = 2, fill = "black") +
  scale_fill_gradient2(
    low = "red", mid = "white", high = "green",
    midpoint = 0, limits = c(-3, 3),
    na.value = "lightgrey",
    name = "Δ Trust\n(R8 - R5)"
  ) +
  theme_minimal() +
  labs(
    title = "Trust Change in Parliament (Round 8 - Round 5)",
    subtitle = "Across African Regions",
    caption = "Grey = No data"
  )
```

```{r}
# Define which countries use which admin level
countries_adm1 <- c("GHA","KEN","MOZ", "MWI","NGA","TZA","UGA","ZMB", "ZWE")
countries_adm2 <- c("BEN", "BFA", "BWA", "CIV", "CMR", "GIN", "LBR", "MLI", "NAM", "NER", "SLE")

# Load required libraries
library(magrittr)
library(dplyr)
library(tidyr)
library(terra)
library(sf)

# Load data
df <- readRDS("data/data_rds/data_final.rds")
df %<>% filter(spl == 1)

# Unique ISO codes in data
country_codes <- unique(df$iso)

# Path to admin shapefiles
shapefile_path <- "data/data_raw/gadm/"

# Store each country's region vector
all_regions_sf <- list()

for (iso_code in country_codes) {
  # Skip if not in our specified lists
  if (!(iso_code %in% c(countries_adm1, countries_adm2))) {
    message("⛔ Skipping ", iso_code, " - not in specified admin level lists.")
    next
  }

  # Determine admin level for this country
  admin_level <- if(iso_code %in% countries_adm1) 1 else 2

  # Build shapefile path based on admin level
  shp_file <- paste0(shapefile_path, "gadm41_", iso_code, "_", admin_level, ".shp")

  if (!file.exists(shp_file)) {
    message("⛔ Skipping ", iso_code, " - shapefile missing at admin level ", admin_level)
    next
  }

  message("✅ Processing ", iso_code, " at admin level ", admin_level)

  # Load shapefile with sf
  adm_sf <- sf::st_read(shp_file, quiet = TRUE)

  # Get the appropriate region name field based on admin level
  region_field <- paste0("NAME_", admin_level)

  # Make sure the field exists in the shapefile
  if(!(region_field %in% names(adm_sf))) {
    message("⚠️ Field ", region_field, " not found in shapefile for ", iso_code)
    next
  }

  # Filter country data to include rounds 5-8
  df_country <- df %>% filter(iso == iso_code, round >= 5, round <= 8)

  if (nrow(df_country) == 0) {
    message("⚠️ No data for rounds 5-8 for ", iso_code)
    adm_sf$variation <- NA_real_
    adm_sf$min_round <- NA_integer_
    adm_sf$max_round <- NA_integer_
    all_regions_sf[[iso_code]] <- adm_sf
    next
  }

  # Ensure we're using the correct region field in the data
  if(!(region_field %in% names(df_country))) {
    message("⚠️ Field ", region_field, " not found in data for ", iso_code)
    adm_sf$variation <- NA_real_
    adm_sf$min_round <- NA_integer_
    adm_sf$max_round <- NA_integer_
    all_regions_sf[[iso_code]] <- adm_sf
    next
  }

  # Calculate average trust by region and round
  region_trust <- df_country %>%
    group_by(!!sym(region_field), round) %>%
    summarise(trust = mean(trust_parl_a, na.rm = TRUE), .groups = "drop")

  # Find regions with at least 2 rounds of data
  region_round_counts <- region_trust %>%
    group_by(!!sym(region_field)) %>%
    summarise(num_rounds = n(), .groups = "drop")

  regions_with_multiple_rounds <- region_round_counts %>%
    filter(num_rounds >= 2) %>%
    pull(!!sym(region_field))

  if (length(regions_with_multiple_rounds) == 0) {
    message("⚠️ No regions with multiple rounds for ", iso_code)
    adm_sf$variation <- NA_real_
    adm_sf$min_round <- NA_integer_
    adm_sf$max_round <- NA_integer_
    all_regions_sf[[iso_code]] <- adm_sf
    next
  }

  # Filter to regions with multiple rounds
  region_trust_filtered <- region_trust %>%
    filter(!!sym(region_field) %in% regions_with_multiple_rounds)

  # For each region, calculate variation between earliest and latest available round
  variation_data <- region_trust_filtered %>%
    group_by(!!sym(region_field)) %>%
    summarise(
      min_round = min(round),
      max_round = max(round),
      earliest_trust = trust[which.min(round)],
      latest_trust = trust[which.max(round)],
      variation = latest_trust - earliest_trust,
      .groups = "drop"
    )

  # Debug info
  message("Found ", nrow(variation_data), " regions with variations for ", iso_code)

  # Convert region names to character for reliable joining
  adm_sf[[region_field]] <- as.character(adm_sf[[region_field]])
  variation_data[[region_field]] <- as.character(variation_data[[region_field]])

  # Join variation data with shapefile
  joined <- left_join(adm_sf, variation_data, by = region_field)

  if(!is.null(joined)) {
    all_regions_sf[[iso_code]] <- joined
    # Debug: Check for NAs in variation
    na_count <- sum(is.na(joined$variation))
    total_count <- nrow(joined)
    message("Variation has ", na_count, " NAs out of ", total_count, " regions")
  }
}

# Only combine non-NULL elements
valid_regions <- all_regions_sf[!sapply(all_regions_sf, is.null)]

if(length(valid_regions) > 0) {
  # Ensure column compatibility
  all_cols <- unique(unlist(lapply(valid_regions, names)))
  for(i in seq_along(valid_regions)) {
    missing_cols <- setdiff(all_cols, names(valid_regions[[i]]))
    for(col in missing_cols) {
      valid_regions[[i]][[col]] <- NA
    }
  }

  # Combine all regions
  combined <- do.call(rbind, valid_regions)

  # Convert back to terra format if needed
  combined_vect <- terra::vect(st_cast(combined))
  combined_df <- as.data.frame(combined_vect)

  # Summary of variations
  message("Variation summary:")
  message("Min: ", min(combined$variation, na.rm = TRUE))
  message("Max: ", max(combined$variation, na.rm = TRUE))
  message("Mean: ", mean(combined$variation, na.rm = TRUE))
  message("Median: ", median(combined$variation, na.rm = TRUE))
  message("NA count: ", sum(is.na(combined$variation)))

  # You can now proceed with further analysis or visualization
} else {
  message("No valid data to combine")
}

capitals <- df %>%
  filter(round %in% c(5, 8)) %>%
  distinct(caplong, caplat, iso) %>%
  na.omit() %>%
  rename(longitude = caplong, latitude = caplat)

africa_outline <- ne_countries(continent = "Africa", returnclass = "sf") %>%
  st_make_valid()

africa_outline <- st_transform(africa_outline, st_crs(combined))

africa_sf <- st_as_sf(combined)

ggplot() +
  geom_sf(data = africa_sf, aes(fill = variation), color = "white", linewidth = 0.1) +
  geom_sf(data = africa_outline, fill = NA, color = "black", linewidth = 0.2) +
    geom_point(data = capitals, aes(x = longitude, y = latitude),
             shape = 21, size = 2, fill = "black") +
  scale_fill_gradient2(
    low = "red", high = "green",
    midpoint = 0, limits = c(-2, 2),
    na.value = "lightgrey",
    name = "Δ Trust\n(R8 - R5)"
  ) +
  theme_minimal() +
  labs(
    title = "Trust Change in Parliament (Round 8 - Round 5)",
    subtitle = "Across African Regions",
    caption = "Grey = No data"
  )
```